---
title: "ccRCC_Tissue"
author: "Christina Schmidt"
date: "March 2026"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
editor_options:
  chunk_output_type: console
---
```{r, warning=FALSE, message=FALSE}
#load the needed:
library(tidyverse) # used for data manipulation
library(rmarkdown) # used for paged_table function
library(dplyr)
library(OmnipathR)
library(MetaProViz)
```

# Samples
The [paper](https://www.cell.com/cancer-cell/comments/S1535-6108(15)00468-7#supplementaryMaterial) "An Integrated Metabolic Atlas of Clear Cell Renal Cell Carcinoma" includes metabolomic profiling on 138 matched clear cell renal cell carcinoma (ccRCC)/normal tissue pairs. Metabolomics was done using The company Metabolon, so this is untargeted metabolomics.\
\
Here we use the median normalised data from the supplementary table 2 of the paper. The median normalisation was done by metaboanalyst.

# The Data
## Load Data
```{r}
data <- readxl::read_excel("Input/Supplementarytable2_mmc2.xlsx", sheet = "Median Normalized")

#Patient MetaData:
MetaData_Sample <- data[c(1:12),-c(1:11)]%>%
  {colnames(.) <- .[12, ]
    .[-12, ]
    }%>%
  column_to_rownames("Group HMDB")
MetaData_Sample <- as.data.frame(t(MetaData_Sample))%>%
    mutate(STAGE = case_when(`TYPE-STAGE`=="TUMOR-STAGE I"  ~ 'EARLY-STAGE',
                             `TYPE-STAGE`=="TUMOR-STAGE II"  ~ 'EARLY-STAGE',
                             `TYPE-STAGE`=="TUMOR-STAGE III"  ~ 'LATE-STAGE',
                             `TYPE-STAGE`=="TUMOR-STAGE IV"  ~ 'LATE-STAGE',
                             `TYPE-STAGE`=="NORMAL-STAGE I"  ~ 'EARLY-STAGE',
                             `TYPE-STAGE`=="NORMAL-STAGE II"  ~ 'EARLY-STAGE',
                             `TYPE-STAGE`=="NORMAL-STAGE III"  ~ 'LATE-STAGE',
                             `TYPE-STAGE`=="NORMAL-STAGE IV"  ~ 'LATE-STAGE',
                                  TRUE ~ 'X'))%>%
    mutate(AGE = case_when(`AGE AT SURGERY`<42  ~ 'Young',
                           `AGE AT SURGERY`>58  ~ 'Old',
                                  TRUE ~ 'Middle'))%>%
    mutate(GENDER = case_when(`GENDER`=="Male"  ~ 'Male',
                           `GENDER`=="male"  ~ 'Male',
                           `GENDER`=="male "  ~ 'Male',
                           `GENDER`=="Female"  ~ 'Female',
                           `GENDER`=="female"  ~ 'Female',
                                  TRUE ~ 'x'))

names(MetaData_Sample) <- gsub(" ", "_", names(MetaData_Sample))

#Metabolite_MetaData
MetaData_Metabolite <-  data[-c(1:11),c(2:12)]%>%
  {colnames(.) <- .[1, ]
    .[-1, ]
  }%>%
  column_to_rownames("BIOCHEMICAL NAME")
names(MetaData_Metabolite ) <- gsub(" ", "_", names(MetaData_Metabolite ))


#Data
Data <-  data[-c(1:11),-c(1, 3:12)]%>%
  {colnames(.) <- .[1, ]
    .[-1, ]
  }%>%
  column_to_rownames("BIOCHEMICAL NAME")
Data <-as.data.frame(t(Data))%>%
  mutate(across(everything(), as.numeric))

```


## Patient sample and value DF
First we removed metabolites that where not identifiable from the dataset.\
```{r, warning=FALSE}
#Remove not identified metabolites:
Data <- Data[, !grepl("^X - ", colnames(Data))]

```
`
Next we remove constant/zero variance columns from Data
```{r, warning=FALSE}
#Remove constant/zero variance columns from Data
Data_Select <- Data[, colSums(Data != 0) > 0]
Data_Select <- Data_Select[, apply(Data_Select, 2, var) > 0]

#Metabolites removed:
Metabolites_removed <- setdiff(colnames(Data), colnames(Data_Select))
message("Metabolites removed due to zero-variance or constant variance: ", paste0(Metabolites_removed, collapse = ", "))

```
\
Now we check the clustering via PCA plot:\
In detail, we want to check for outliers, how the patients samples and normal tissue samples cluster and if the different demographics separate. \
```{r, warning=FALSE}
#Make plot:
MetaProViz::viz_pca(metadata_info= c(color="TISSUE_TYPE"),
                   metadata_sample = MetaData_Sample ,
                   data= Data_Select,
                   plot_name = "Outlier",
                   save_plot=NULL)
```
\
Merge Data with MetaData_Sample
```{r}
Data_Sample <- merge( MetaData_Sample, Data_Select, by = "row.names")%>%
  column_to_rownames("Row.names")

```

## Feature metadata table
Data wrangling
```{r, warning=FALSE}
MetaboliteIDs <- MetaData_Metabolite %>%
  tibble::rownames_to_column("Metabolite") %>%
  dplyr::rename(HMDB = Group_HMDB) %>%
  dplyr::mutate(
    PUBCHEM = as.integer(as.numeric(PUBCHEM)),
    KEGG = trimws(KEGG),
    HMDB = toupper(trimws(HMDB)),
    CAS = trimws(CAS)
  ) %>%
  tidyr::separate_rows(CAS, sep = ";") %>%
  dplyr::mutate(CAS = trimws(CAS)) %>%
  dplyr::filter(is.na(CAS) | CAS != "")

```
\
There are PubChem entries in the excel file that are not PubChem IDs. It is likely that this is an error caused by excel conversion as it seems in some cases the first part of the ID corresponds to one of the other IDs, so maybe these where multiple potential PubChem IDs.\
First we will extract the PubChem IDs that are present in the metadata table and check if they are valid PubChem IDs. We can do this using the `ramp_id_mapping_table` function, which includes all the valid PubChem IDs that are present in RaMP, which is a database that integrates multiple databases including KEGG, HMDB and PubChem.\
```{r}
#Get RaMP table
RaMP_PubChem <- OmnipathR::ramp_id_mapping_table(from = "pubchem",
                                         to= "pubchem")
# Get RaMP mappings for alternative ID types
RaMP_KEGG_PubChem <- OmnipathR::ramp_id_mapping_table(from = "kegg", 
                                                       to = "pubchem")
RaMP_HMDB_PubChem <- OmnipathR::ramp_id_mapping_table(from = "hmdb", 
                                                       to = "pubchem")
RaMP_CAS_PubChem <- OmnipathR::ramp_id_mapping_table(from = "cas", 
                                                      to = "pubchem")

##------------------------------------------------------------------------
#Check which PubChem IDs in our metadata are valid:
MetaboliteIDs_PUBCHEM <- MetaboliteIDs%>%
  dplyr::filter(!is.na(PUBCHEM)) %>%
  dplyr::mutate(PUBCHEM_valid = PUBCHEM %in% RaMP_PubChem$From)

MetaboliteIDs_PUBCHEM_FALSE <- MetaboliteIDs_PUBCHEM %>%
  dplyr::filter(PUBCHEM_valid == FALSE)

###------------------------------------------------------------------------
# Try KEGG first, then HMDB, then CAS as fallback
MetaboliteIDs_PUBCHEM_FALSE <- MetaboliteIDs_PUBCHEM_FALSE %>%
  dplyr::left_join(
    RaMP_KEGG_PubChem %>% 
      dplyr::select(From, To) %>% 
      dplyr::rename(KEGG = From, PUBCHEM_from_KEGG = To),
    by = "KEGG",
    relationship = "many-to-many"
  ) %>%
  dplyr::left_join(
    RaMP_HMDB_PubChem %>% 
      dplyr::select(From, To) %>% 
      dplyr::rename(HMDB = From, PUBCHEM_from_HMDB = To),
    by = "HMDB",
    relationship = "many-to-many"
  ) %>%
  tidyr::separate_rows(CAS, sep = ";") %>%
  dplyr::mutate(CAS = trimws(CAS)) %>%
  dplyr::filter(CAS != "") %>%
  dplyr::left_join(
    RaMP_CAS_PubChem %>% 
      dplyr::select(From, To) %>% 
      dplyr::rename(CAS = From, PUBCHEM_from_CAS = To),
    by = "CAS",
    relationship = "many-to-many"
  ) %>%
  dplyr::mutate(
    PUBCHEM_corrected = as.integer(
      dplyr::coalesce(PUBCHEM_from_KEGG, PUBCHEM_from_HMDB, PUBCHEM_from_CAS)
    )
  ) %>%
  dplyr::select(-PUBCHEM_from_KEGG, -PUBCHEM_from_HMDB, -PUBCHEM_from_CAS) %>%
  dplyr::group_by(Metabolite) %>%
  dplyr::summarise(
    PUBCHEM_corrected = paste(unique(PUBCHEM_corrected), collapse = "; "),
    CAS = paste(unique(CAS), collapse = "; "),
    dplyr::across(-c(PUBCHEM_corrected, CAS), dplyr::first),
    .groups = "drop"
  )

```

For the final PubChem IDs we will check manually:
1. 64947 = Correct (Via Name and CAS)
2. 5656 = correct (Via Name and CAS)
3. 170746 = 6209 (Via CAS)
4. 761 = 135414246 (Via CAS)
5. 5999 = 19648 (Via CAS)
6. 139506 = 139506 (Via Name)
7. 5057430 = 70590 (Via CAS), 5057430 (Via Name) --> Structurally very similar

```{r}
# Manual corrections: keyed by the current PUBCHEM_corrected value
manual_corrections <- tibble::tibble(
  PUBCHEM = c(64947L, 5656L, 170746L, 761L, 5999L, 139506L, 5057430L),
  PUBCHEM_manual = c("64947", "5656", "6209", "135414246", "19648", "139506", "70590; 5057430")
)

MetaboliteIDs_PUBCHEM_FALSE <- MetaboliteIDs_PUBCHEM_FALSE %>%
  dplyr::left_join(manual_corrections, by = "PUBCHEM") %>%
  dplyr::mutate(
    PUBCHEM_corrected = dplyr::case_when(
      !is.na(PUBCHEM_manual) ~ PUBCHEM_manual,
      !is.na(PUBCHEM_corrected) ~ as.character(PUBCHEM_corrected),
      TRUE ~ NA_character_
    ),
    PUBCHEM_Correction_Source = dplyr::case_when(
      !is.na(PUBCHEM_manual) ~ "Manual",
      !is.na(PUBCHEM_corrected) ~ "RaMP-DB",
      TRUE ~ "Not found"
    )
  ) %>%
  dplyr::select(-PUBCHEM_manual)


##-------------------------------------------------
#Add into original table:
# Add corrected invalid rows back into the original table
MetaboliteIDs_PUBCHEM <- MetaboliteIDs %>%
  dplyr::mutate(
    PUBCHEM_valid = !is.na(PUBCHEM) & PUBCHEM %in% RaMP_PubChem$From
  ) %>%
  dplyr::left_join(
    MetaboliteIDs_PUBCHEM_FALSE %>%
      dplyr::select(Metabolite, PUBCHEM, PUBCHEM_corrected, PUBCHEM_Correction_Source),
    by = c("Metabolite", "PUBCHEM")
  ) %>%
  dplyr::mutate(
    PUBCHEM = as.character(PUBCHEM),
    PUBCHEM_corrected = dplyr::case_when(
      PUBCHEM_valid ~ NA_character_,                      # valid original, no correction needed
      !is.na(PUBCHEM_corrected) ~ PUBCHEM_corrected,      # corrected via RaMP/manual
      TRUE ~ NA_character_
    ),
    PUBCHEM_Correction_Source = dplyr::case_when(
      is.na(PUBCHEM) ~ "NA",                              # no original PubChem in metadata
      PUBCHEM_valid ~ "None",                             # original ID already valid
      !is.na(PUBCHEM_Correction_Source) ~ PUBCHEM_Correction_Source,
      TRUE ~ "NA"
    ),
    # final PubChem to use
    PUBCHEM_corrected = dplyr::coalesce(PUBCHEM_corrected, PUBCHEM)
  ) %>%
  dplyr::rename(
    `PUBCHEM original` = PUBCHEM,
    PUBCHEM = PUBCHEM_corrected
  )%>%
  dplyr::group_by(Metabolite) %>%
  dplyr::summarise(
    CAS = paste(unique(stats::na.omit(trimws(CAS))), collapse = "; "),
    dplyr::across(-CAS, dplyr::first),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    CAS = dplyr::na_if(CAS, "")
  )
```

Some details on how many IDs where affected:
```{r}
# Quick counts from your final table
n_total <- nrow(MetaboliteIDs_PUBCHEM)

n_false <- MetaboliteIDs_PUBCHEM %>%
  dplyr::summarise(n = sum(PUBCHEM_valid == FALSE, na.rm = TRUE)) %>%
  dplyr::pull(n)

n_ramp <- MetaboliteIDs_PUBCHEM %>%
  dplyr::summarise(n = sum(PUBCHEM_Correction_Source == "RaMP-DB", na.rm = TRUE)) %>%
  dplyr::pull(n)

n_manual <- MetaboliteIDs_PUBCHEM %>%
  dplyr::summarise(n = sum(PUBCHEM_Correction_Source == "Manual", na.rm = TRUE)) %>%
  dplyr::pull(n)

n_not_found <- MetaboliteIDs_PUBCHEM %>%
  dplyr::summarise(n = sum(PUBCHEM_Correction_Source == "Not found", na.rm = TRUE)) %>%
  dplyr::pull(n)

n_with_pubchem_final <- MetaboliteIDs_PUBCHEM %>%
  dplyr::summarise(n = sum(!is.na(PUBCHEM) & trimws(PUBCHEM) != "", na.rm = TRUE)) %>%
  dplyr::pull(n)

cat("Total features:", n_total, "\n")
cat("Initially invalid PubChem (PUBCHEM_valid == FALSE):", n_false, "\n")
cat("Corrected using RaMP-DB:", n_ramp, "\n")
cat("Corrected manually:", n_manual, "\n")
cat("Invalid and still not found:", n_not_found, "\n")
cat("Features with final PubChem ID:", n_with_pubchem_final, "\n")

```

## Save
Save DF as .csv file
```{r}
write.csv(Data_Sample, "Hakimi_ccRCC-Tissue_Data.csv")
write.csv(MetaboliteIDs_PUBCHEM, "Hakimi_ccRCC-Tissue_FeatureMetaData_Full-Pubchem-Info.csv")

tissue_meta <- MetaboliteIDs_PUBCHEM %>%
  dplyr::select(1:8, 13, 10:11)

write.csv(tissue_meta, "Hakimi_ccRCC-Tissue_FeatureMetaData.csv")
save(tissue_meta, file = "tissue_meta.rda")

```

# Information about package used and versions 
```{r session_info, echo=FALSE}
options(width = 120)
sessioninfo::session_info()
```
