---
title: "ccRCC_Tissue"
author: "Christina Schmidt"
date: "July 2024"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
editor_options:
  chunk_output_type: console
---
```{r, warning=FALSE, message=FALSE}
#load the needed:
library(tidyverse) # used for data manipulation
library(rmarkdown) # used for paged_table function
```

# Samples
The [paper](https://www.cell.com/cancer-cell/comments/S1535-6108(15)00468-7#supplementaryMaterial) "An Integrated Metabolic Atlas of Clear Cell Renal Cell Carcinoma" includes metabolomic profiling on 138 matched clear cell renal cell carcinoma (ccRCC)/normal tissue pairs. Metabolomics was done using The company Metabolon, so this is untargeted metabolomics.\
\
Here we use the median normalised data from the supplementary table 2 of the paper. The median normalisation was done by metaboanalyst.

# The Data
```{r}
data <- readxl::read_excel("Input/Supplementarytable2_mmc2.xlsx", sheet = "Median Normalized")

#Patient MetaData:
MetaData_Sample <- data[c(1:12),-c(1:11)]%>%
  {colnames(.) <- .[12, ]
    .[-12, ]
    }%>%
  column_to_rownames("Group HMDB")
MetaData_Sample <- as.data.frame(t(MetaData_Sample))%>%
    mutate(STAGE = case_when(`TYPE-STAGE`=="TUMOR-STAGE I"  ~ 'EARLY-STAGE',
                             `TYPE-STAGE`=="TUMOR-STAGE II"  ~ 'EARLY-STAGE',
                             `TYPE-STAGE`=="TUMOR-STAGE III"  ~ 'LATE-STAGE',
                             `TYPE-STAGE`=="TUMOR-STAGE IV"  ~ 'LATE-STAGE',
                             `TYPE-STAGE`=="NORMAL-STAGE I"  ~ 'EARLY-STAGE',
                             `TYPE-STAGE`=="NORMAL-STAGE II"  ~ 'EARLY-STAGE',
                             `TYPE-STAGE`=="NORMAL-STAGE III"  ~ 'LATE-STAGE',
                             `TYPE-STAGE`=="NORMAL-STAGE IV"  ~ 'LATE-STAGE',
                                  TRUE ~ 'X'))%>%
    mutate(AGE = case_when(`AGE AT SURGERY`<42  ~ 'Young',
                           `AGE AT SURGERY`>58  ~ 'Old',
                                  TRUE ~ 'Middle'))%>%
    mutate(GENDER = case_when(`GENDER`=="Male"  ~ 'Male',
                           `GENDER`=="male"  ~ 'Male',
                           `GENDER`=="male "  ~ 'Male',
                           `GENDER`=="Female"  ~ 'Female',
                           `GENDER`=="female"  ~ 'Female',
                                  TRUE ~ 'x'))

names(MetaData_Sample) <- gsub(" ", "_", names(MetaData_Sample))

#Metabolite_MetaData
MetaData_Metabolite <-  data[-c(1:11),c(2:12)]%>%
  {colnames(.) <- .[1, ]
    .[-1, ]
  }%>%
  column_to_rownames("BIOCHEMICAL NAME")
names(MetaData_Metabolite ) <- gsub(" ", "_", names(MetaData_Metabolite ))


#Data
Data <-  data[-c(1:11),-c(1, 3:12)]%>%
  {colnames(.) <- .[1, ]
    .[-1, ]
  }%>%
  column_to_rownames("BIOCHEMICAL NAME")
Data <-as.data.frame(t(Data))%>%
  mutate(across(everything(), as.numeric))

```
\
First we removed metabolites that where not identifiable from the dataset.\
```{r, warning=FALSE}
#Remove not identified metabolites:
Data <- Data[, !grepl("^X - ", colnames(Data))]

```
`
Next we remove constant/zero variance columns from Data
```{r, warning=FALSE}
#Remove constant/zero variance columns from Data
Data_Select <- Data[, colSums(Data != 0) > 0]
Data_Select <- Data_Select[, apply(Data_Select, 2, var) > 0]

#Metabolites removed:
Metabolites_removed <- setdiff(colnames(Data), colnames(Data_Select))
message("Metabolites removed due to zero-variance or constant variance: ", paste0(Metabolites_removed, collapse = ", "))

```
\
Now we check the clustering via PCA plot:\
In detail, we want to check for outliers, how the patients samples and normal tissue samples cluster and if the different demographics separate. \
```{r, warning=FALSE}
#Make plot:
MetaProViz::VizPCA(SettingsInfo= c(color="TISSUE_TYPE"),
                   SettingsFile_Sample = MetaData_Sample ,
                   InputData= Data_Select,
                   PlotName = "Outlier",
                   SaveAs_Plot=NULL)
```
\
Merge Data with MetaData_Sample
```{r}
Data_Sample <- merge( MetaData_Sample, Data_Select, by = "row.names")%>%
  column_to_rownames("Row.names")

```
\
Save DF as .csv file
```{r}
write.csv(Data_Sample, "Hakimi_ccRCC-Tissue_Data.csv")
write.csv(MetaData_Metabolite, "Hakimi_ccRCC-Tissue_FeatureMetaData.csv")
```

# Information about package used and versions 
```{r session_info, echo=FALSE}
options(width = 120)
sessioninfo::session_info()
```
