---
title: "Create_SE_Input"
author: "Christina Schmidt, Jannik Franken, Giulia Roeth"
date: "September 2026"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
editor_options:
  chunk_output_type: console
---
```{r, warning=FALSE, message=FALSE}
#load the needed:
library(tidyverse) # used for data manipulation
library(rmarkdown) # used for paged_table function
library(dplyr)
library(tibble)
library(SummarizedExperiment)
library(MetaProViz)
```

# Purpose
This notebook documents how the three `SummarizedExperiment` (SE) example
datasets are generated from the published `data()` objects, reusing
`MetaProViz:::build_se_from_df()` to assemble the assay, sample metadata,
and feature metadata into one object.\
\
`1.` `tissue_norm_se`: ccRCC tissue metabolomics, Hakimi et al.
[@Hakimi2016].\
`2.` `intracell_raw_se`: intracellular cell-line metabolomics,
metabolomics workbench project PR001418, study ST002224
[@Sciacovelli_Dugourd2022].\
`3.` `medium_raw_se`: exometabolomics (CoRe) of the same cell lines,
metabolomics workbench project PR001418, study ST002226.\
\
Full descriptions of each dataset, including download links, are given in
the corresponding section below.
```{r}
#align_feature_metadata() works around build_se_from_df() erroring when
#metadata_feature does not cover every feature in the assay: safe join,
#NA-padding the ones without an annotation entry:
align_feature_metadata <- function(assay_df, metadata_feature) {
  all_features <- colnames(assay_df)
  merge(
    data.frame(FeatureID = all_features, stringsAsFactors = FALSE),
    metadata_feature %>% tibble::rownames_to_column("FeatureID"),
    by = "FeatureID", all.x = TRUE
  ) %>%
    tibble::column_to_rownames("FeatureID") %>%
    .[all_features, , drop = FALSE]
}
```

# 1. `tissue_norm_se`
The [paper](https://www.cell.com/cancer-cell/comments/S1535-6108(15)00468-7#supplementaryMaterial)
"An Integrated Metabolic Atlas of Clear Cell Renal Cell Carcinoma" includes
metabolomic profiling on 138 matched clear cell renal cell carcinoma
(ccRCC)/normal tissue pairs [@Hakimi2016]. Metabolomics was done using the
company Metabolon, so this is untargeted metabolomics. Here we use the
median normalised data from supplementary table 2 of the paper.\
\
`1.` Sample information + median-normalised metabolite peaks **(tissue_norm)**\
`2.` Metabolite IDs (CAS, HMDB, KEGG, PubChem), pathway assignment and
measurement platform **(tissue_meta)**, built in `ccRCC_Hakimi-et-all.Rmd`.

## Load Data
```{r}
data(tissue_norm)
data(tissue_meta)

Tissue_Norm <- tissue_norm %>% column_to_rownames("Code")

sample_cols <- c("TISSUE_TYPE", "GENDER", "RACE", "STAGE", "AGE")
metab_cols  <- setdiff(colnames(Tissue_Norm), colnames(tissue_norm)[2:14])

assay_df    <- Tissue_Norm[, metab_cols]
meta_sample <- Tissue_Norm[, sample_cols]
meta_feat   <- tissue_meta %>%
  column_to_rownames("Metabolite") %>%
  align_feature_metadata(assay_df, .)
```

## Build the SE
```{r}
tissue_norm_se <- build_se_from_df(
  assay_df = assay_df,
  metadata_sample = meta_sample,
  metadata_feature = meta_feat
)

tissue_norm_se
```

## Feature-metadata coverage
```{r}
rd <- rowData(tissue_norm_se)
fully_na <- rownames(rd)[apply(as.data.frame(rd), 1, function(r) all(is.na(r)))]
cat("features with no tissue_meta entry:", length(fully_na), "of", nrow(rd), "\n")
```
All 570 features have a `tissue_meta` entry.

# 2. `intracell_raw_se`
Publicly available on [metabolomics workbench project PR001418](https://www.metabolomicsworkbench.org/data/DRCCMetadata.php?Mode=Project&ProjectID=PR001418)
including metabolic profiles of human renal epithelial cells HK2 and ccRCC
cell lines cultured in Plasmax cell culture media [@Sciacovelli_Dugourd2022].
The raw data are available via [metabolomics workbench study ST002224](https://www.metabolomicsworkbench.org/data/DRCCMetadata.php?Mode=Study&StudyID=ST002224&StudyType=MS&ResultType=1)
were intracellular metabolomics of HK2 and ccRCC cell lines 786-O, 786-M1A
and 786-M2A were performed.\
\
`1.` Sample information + raw integrated metabolite peaks **(intracell_raw)**\
`2.` Metabolite trivial name mapped to HMDB ID, KEGG ID and one main
metabolic pathway **(cellular_meta)**, shared with `medium_raw` below (same
project).

## Load Data
```{r}
data(intracell_raw)
data(cellular_meta)

Intra <- intracell_raw %>% column_to_rownames("Code")

sample_cols2 <- c("Conditions", "Analytical_Replicates", "Biological_Replicates")
metab_cols2  <- setdiff(colnames(Intra), sample_cols2)

assay_df2    <- Intra[, metab_cols2]
meta_sample2 <- Intra[, sample_cols2]
meta_feat2   <- cellular_meta %>%
  column_to_rownames("Metabolite") %>%
  align_feature_metadata(assay_df2, .)
```

## Build the SE
```{r}
intracell_raw_se <- build_se_from_df(
  assay_df = assay_df2,
  metadata_sample = meta_sample2,
  metadata_feature = meta_feat2
)

intracell_raw_se
```

## Feature-metadata coverage
```{r}
rd2 <- rowData(intracell_raw_se)
fully_na2 <- rownames(rd2)[apply(as.data.frame(rd2), 1, function(r) all(is.na(r)))]
cat("features with no cellular_meta entry:", length(fully_na2), "of", nrow(rd2), "\n")
cat("unmatched:", paste(fully_na2, collapse = ", "), "\n")
```
`valine-d8` and `hippuric acid-d5` are internal standards,
so their absence from a pathway/ID annotation table is expected.
`hexanoylcarnitine C6` is a coverage gap in `cellular_meta` (left as `NA`).
See Create_cellular_meta.Rmd to fill in gaps.

# 3. `medium_raw_se`
Same project as above, [metabolomics workbench project PR001418](https://www.metabolomicsworkbench.org/data/DRCCMetadata.php?Mode=Project&ProjectID=PR001418).
The raw data are available via [metabolomics workbench study ST002226](https://www.metabolomicsworkbench.org/data/DRCCMetadata.php?Mode=Study&StudyID=ST002226&StudyType=MS&ResultType=1)
were exometabolomics (consumption-release, "CoRe") of HK2 and ccRCC cell
lines 786-O, 786-M1A, 786-M2A, OS-RC-2, OS-LM1 and RFX-631 were performed.
\
`1.` Sample information + raw integrated metabolite peaks from the culture
media **(medium_raw)**\
`2.` Same feature metadata table as used for `intracell_raw_se` above
**(cellular_meta)**.

## Load Data
```{r}
data(medium_raw)
#cellular_meta already loaded above

Media <- medium_raw %>% column_to_rownames("Code")

sample_cols3 <- c("Conditions", "Biological_Replicates", "GrowthFactor")
metab_cols3  <- setdiff(colnames(Media), sample_cols3)

assay_df3    <- Media[, metab_cols3]
meta_sample3 <- Media[, sample_cols3]
meta_feat3   <- cellular_meta %>%
  column_to_rownames("Metabolite") %>%
  align_feature_metadata(assay_df3, .)
```

## Build the SE
```{r}
medium_raw_se <- build_se_from_df(
  assay_df = assay_df3,
  metadata_sample = meta_sample3,
  metadata_feature = meta_feat3
)

medium_raw_se
```

## Feature-metadata coverage
```{r}
rd3 <- rowData(medium_raw_se)
fully_na3 <- rownames(rd3)[apply(as.data.frame(rd3), 1, function(r) all(is.na(r)))]
cat("features with no cellular_meta entry:", length(fully_na3), "of", nrow(rd3), "\n")
cat("unmatched:", paste(fully_na3, collapse = ", "), "\n")
```
`valine-d8` and `hipppuric acid-d5` are internal standards, 
so their absence from a pathway/ID annotation table is expected.

# Save
Save as .rda files.
```{r}
save(tissue_norm_se, file = "tissue_norm_se.rda")
save(intracell_raw_se, file = "intracell_raw_se.rda")
save(medium_raw_se, file = "medium_raw_se.rda")
```

# Information about package used and versions
```{r session_info, echo=FALSE}
sessionInfo()
```
