---
title: "Create_cellular_meta"
author: "Giulia Roeth"
date: "September 2026"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
editor_options:
  chunk_output_type: console
---
```{r, warning=FALSE, message=FALSE}
#load the needed:
library(tidyverse) # used for data manipulation
library(rmarkdown) # used for paged_table function
library(dplyr)
library(tibble)
library(SummarizedExperiment)
library(MetaProViz)
```

# Purpose
`cellular_meta` is missing an entry for `hexanoylcarnitine C6` entirely (only
its unsaturated relative `hexenoylcarnitine C6:1` is present), which leaves
it unmatched in `intracell_raw_se` (see `Create_SE_Input.Rmd`). Investigation
into this gap found:\
\
`1.` The original source (metabolomics workbench study ST002224) lists this
compound under a different name, `carnitine C6` (WorkBench ID `ME538459`),
and it already has **no PubChem or KEGG ID** there either -- the original
submission never carried these identifiers.\
`2.` `cellular_meta` has no build script anywhere in this branch's history,
so there is no processing step that could have caused the gap.\
\
Since the missingness does not stem from anything MetaProViz did, this
notebook backfills `hexanoylcarnitine C6` into `cellular_meta`, using
`MetaProViz::translate_id()` (the RaMP-based ID-processing workflow) to look
up cross-referenced identifiers, and documents where those identifiers come
from and why `KEGG` is left empty.

# Keep the pre-backfill version
```{r}
data(cellular_meta)

cellular_meta_original <- cellular_meta
save(cellular_meta_original, file = "cellular_meta_original.rda")
```
`cellular_meta_original` is a fixed snapshot of `cellular_meta` as currently
shipped, kept as a stable baseline to diff future changes against -- unlike
an "_old" label, its meaning does not become ambiguous if `cellular_meta` is
backfilled again later.

# Look up IDs for `hexanoylcarnitine C6`
Seed ID: `HMDB0000756`. Metabolomics Workbench's RefMet database assigns
the standardized name `CAR 6:0` to this exact compound (the same identity
behind `ME538459`/`carnitine C6` in the ST002224 source table), and its
[RefMet entry](https://www.metabolomicsworkbench.org/databases/refmet/refmet_details.php?REFMET_NAME=CAR+6%3A0)
lists `PubChem CID 3246938`, `ChEBI 84834`, and `HMDB0000756`. ChEBI:84834
("O-hexanoyl-L-carnitine") independently confirms `HMDB0000756` as its own
HMDB cross-reference -- so all three sources triangulate on the same ID.
```{r}
Hexanoylcarnitine_Seed <- data.frame(
  MetaboliteID = "HMDB0000756",
  term = "hexanoylcarnitine C6",
  stringsAsFactors = FALSE
)

Hexanoylcarnitine_IDs <- translate_id(
  data = Hexanoylcarnitine_Seed,
  metadata_info = c(InputID = "MetaboliteID", grouping_variable = "term"),
  from = "hmdb",
  to = c("pubchem", "chebi"),
  summary = TRUE,
  save_table = NULL
)

Hexanoylcarnitine_IDs$TranslatedDF
```
Note: `to = "kegg"` was also attempted here, but `translate_id()` errors on
this seed ID (`object 'as.NULL' not found`, inside `OmnipathR`) instead of
returning an empty result -- this happens because RaMP has **no KEGG
mapping at all** for `HMDB0000756` (or its closely-related achiral parent,
`HMDB0000705` -- both give the same error). That failure is itself useful
confirmation: neither RaMP nor a direct KEGG COMPOUND search (checked
manually) has an entry for hexanoylcarnitine, so `KEGG.ID`/`KEGGCompound`
are left as `NA` below -- consistent with several other acylcarnitines
already in `cellular_meta` (e.g. `hexenoylcarnitine C6:1`,
`tetradecenoylcarnitine C14:1`).\
\
RaMP's graph returns two linked PubChem/ChEBI candidates regardless of
which of the two related HMDB IDs is used as seed: `3246938`/`84834` (the
L-stereoisomer, matching RefMet `CAR 6:0`/`HMDB0000756` used here) and
`6426853`/`70749` (its achiral parent compound, `HMDB0000705`). `cellular_meta`
has no `PubChem`/`ChEBI` columns (only `HMDB` and `KEGG.ID`/`KEGGCompound`),
so only `HMDB` is filled in below; both candidate pairs are documented here
for reference should the schema be extended later.

# Add the new row
```{r}
cellular_meta <- cellular_meta %>%
  bind_rows(
    data.frame(
      Metabolite = "hexanoylcarnitine C6",
      HMDB = "HMDB0000756",
      KEGG.ID = NA_character_,
      KEGGCompound = NA_character_,
      Pathway = "Fatty acyl carnitines",
      stringsAsFactors = FALSE
    )
  )

save(cellular_meta, file = "cellular_meta.rda")
```
`Pathway` is set to `"Fatty acyl carnitines"`, matching every other
acylcarnitine already in the table (e.g. `acetylcarnitine`,
`propionylcarnitine`, `hexenoylcarnitine C6:1`).

# Update `intracell_raw_se`
Rebuilt from the updated `cellular_meta` above, the same way as in
`Create_SE_Input.Rmd` (reusing `MetaProViz:::build_se_from_df()` and the
same `align_feature_metadata()` NA-padding helper). `medium_raw_se` is not
rebuilt here -- `medium_raw` does not measure `hexanoylcarnitine C6` at all,
so this change does not affect it.
```{r}
align_feature_metadata <- function(assay_df, metadata_feature) {
  all_features <- colnames(assay_df)
  merge(
    data.frame(FeatureID = all_features, stringsAsFactors = FALSE),
    metadata_feature %>% tibble::rownames_to_column("FeatureID"),
    by = "FeatureID", all.x = TRUE
  ) %>%
    tibble::column_to_rownames("FeatureID") %>%
    .[all_features, , drop = FALSE]
}

data(intracell_raw)

Intra <- intracell_raw %>% column_to_rownames("Code")

sample_cols2 <- c("Conditions", "Analytical_Replicates", "Biological_Replicates")
metab_cols2  <- setdiff(colnames(Intra), sample_cols2)

assay_df2    <- Intra[, metab_cols2]
meta_sample2 <- Intra[, sample_cols2]
meta_feat2   <- cellular_meta %>%
  column_to_rownames("Metabolite") %>%
  align_feature_metadata(assay_df2, .)

intracell_raw_se <- build_se_from_df(
  assay_df = assay_df2,
  metadata_sample = meta_sample2,
  metadata_feature = meta_feat2
)

intracell_raw_se
```

## Feature-metadata coverage after backfill
```{r}
rd2 <- rowData(intracell_raw_se)
fully_na2 <- rownames(rd2)[apply(as.data.frame(rd2), 1, function(r) all(is.na(r)))]
cat("features with no cellular_meta entry:", length(fully_na2), "of", nrow(rd2), "\n")
cat("unmatched:", paste(fully_na2, collapse = ", "), "\n")
```
`hexanoylcarnitine C6` is now matched; only `valine-d8` and
`hippuric acid-d5` (internal standards) remain unmatched, as expected.

## Save
```{r}
save(intracell_raw_se, file = "intracell_raw_se.rda")
```

# Information about package used and versions
```{r session_info, echo=FALSE}
sessionInfo()
```
