---
title: "SourceDataTable"
author:
  - name: Christina Schmidt
    affiliation:
    - Heidelberg University
output:
  html_document:
    self_contained: true
    toc: true
    toc_float: true
    toc_depth: 5
    code_folding: show
vignette: >
  %\VignetteIndexEntry{Standard Metabolomics}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmdarkdown}
bibliography: bibliography.bib
editor_options:
  chunk_output_type: console
---

<style>
.vscroll-plot {
    width: 850px;
    height: 500px;
    overflow-y: scroll;
    overflow-x: hidden;
}
</style>

```{r chunk_setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

install the required dependencies and load the libraries:
```{r, message=FALSE, warning=FALSE}
#dependencies that need to be loaded:
library(dplyr)
library(tidyverse)
library(readxl)
library(MetaProViz)
```

# Compile the source data tables for the figures
```{r}
#Create output folder
Output_folder <- file.path(getwd(), "SourceData")
if (!dir.exists(Output_folder)) {dir.create(Output_folder)} # Make Results folder
```

## Figure 1
Not applicable to Fig. 1a-b as Schematics.
```{r}
# Create needed output folders
Fig1 <- file.path(Output_folder, "Figure 1")
if (!dir.exists(Fig1)) {dir.create(Fig1)} 

Fig1C <- file.path(Fig1 , "1C")
if (!dir.exists(Fig1C)) {dir.create(Fig1C)}

Fig1D <- file.path(Fig1 , "1D")
if (!dir.exists(Fig1D)) {dir.create(Fig1D)}

#---------------------------------------------------------------------------------
# Add Figure Source data and save as .csv files within the respective folders:

# Fig1C
Table_EV1_Resource_Level <- read_excel("Table_EV1.xlsx", sheet = "Resource_Level")
write.csv(Table_EV1_Resource_Level , 
          paste0(Fig1C ,"/Fig.1c.csv"), 
          row.names = FALSE)

# Fig1D
Table_EV1_Term_Size_Distribution <- read_excel("Table_EV1.xlsx", sheet = "Term_Size_Distribution")
write.csv(Table_EV1_Term_Size_Distribution , 
          paste0(Fig1D ,"/Fig.1d.csv"), 
          row.names = FALSE)

```

## Figure 2
Not applicable to Fig 2D-E as Schematics.
```{r}
# Create needed output folders
Fig2 <- file.path(Output_folder, "Figure 2")
if (!dir.exists(Fig2)) {dir.create(Fig2)} 

Fig2A <- file.path(Fig2 , "2A")
if (!dir.exists(Fig2A)) {dir.create(Fig2A)}  

Fig2B <- file.path(Fig2 , "2B")
if (!dir.exists(Fig2B)) {dir.create(Fig2B)}  

Fig2C <- file.path(Fig2 , "2C")
if (!dir.exists(Fig2C)) {dir.create(Fig2C)}  


#---------------------------------------------------------------------------------
# Add Figure Source data and save as .csv files within the respective folders:

## Fig2A = ExtendedDataTable 1, sheet 2 "Feature metadata"

## Fig2B = ExtendedDataTable 2, sheet 2 "Alanine IDs"

## Fig2C = ExtendedDataTable 2, sheet 3 "Mapping ambiguities"

```

## Figure 3
Not applicable to Fig 3A as Schmatics
```{r}
# Create needed output folders
Fig3 <- file.path(Output_folder, "Figure 3")
if (!dir.exists(Fig3)) {dir.create(Fig3)} 

Fig3B <- file.path(Fig3 , "3B")
if (!dir.exists(Fig3B)) {dir.create(Fig3B)}  

Fig3C <- file.path(Fig3 , "3C")
if (!dir.exists(Fig3C)) {dir.create(Fig3C)}  

Fig3D <- file.path(Fig3 , "3D")
if (!dir.exists(Fig3D)) {dir.create(Fig3D)}

Fig3E <- file.path(Fig3 , "3E")
if (!dir.exists(Fig3E)) {dir.create(Fig3E)}

Fig3F <- file.path(Fig3 , "3F")
if (!dir.exists(Fig3F)) {dir.create(Fig3F)}

Fig3G <- file.path(Fig3 , "3G")
if (!dir.exists(Fig3G)) {dir.create(Fig3G)}

#---------------------------------------------------------------------------------
# Add Figure Source data and save as .csv files within the respective folders:

## Fig3B = Table EV3, sheet  1 "Raw data." --> needs data wrangling
data(intracell_raw)

CellPlot_Intra <- intracell_raw%>%
  filter(Conditions != "Pool")%>%
   mutate(Celltype = case_when(Conditions=="HK2" ~ 'Healthy',
                               Conditions=="786-O" ~ 'PrimaryTumour',
                               TRUE ~ 'MetastaticTumour'))

CellPlot_Intra$Count <- 1
CellPlot_Intra$CelltypeFactor <- factor(CellPlot_Intra$Celltype, levels = c("Healthy","PrimaryTumour", "MetastaticTumour"))

CellPlot_Intra_select <- CellPlot_Intra %>%
  select(Code, Conditions, Analytical_Replicates, Biological_Replicates, Celltype, Count)

write.csv(CellPlot_Intra_select, 
          paste0(Fig3B ,"/Fig.3b.csv"), 
          row.names = FALSE)

## Fig3C = Table EV3, sheet 2
PoolEstimationCV <- read_excel("Table_EV3.xlsx", sheet = "PoolEstimationCV")

write.csv(PoolEstimationCV, 
          paste0(Fig3C ,"/Fig.3c.csv"), 
          row.names = FALSE)

## Fig3D = Table EV3, sheet 3
MediaBlankCV <- read_excel("Table_EV3.xlsx", sheet = "MediaBlankCV")

write.csv(MediaBlankCV, 
          paste0(Fig3D ,"/Fig.3d.csv"), 
          row.names = FALSE)

## Fig3E = PCA Pool QC --> Raw data is used at this stage!
#Open: Pool_Estimation_result[["DF"]][["data"]] from BioRxiv Branch and run: 
PCA_RES <- prcomp(as.matrix(Pool_Estimation_result[["DF"]][["data"]]), scale. = as.logical(TRUE))

autoplot(PCA_RES)


## Fig3F = Before/After TIC
RawData <- read_excel("Table_EV3.xlsx", sheet = "RawData")
write.csv(RawData , 
          paste0(Fig3F ,"/Fig.3f_LeftPanel.csv"), 
          row.names = FALSE)

ProcessedData <- read_excel("Table_EV3.xlsx", sheet = "ProcessedData")
write.csv(ProcessedData%>%select(-(3:5)),
          paste0(Fig3F ,"/Fig.3f_RightPanel.csv"), 
          row.names = FALSE)

## Fig3G = PCA outliers --> Sheet 5, processed data is used with VizPCA!
ProcessedData$Outliers <- relevel(as.factor(ProcessedData$Outliers), ref = "no")

ProcessedData_Mat <- ProcessedData %>%
  select( -Conditions, -Biological_Replicates, -GrowthFactor, -Outliers)%>%
  column_to_rownames("Code")

PCA_RES <- prcomp(as.matrix(ProcessedData_Mat), scale. = as.logical(TRUE))

autoplot(PCA_RES)

pca_QC <- invisible(viz_pca(data = as.data.frame(ProcessedData_Mat),
                                    metadata_info = c(color = "Conditions", shape = "Outliers"),
                                    metadata_sample = ProcessedData[,c(1:5)]%>%column_to_rownames("Code"),
                                    plot_name = "Quality Control PCA Condition clustering and outlier check",
                                    save_plot = NULL))

# It is not the exact same --> What do I do here? 

write.csv(ProcessedData , 
          paste0(Fig3G ,"/Fig.3g.csv"), 
          row.names = FALSE)
```

## Figure 4
```{r}
# Create needed output folders
Fig4 <- file.path(Output_folder, "Figure 4")
if (!dir.exists(Fig4)) {dir.create(Fig4)} 

Fig4A <- file.path(Fig4 , "4A")
if (!dir.exists(Fig4A)) {dir.create(Fig4A)}

Fig4B <- file.path(Fig4 , "4B")
if (!dir.exists(Fig4B)) {dir.create(Fig4B)}  

Fig4C <- file.path(Fig4 , "4C")
if (!dir.exists(Fig4C)) {dir.create(Fig4C)}  

Fig4D <- file.path(Fig4 , "4D")
if (!dir.exists(Fig4D)) {dir.create(Fig4D)}

Fig4E <- file.path(Fig4 , "4E")
if (!dir.exists(Fig4E)) {dir.create(Fig4E)}

Fig4F <- file.path(Fig4 , "4F")
if (!dir.exists(Fig4F)) {dir.create(Fig4F)}

Fig4G <- file.path(Fig4 , "4G")
if (!dir.exists(Fig4G)) {dir.create(Fig4G)}

Fig4H <- file.path(Fig4 , "4H")
if (!dir.exists(Fig4H)) {dir.create(Fig4H)}

#---------------------------------------------------------------------------------
# Add Figure Source data and save as .csv files within the respective folders:

## Fig4A = Table_EV3
DMA_786O_vs_HK2<- read_excel("Table_EV3.xlsx", sheet = "DMA_786-O_vs_HK2")

write.csv(DMA_786O_vs_HK2, 
          paste0(Fig4A ,"/Fig.4a.csv"), 
          row.names = FALSE)

## Fig4B = Table_EV3
DMAFeatureMetadata<- read_excel("Table_EV3.xlsx", sheet = "DMA_FeatureMetadata")%>%
  select(-HMDB, -KEGG.ID, -KEGGCompound)%>%
  drop_na()

write.csv(DMAFeatureMetadata, 
          paste0(Fig4B ,"/Fig.4b.csv"), 
          row.names = FALSE)

## Fig4C = Table_EV4, sheet ""
ProcessedData <- read_excel("Table_EV3.xlsx", sheet = "ProcessedData")

# 1) Columns in ProcessedData that are sample metadata (not metabolites)
sample_cols <- c("Code", "Conditions", "Biological_Replicates", "GrowthFactor", "Outliers")

# 2) Convert metabolite columns to long format (one row = one heatmap cell)
heatmap_plot_df <- ProcessedData %>%
  pivot_longer(
    cols = -all_of(sample_cols),
    names_to = "Metabolite",
    values_to = "Abundance"
  ) %>%
  left_join(
    DMAFeatureMetadata %>%
      select(all_of(c("Metabolite", "Pathway", "core_HK2", "core_786-O", "core_786-M1A"))),
    by = "Metabolite"
  )%>%
  filter(Outliers != "Outlier_filtering_round_1") %>%
  filter(!Metabolite %in% c("valine-d8", "hipppuric acid-d5"))  %>%
  filter(Pathway == "Amino acid metabolism") %>%
  select(-Biological_Replicates, -GrowthFactor, -Outliers)

write.csv(heatmap_plot_df, 
          paste0(Fig4C ,"/Fig.4c.csv"), 
          row.names = FALSE)

## Fig4D
DMA_786O_vs_HK2_Intra <- read_excel("Table_EV4.xlsx", sheet = "DMA_786-O_vs_HK2")

write.csv(DMA_786O_vs_HK2_Intra, 
          paste0(Fig4D ,"/Fig.4d.csv"), 
          row.names = FALSE)

## Fig4E Alluvial plot
data(mca_core_rules)

MCA_core_Compressed <- mca_core_rules %>%
  mutate(Intra = case_when(
    Intra %in% c("UP", "DOWN") ~ Intra,  # Keep "UP" and "DOWN" as they are
    Intra %in% c("Not Detected", "Significant Negative", "Significant Positive", "Not Significant") ~ "No Change",
    TRUE ~ NA_character_  # Handle unexpected values
  )) %>%
  mutate(CoRe = case_when(
    CoRe %in% c("UP", "DOWN") ~ CoRe,  # Keep "UP" and "DOWN" as they are
    CoRe %in% c("Not Detected", "Significant Negative", "Significant Positive", "Not Significant") ~ "No Change",
    TRUE ~ NA_character_  # Handle unexpected values
  ))%>%
  distinct(Intra, CoRe, Core_Direction, .keep_all = TRUE)%>%  # Keep only unique rows based on these columns
  filter(!Core_Direction=="Not Detected")#remove the "Not Detected" as we are only interested in the released/consumed metabolites

write.csv(MCA_core_Compressed, 
          paste0(Fig4E ,"/Fig.4e.csv"), 
          row.names = FALSE)

## Fig4F MCA_Summary Table
MCA2Cond_Summary <- read_excel("Table_EV3.xlsx", sheet = "MCA2Cond_Summary")%>%
  filter(`Regulation.Grouping`=="RG2_Significant")%>%
  filter(!`SiRCle.cluster.Name`=="Background = FALSE")%>%
  filter(!`SiRCle.cluster.Name`=="None")

# this is not the exact same --> What happened? --> plot again!


## Fig4G
MCA2Cond_786M1A_vs_HK2 <- read_excel("Table_EV3.xlsx", sheet = "MCA2Cond_786M1A_vs_HK2")%>%
  filter(RG2_Significant =="Both_DOWN (Released/Consumed)")

Heatmap_Meatbs <- MCA2Cond_786M1A_vs_HK2$Metabolite

heatmap_plot_df <- ProcessedData %>%
  pivot_longer(
    cols = -all_of(sample_cols),
    names_to = "Metabolite",
    values_to = "Abundance"
  ) %>%
  left_join(
    DMAFeatureMetadata %>%
      select(all_of(c("Metabolite", "Pathway", "core_HK2", "core_786-O", "core_786-M1A"))),
    by = "Metabolite"
  )%>%
  filter(Outliers != "Outlier_filtering_round_1") %>%
  filter(!Metabolite %in% c("valine-d8", "hipppuric acid-d5"))  %>%
  #filter(Pathway == "Amino acid metabolism") %>% #Select for metabolites from cluster!
  select(-Biological_Replicates, -GrowthFactor, -Outliers)


## Fig4H
#Needs to be compiled from both processed intra and processed core data. 


```

## Figure 5
```{r}
# Create needed output folders
Fig5 <- file.path(Output_folder, "Figure 5")
if (!dir.exists(Fig5)) {dir.create(Fig5)} 

Fig5A <- file.path(Fig5 , "5A")
if (!dir.exists(Fig5A)) {dir.create(Fig5A)}

Fig5B <- file.path(Fig5 , "5B")
if (!dir.exists(Fig5B)) {dir.create(Fig5B)}  

Fig5C <- file.path(Fig5 , "5C")
if (!dir.exists(Fig5C)) {dir.create(Fig5C)}  

Fig5D <- file.path(Fig5 , "5D")
if (!dir.exists(Fig5D)) {dir.create(Fig5D)}

Fig5E <- file.path(Fig5 , "5E")
if (!dir.exists(Fig5E)) {dir.create(Fig5E)}

Fig5F <- file.path(Fig5 , "5F")
if (!dir.exists(Fig5F)) {dir.create(Fig5F)}

#-----------------------------------------------------------------------------
# Add Figure Source data and save as .csv files within the respective folders:

## Fig5A =

## Fig5B = ExtendedDataTable 5, sheet  1 "Raw data." --> needs data wrangling

## Fig5C = ExtendedDataTable 5, sheet ""

## Fig5D 

## Fig5E

## Fig5F
```


## EV Figure 1
```{r}
# Create needed output folders
FigEV1 <- file.path(Output_folder, "Figure EV1")
if (!dir.exists(FigEV1)) {dir.create(FigEV1)} 

FigEV1A <- file.path(FigEV1, "Figure EV1A")
if (!dir.exists(FigEV1A)) {dir.create(FigEV1A)} 

FigEV1B <- file.path(FigEV1, "Figure EV1B")
if (!dir.exists(FigEV1B)) {dir.create(FigEV1B)} 

FigEV1C <- file.path(FigEV1, "Figure EV1C")
if (!dir.exists(FigEV1C)) {dir.create(FigEV1C)} 


# Fig EV1a



# Fig EV1b
Table_EV2_ID_Summary <- read_excel("Table_EV2.xlsx", sheet = "ID_Summary")
Table_EV2_ID_Summary_Deltas <- read_excel("Table_EV2.xlsx", sheet = "ID_Summary_Deltas")
combined <- rbind(Table_EV2_ID_Summary, Table_EV2_ID_Summary_Deltas)

write.csv(combined , 
          paste0(FigEV1B ,"/Fig.EV1b.csv"), 
          row.names = FALSE)



# Fig EV1c










```


## EV Figure 2

```{r}
# Create needed output folders
FigEV2 <- file.path(Output_folder, "Figure EV2")
if (!dir.exists(FigEV2)) {dir.create(FigEV2)} 

FigEV2A <- file.path(FigEV2 , "2A")
if (!dir.exists(FigEV2A)) {dir.create(FigEV2A)}

FigEV2B <- file.path(FigEV2 , "2B")
if (!dir.exists(FigEV2B)) {dir.create(FigEV2B)}  

FigEV2C <- file.path(FigEV2 , "2C")
if (!dir.exists(FigEV2C)) {dir.create(FigEV2C)}  

FigEV2D <- file.path(FigEV2 , "2D")
if (!dir.exists(FigEV2D)) {dir.create(FigEV2D)}

FigEV2E <- file.path(FigEV2 , "2E")
if (!dir.exists(FigEV2E)) {dir.create(FigEV2E)}

#-----------------------------------------------------------------------------
# Add Figure Source data and save as .csv files within the respective folders:

## FigEV 2A = Table EV3, sheet 2
PoolEstimationCV <- read_excel("Table_EV3.xlsx", sheet = "PoolEstimationCV")

write.csv(PoolEstimationCV, 
          paste0(FigEV2A ,"/Fig.EV2a.csv"), 
          row.names = FALSE)

## FigEV 2B = Table EV3, sheet 3
MediaBlankCV <- read_excel("Table_EV3.xlsx", sheet = "MediaBlankCV")

write.csv(MediaBlankCV, 
          paste0(FigEV2B ,"/Fig.EV2b.csv"), 
          row.names = FALSE)

## FigEV 2C = PCA plot --> Sheet 5, processed data
ProcessedData <- read_excel("Table_EV3.xlsx", sheet = "ProcessedData")

write.csv(ProcessedData , 
          paste0(FigEV2C ,"/Fig.EV2c.csv"), 
          row.names = FALSE)




## Screeplot uses PCA results:
inflect_df <- as.data.frame(c(seq_along(PCA.res$sdev)))
        colnames(inflect_df) <- "x"
        inflect_df$y <- summary(PCA.res)$importance[2, ]
        inflect_df$Cumulative <- summary(PCA.res)$importance[3, ]
        screeplot_cumul <-
            format(
                round(inflect_df$Cumulative[seq_len(20)]*100, 1),
                nsmall = 1
            )  # make cumulative variation labels for plot
        knee <-
            uik(
                inflect_df$x,
                inflect_df$y
            )  # Calculate the knee and select optimal number of components
        # Note: we subtract 1 components from the knee cause the root of the
        # knee is the PC that does not add something. npcs = 30
        npcs <- knee -1
```










# Session information

```{r session_info, echo=FALSE}
options(width = 120)
sessionInfo()
```
