---
title: "SourceDataTable"
author:
  - name: Christina Schmidt
    affiliation:
    - Heidelberg University
output:
  html_document:
    self_contained: true
    toc: true
    toc_float: true
    toc_depth: 5
    code_folding: show
vignette: >
  %\VignetteIndexEntry{Standard Metabolomics}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmdarkdown}
editor_options:
  chunk_output_type: console
---

<style>
.vscroll-plot {
    width: 850px;
    height: 500px;
    overflow-y: scroll;
    overflow-x: hidden;
}
</style>

```{r chunk_setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

install the required dependencies and load the libraries:
```{r, message=FALSE, warning=FALSE}
#dependencies that need to be loaded:
library(dplyr)
library(magrittr)
library(tidyverse)
library(readxl)
library(MetaProViz)
```

```{r}
read_xlsx_all <- function(url) {
  tmp <- tempfile(fileext = ".xlsx")
  download.file(url, tmp, mode = "wb")
  
  sheets <- excel_sheets(tmp)
  lapply(sheets, function(s) read_excel(tmp, sheet = s)) |>
    setNames(sheets)
}
```

# Compile the source data tables for the figures
```{r}
#Create output folder
Output_folder <- file.path(getwd(), "SourceData")
if (!dir.exists(Output_folder)) {dir.create(Output_folder)} # Make Results folder
```

## Figure 1
Not applicable to Fig. 1a-b as Schematics.
```{r}
# Create needed output folders
Fig1 <- file.path(Output_folder, "Figure 1")
if (!dir.exists(Fig1)) {dir.create(Fig1)} 

Fig1C <- file.path(Fig1 , "1C")
if (!dir.exists(Fig1C)) {dir.create(Fig1C)}

Fig1D <- file.path(Fig1 , "1D")
if (!dir.exists(Fig1D)) {dir.create(Fig1D)}

#---------------------------------------------------------------------------------
# Add Figure Source data and save as .csv files within the respective folders:

# Fig1C
Table_EV1_Resource_Level <- read_excel("Table_EV1.xlsx", sheet = "Resource_Level")
write.csv(Table_EV1_Resource_Level , 
          paste0(Fig1C ,"/Fig.1c.csv"), 
          row.names = FALSE)

# Fig1D
Table_EV1_Term_Size_Distribution <- read_excel("Table_EV1.xlsx", sheet = "Term_Size_Distribution")
write.csv(Table_EV1_Term_Size_Distribution , 
          paste0(Fig1D ,"/Fig.1d.csv"), 
          row.names = FALSE)

```

## Figure 2
Not applicable to Fig 2A and D as Schematics.
```{r}
# Create needed output folders
Fig2 <- file.path(Output_folder, "Figure 2")
if (!dir.exists(Fig2)) {dir.create(Fig2)} 

Fig2B <- file.path(Fig2 , "2B")
if (!dir.exists(Fig2B)) {dir.create(Fig2B)}  

Fig2C <- file.path(Fig2 , "2C")
if (!dir.exists(Fig2C)) {dir.create(Fig2C)}  


#---------------------------------------------------------------------------------
# Add Figure Source data and save as .csv files within the respective folders:

# Fig 2B
Table_EV2_Raw_Feature_Metadata <- read_excel("Table_EV2.xlsx", sheet = "Raw_Feature_Metadata")
Table_EV2_Feature_Metadata_EquivalentID <- read_excel("Table_EV2.xlsx", sheet = "Feature_Metadata_EquivalentID")

# These two tables were used to generate both upsetplots
# append them using rbind, making same column order and adding a "Version" column, 
# indicating whether it is the "Original" (left) or "Adapted" (right) plot

Table_EV2_Raw_Feature_Metadata <- 
  Table_EV2_Raw_Feature_Metadata %>%
    mutate(
      CHEBI = NA_character_,
      manual_correction = NA_character_,
      correction_tbl = NA_character_,
      Plot = "Original"
    ) %>%
    select(
        Plot,
        Metabolite,
        CAS,
        SUPER_PATHWAY,
        SUB_PATHWAY,
        COMP_ID,
        PLATFORM,
        RI,
        MASS,
        HMDB,
        KEGG,
        PUBCHEM,
        CHEBI,
        manual_correction,
        correction_tbl
    )

Table_EV2_Feature_Metadata_EquivalentID <- 
  Table_EV2_Feature_Metadata_EquivalentID %>%
    mutate(
      Plot = "Adapted"
    ) %>%
    select(
        Plot,
        Metabolite,
        CAS,
        SUPER_PATHWAY,
        SUB_PATHWAY,
        COMP_ID,
        PLATFORM,
        RI,
        MASS,
        HMDB,
        KEGG,
        PUBCHEM,
        CHEBI,
        manual_correction,
        correction_tbl
    )

combined <- rbind(Table_EV2_Raw_Feature_Metadata, Table_EV2_Feature_Metadata_EquivalentID)

write.csv(combined, 
          paste0(Fig2B ,"/Fig.2b.csv"), 
          row.names = FALSE)


# Fig 2C

Table_EV2_ID_Summary <- read_excel("Table_EV2.xlsx", sheet = "ID_Summary")
Table_EV2_ID_Summary_Deltas <- read_excel("Table_EV2.xlsx", sheet = "ID_Summary_Deltas")
combined <- rbind(Table_EV2_ID_Summary, Table_EV2_ID_Summary_Deltas)
combined <- combined %>% select(row_name, Total_IDs)

write.csv(combined, 
          paste0(Fig2C ,"/Fig.2c.csv"), 
          row.names = FALSE)

```

## Figure 3
Not applicable to Fig 3A as Schematics
```{r}
# Create needed output folders
Fig3 <- file.path(Output_folder, "Figure 3")
if (!dir.exists(Fig3)) {dir.create(Fig3)} 

Fig3B <- file.path(Fig3 , "3B")
if (!dir.exists(Fig3B)) {dir.create(Fig3B)}  

Fig3C <- file.path(Fig3 , "3C")
if (!dir.exists(Fig3C)) {dir.create(Fig3C)}  

Fig3D <- file.path(Fig3 , "3D")
if (!dir.exists(Fig3D)) {dir.create(Fig3D)}

Fig3E <- file.path(Fig3 , "3E")
if (!dir.exists(Fig3E)) {dir.create(Fig3E)}

Fig3F <- file.path(Fig3 , "3F")
if (!dir.exists(Fig3F)) {dir.create(Fig3F)}

Fig3G <- file.path(Fig3 , "3G")
if (!dir.exists(Fig3G)) {dir.create(Fig3G)}

#---------------------------------------------------------------------------------
# Add Figure Source data and save as .csv files within the respective folders:

## Fig3B = Table EV3, sheet  1 "Raw data." --> needs data wrangling
data(intracell_raw)

CellPlot_Intra <- intracell_raw%>%
  filter(Conditions != "Pool")%>%
   mutate(Celltype = case_when(Conditions=="HK2" ~ 'Healthy',
                               Conditions=="786-O" ~ 'PrimaryTumour',
                               TRUE ~ 'MetastaticTumour'))

CellPlot_Intra$Count <- 1
CellPlot_Intra$CelltypeFactor <- factor(CellPlot_Intra$Celltype, levels = c("Healthy","PrimaryTumour", "MetastaticTumour"))

CellPlot_Intra_select <- CellPlot_Intra %>%
  select(Code, Conditions, Analytical_Replicates, Biological_Replicates, Celltype, Count)

write.csv(CellPlot_Intra_select, 
          paste0(Fig3B ,"/Fig.3b.csv"), 
          row.names = FALSE)

## Fig3C = Table EV3, sheet 2
PoolEstimationCV <- read_excel("Table_EV3.xlsx", sheet = "PoolEstimationCV")

write.csv(PoolEstimationCV, 
          paste0(Fig3C ,"/Fig.3c.csv"), 
          row.names = FALSE)

## Fig3D = Table EV3, sheet 3
MediaBlankCV <- read_excel("Table_EV3.xlsx", sheet = "MediaBlankCV")

write.csv(MediaBlankCV, 
          paste0(Fig3D ,"/Fig.3d.csv"), 
          row.names = FALSE)

## Fig3E = PCA Pool QC
# Load from BoRxiv branch:
Pool_Estimation_result <- read.csv(
    "https://raw.githubusercontent.com/saezlab/MetaProViz/BioRxiv/Cells_CoRe/MetaProViz_Results/Processing/pool_estimation/pool_estimation_data_2025-08-01.csv",
    sep=","
  )%>%
  column_to_rownames("Code")

#Run prcomp to generate PCA results and check plot
PCA_RES <- prcomp(as.matrix(Pool_Estimation_result), scale. = as.logical(TRUE))
autoplot(PCA_RES)

#extract PCA loading:
pca_var <- (PCA_RES$sdev^2) / sum(PCA_RES$sdev^2)

pca_pc1_pc2_df <- as.data.frame(PCA_RES$x) %>%
  rownames_to_column("Code") %>%
  transmute(
    Code,
    PC1,
    PC2,
    PC1_variance_pct = round(pca_var[1] * 100, 2),
    PC2_variance_pct = round(pca_var[2] * 100, 2)
  )

ggplot(pca_pc1_pc2_df, aes(x = PC1, y = PC2)) +
  geom_point(size = 3, alpha = 0.8, color = "#2C7FB8") +
  labs(x = paste0("PC1 (", sprintf("%.2f", unique(pca_pc1_pc2_df$PC1_variance_pct)[1]), "%)"),
       y = paste0("PC2 (", sprintf("%.2f", unique(pca_pc1_pc2_df$PC2_variance_pct)[1]), "%)"))

#sAVE:
write.csv(pca_pc1_pc2_df , 
          paste0(Fig3E ,"/Fig.3e.csv"), 
          row.names = FALSE)

## Fig3F = Before/After TIC
RawData <- read_excel("Table_EV3.xlsx", sheet = "RawData")%>%
  mutate(Panel = "LEFT", .before = 1)

ProcessedData <- read_excel("Table_EV3.xlsx", sheet = "ProcessedData")%>%
  mutate(Panel = "RIGHT", .before = 1)

CombinedData <- bind_rows(RawData, ProcessedData)

write.csv(CombinedData,
          paste0(Fig3F ,"/Fig.3f.csv"), 
          row.names = FALSE)

## Fig3G = PCA outliers
# The figures have been generated within MetaProViz functions and the data tables have not been saved. 
#Hence, we will extract the data now:

###-----------------------------
# Prepare the input:
data(medium_raw)

Media <- medium_raw%>%
column_to_rownames("Code")

Media_input <- Media%>%
subset(!Conditions=="Pool", select = -c(1:3))#remove pool samples and remove the information columns

Media_Metadata <- Media%>%
subset(!Conditions=="Pool", select = c(1:3))#remove pool samples and keep the information columns only

# data files:
input_data <- Media_input
data <- Media_input
data %<>% as.data.frame()
data %<>% mutate_all(~ ifelse(grepl("^0*(\\.0*)?$", as.character(.)), NA, .))
data <- as.data.frame(mutate_all(as.data.frame(data), function(x) as.numeric(as.character(x))))

metadata_sample <- Media_Metadata

# parameter info
metadata_info = c(Conditions = "Conditions",
                  Biological_Replicates = "Biological_Replicates",
                  core_norm_factor = "GrowthFactor",
                  core_media = "blank")
                                                
featurefilt = "Modified"
                            
cutoff_featurefilt = 0.8
                                                
mvi_percentage=50
                                                
hotellins_confidence = 0.99# We perform outlier testing using 0.99 confidence interval
                                                
core = TRUE

tic = TRUE

#1 Feature filtering:
data_Filtered <- MetaProViz:::feature_filtering(
            data = data,
            featurefilt = featurefilt,
            cutoff_featurefilt = cutoff_featurefilt,
            metadata_sample = metadata_sample,
            metadata_info = metadata_info,
            core = core)

data_Filt <- data_Filtered[["DF"]]

#MVI
mviRes <- MetaProViz:::mvi_imputation(
            data = data_Filt,
            metadata_sample = metadata_sample,
            metadata_info = metadata_info,
            core = core,
            mvi_percentage = mvi_percentage)

# Perform tic
ticRes_List <- MetaProViz:::tic_norm(
            data = mviRes,
            metadata_sample = metadata_sample,
            metadata_info = metadata_info,
            tic = tic)
ticRes <- ticRes_List[["DF"]][["data_tic"]]

#Core
data_coreNorm <- MetaProViz:::core_norm(
            data = ticRes,
            metadata_sample = metadata_sample,
            metadata_info = metadata_info)

ticRes <- data_coreNorm[["DF"]][["core_Norm"]]
data_norm <- ticRes %>% as.data.frame()

#Outlier Analysis
OutlierRes <- MetaProViz:::outlier_detection(
        data = data_norm,
        metadata_sample = metadata_sample,
        metadata_info = metadata_info,
        core = core,
        hotellins_confidence = hotellins_confidence)
##----------------------------------------------------------------------------------

#Get the PCA scores and variance explained for the QC PCA plot:
qc_scores <- OutlierRes$DF$QC_PCA_scores_Final
qc_variance <- OutlierRes$DF$QC_PCA_variance_Final

qc_scores_meta <- qc_scores %>%
  left_join(
    metadata_sample %>%
      rownames_to_column("Sample") %>%
      select(Sample, Conditions, Biological_Replicates),
    by = "Sample"
  ) %>%
  left_join(
    OutlierRes$DF$data_outliers %>%
      rownames_to_column("Sample") %>%
      select(Sample, Outliers),
    by = "Sample"
  )

ggplot(qc_scores_meta, aes(x = PC1, y = PC2, color = Conditions, shape = Outliers)) +
  geom_point(size = 3) +
  labs(
    title = "QC PCA (Final - Colored by Condition)",
    x = paste0("PC1 (", round(qc_variance$VarianceExplained[1]*100, 2), "%)"),
    y = paste0("PC2 (", round(qc_variance$VarianceExplained[2]*100, 2), "%)")
  ) +
  theme_classic()

#Save
write.csv(qc_scores_meta , 
          paste0(Fig3G ,"/Fig.3g.csv"), 
          row.names = FALSE)
```

## Figure 4
```{r}
# Create needed output folders
Fig4 <- file.path(Output_folder, "Figure 4")
if (!dir.exists(Fig4)) {dir.create(Fig4)} 

Fig4A <- file.path(Fig4 , "4A")
if (!dir.exists(Fig4A)) {dir.create(Fig4A)}

Fig4B <- file.path(Fig4 , "4B")
if (!dir.exists(Fig4B)) {dir.create(Fig4B)}  

Fig4C <- file.path(Fig4 , "4C")
if (!dir.exists(Fig4C)) {dir.create(Fig4C)}  

Fig4D <- file.path(Fig4 , "4D")
if (!dir.exists(Fig4D)) {dir.create(Fig4D)}

Fig4E <- file.path(Fig4 , "4E")
if (!dir.exists(Fig4E)) {dir.create(Fig4E)}

Fig4F <- file.path(Fig4 , "4F")
if (!dir.exists(Fig4F)) {dir.create(Fig4F)}

Fig4G <- file.path(Fig4 , "4G")
if (!dir.exists(Fig4G)) {dir.create(Fig4G)}

Fig4H <- file.path(Fig4 , "4H")
if (!dir.exists(Fig4H)) {dir.create(Fig4H)}

#---------------------------------------------------------------------------------
# Add Figure Source data and save as .csv files within the respective folders:

## Fig4A = Table_EV3
DMA_786O_vs_HK2<- read_excel("Table_EV3.xlsx", sheet = "DMA_786-O_vs_HK2")

write.csv(DMA_786O_vs_HK2, 
          paste0(Fig4A ,"/Fig.4a.csv"), 
          row.names = FALSE)

## Fig4B = Table_EV3
DMAFeatureMetadata<- read_excel("Table_EV3.xlsx", sheet = "DMA_FeatureMetadata")%>%
  select(-HMDB, -KEGG.ID, -KEGGCompound)%>%
  drop_na()

write.csv(DMAFeatureMetadata, 
          paste0(Fig4B ,"/Fig.4b.csv"), 
          row.names = FALSE)

## Fig4C = Table_EV4, sheet ""
ProcessedData <- read_excel("Table_EV3.xlsx", sheet = "ProcessedData")

# 1) Columns in ProcessedData that are sample metadata (not metabolites)
sample_cols <- c("Code", "Conditions", "Biological_Replicates", "GrowthFactor", "Outliers")

# 2) Convert metabolite columns to long format (one row = one heatmap cell)
heatmap_plot_df <- ProcessedData %>%
  pivot_longer(
    cols = -all_of(sample_cols),
    names_to = "Metabolite",
    values_to = "Abundance"
  ) %>%
  left_join(
    DMAFeatureMetadata %>%
      select(all_of(c("Metabolite", "Pathway", "core_HK2", "core_786-O", "core_786-M1A"))),
    by = "Metabolite"
  )%>%
  filter(Outliers != "Outlier_filtering_round_1") %>%
  filter(!Metabolite %in% c("valine-d8", "hipppuric acid-d5"))  %>%
  filter(Pathway == "Amino acid metabolism") %>%
  select(-Biological_Replicates, -GrowthFactor, -Outliers)

write.csv(heatmap_plot_df, 
          paste0(Fig4C ,"/Fig.4c.csv"), 
          row.names = FALSE)

## Fig4D
DMA_786O_vs_HK2_Intra <- read_excel("Table_EV4.xlsx", sheet = "DMA_786-O_vs_HK2")

write.csv(DMA_786O_vs_HK2_Intra, 
          paste0(Fig4D ,"/Fig.4d.csv"), 
          row.names = FALSE)

## Fig4E Alluvial plot
data(mca_core_rules)

MCA_core_Compressed <- mca_core_rules %>%
  mutate(Intra = case_when(
    Intra %in% c("UP", "DOWN") ~ Intra,  # Keep "UP" and "DOWN" as they are
    Intra %in% c("Not Detected", "Significant Negative", "Significant Positive", "Not Significant") ~ "No Change",
    TRUE ~ NA_character_  # Handle unexpected values
  )) %>%
  mutate(CoRe = case_when(
    CoRe %in% c("UP", "DOWN") ~ CoRe,  # Keep "UP" and "DOWN" as they are
    CoRe %in% c("Not Detected", "Significant Negative", "Significant Positive", "Not Significant") ~ "No Change",
    TRUE ~ NA_character_  # Handle unexpected values
  ))%>%
  distinct(Intra, CoRe, Core_Direction, .keep_all = TRUE)%>%  # Keep only unique rows based on these columns
  filter(!Core_Direction=="Not Detected")#remove the "Not Detected" as we are only interested in the released/consumed metabolites

write.csv(MCA_core_Compressed, 
          paste0(Fig4E ,"/Fig.4e.csv"), 
          row.names = FALSE)

## Fig4F MCA_Summary Table
MCA2Cond_Summary <- read_excel("Table_EV3.xlsx", sheet = "MCA2Cond_Summary")%>%
  filter(`Regulation.Grouping`=="RG2_Significant")%>%
  filter(!`SiRCle.cluster.Name`=="Background = FALSE")%>%
  filter(!`SiRCle.cluster.Name`=="None")

write.csv(MCA2Cond_Summary, 
          paste0(Fig4F ,"/Fig.4f.csv"), 
          row.names = FALSE)

## Fig4G
MCA2Cond_786M1A_vs_HK2 <- read_excel("Table_EV3.xlsx", sheet = "MCA2Cond_786M1A_vs_HK2")%>%
  filter(RG2_Significant =="Both_DOWN (Released/Consumed)")

Heatmap_Meatbs <- MCA2Cond_786M1A_vs_HK2$Metabolite

heatmap_plot_df <- ProcessedData %>%
  pivot_longer(
    cols = -all_of(sample_cols),
    names_to = "Metabolite",
    values_to = "Abundance"
  ) %>%
  left_join(
    DMAFeatureMetadata %>%
      select(all_of(c("Metabolite", "Pathway", "core_HK2", "core_786-O", "core_786-M1A"))),
    by = "Metabolite"
  )%>%
  filter(Outliers != "Outlier_filtering_round_1") %>%
  filter(!Metabolite %in% c("valine-d8", "hipppuric acid-d5"))  %>%
  #filter(Pathway == "Amino acid metabolism") %>% #Select for metabolites from cluster!
  select(-Biological_Replicates, -GrowthFactor, -Outliers)

write.csv(heatmap_plot_df, 
          paste0(Fig4G ,"/Fig.4g.csv"), 
          row.names = FALSE)

## Fig4H
#Needs to be compiled from both processed intra and processed core data. 
ProcessedData_Core <- read_excel("Table_EV3.xlsx", sheet = "ProcessedData")%>%
  select(Code, Conditions, Biological_Replicates, methionine)%>%
  mutate(Panel = "CoRe", .before = 1)


ProcessedData_Intra <- read_excel("Table_EV4.xlsx", sheet = "ProcessedData")%>%
  select(Code, Conditions, Biological_Replicates, methionine)%>%
  mutate(Panel = "Intra", .before = 1)

Processed_CombinedData <- bind_rows(ProcessedData_Core, ProcessedData_Intra)

write.csv(Processed_CombinedData,
          paste0(Fig4H ,"/Fig.4h.csv"), 
          row.names = FALSE)

```

## Figure 5
```{r}
# Create needed output folders
Fig5 <- file.path(Output_folder, "Figure 5")
if (!dir.exists(Fig5)) {dir.create(Fig5)} 

Fig5A <- file.path(Fig5 , "5A")
if (!dir.exists(Fig5A)) {dir.create(Fig5A)}

Fig5B <- file.path(Fig5 , "5B")
if (!dir.exists(Fig5B)) {dir.create(Fig5B)}  

Fig5C <- file.path(Fig5 , "5C")
if (!dir.exists(Fig5C)) {dir.create(Fig5C)}  

Fig5D <- file.path(Fig5 , "5D")
if (!dir.exists(Fig5D)) {dir.create(Fig5D)}

Fig5E <- file.path(Fig5 , "5E")
if (!dir.exists(Fig5E)) {dir.create(Fig5E)}

Fig5F <- file.path(Fig5 , "5F")
if (!dir.exists(Fig5F)) {dir.create(Fig5F)}

#-----------------------------------------------------------------------------
# Add Figure Source data and save as .csv files within the respective folders:

## Fig5A
data(tissue_norm)

TissuePlot <- tissue_norm%>%
  column_to_rownames("Code")%>%
  filter(TISSUE_TYPE != "NORMAL")%>%
  separate(`TYPE-STAGE`, into=c("Type", "Stage"), sep="-", remove=FALSE)
TissuePlot$Count <- 1

TissuePlot$AGEFactor <- factor(TissuePlot$AGE, levels = c("Young","Middle", "Old"))

ggplot2::ggplot(TissuePlot, ggplot2::aes(fill=Stage, y=Count, x=AGEFactor)) + 
        ggplot2::geom_bar(position="stack", stat="identity")+
        ggplot2::scale_fill_manual(values = c("STAGE I" = "#006384"  , "STAGE II" = "#9F0162", "STAGE III" = "#00735C" , "STAGE IV" = "grey" ))+
        ggplot2::theme_classic()+
        ggplot2::ggtitle("ccRCC patients metadata")+
        ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, hjust = 1))+
        ggplot2::labs(subtitle="Age at surgery <42 = Young,  >58 = Old, else = Middle", x = "Age", y = "Number of Patients")

write.csv(TissuePlot , 
          paste0(Fig5A ,"/Fig.5a.csv"), 
          row.names = FALSE)

## Fig5B
Stat_results <- read_excel("Table_EV2.xlsx", sheet = "MetaAnalysis_AOV")

data_Heat <-
        Stat_results %>%
        filter(tukeyHSD_p.adjusted < 0.05) %>%
        # Filter for significant results
        # Exclude Residuals row
        filter(Explained_Variance > 1) %>%
        distinct(
            term,
            PC,
            .keep_all = TRUE
        ) %>%  # only keep unique term~PC combinations AND STATS
        select(term, PC, Explained_Variance) %>%
        pivot_wider(names_from = PC, values_from = Explained_Variance) %>%
        column_to_rownames("term") %>%
        mutate_all(~ replace(., is.na(.), 0L))

write.csv(data_Heat,
          paste0(Fig5B ,"/Fig.5b.csv"),
          row.names = TRUE)

## Fig5C alluvial plot
data("mca_twocond_rules")

MCA_twocond_Compressed <- mca_twocond_rules %>%
  mutate(Cond1 = case_when(
    Cond1 %in% c("UP", "DOWN") ~ Cond1,  # Keep "UP" and "DOWN" as they are
    Cond1 %in% c("Not Detected", "Significant Negative", "Significant Positive", "Not Significant") ~ "No Change",
    TRUE ~ NA_character_  # Handle unexpected values
  )) %>%
  mutate(Cond2 = case_when(
    Cond2 %in% c("UP", "DOWN") ~ Cond2,  # Keep "UP" and "DOWN" as they are
    Cond2 %in% c("Not Detected", "Significant Negative", "Significant Positive", "Not Significant") ~ "No Change",
    TRUE ~ NA_character_  # Handle unexpected values
  ))%>%
  distinct(Cond1, Cond2, .keep_all = TRUE)

write.csv(MCA_twocond_Compressed, 
          paste0(Fig5C ,"/Fig.5c.csv"), 
          row.names = FALSE)

## Fig5D 
#Select Dipeptides
Dipeptides <- read_excel("Table_EV2.xlsx", sheet = "Raw_Feature_Metadata")%>%
  filter(SUB_PATHWAY == "Dipeptide")%>%
  select(Metabolite)

#Tumour versus Normal Young --> Selection Dipeptides
DMA_Young <- read_excel("Table_EV2.xlsx", sheet = "DMA_TUMOR_vs_NORMAL_Young")%>%
  select(Metabolite, Log2FC, p.adj)%>%
  filter(Metabolite %in% Dipeptides$Metabolite)%>%
  mutate(Panel = "Young", .before = 1)

#Tumour versus Normal Old --> Selection Dipeptides
DMA_Old <- read_excel("Table_EV2.xlsx", sheet = "DMA_TUMOR_vs_NORMAL_Old")%>%
  select(Metabolite, Log2FC, p.adj)%>%
  filter(Metabolite %in% Dipeptides$Metabolite)%>%
  mutate(Panel = "Old", .before = 1)

# Addition biological clusters
MCA_2Cond <- read_excel("Table_EV2.xlsx", sheet = "MCA2Cond_Old-Young") %>%
  select(Metabolite, RG2_Significant)

DMA_Young <- merge(DMA_Young, MCA_2Cond, by = "Metabolite", all.x = TRUE)
DMA_Old <- merge(DMA_Old, MCA_2Cond, by = "Metabolite", all.x = TRUE)

#Cobine and save
CombinedData_YO <- bind_rows(DMA_Young, DMA_Old)

write.csv(CombinedData_YO,
          paste0(Fig5D ,"/Fig.5d.csv"), 
          row.names = FALSE)


## Fig5E
# Select pathway metabolites

volcano_plot_source_data_Young <- 
  read.csv2("Patients_Metadata/MetaProViz_Results/Enrichment/volcano_plot_source_data_Young.csv") %>%
  select(-X) %>%
  filter(
      term == "Mineral absorption" | term == "Protein digestion and absorption"
    ) 

write.csv(volcano_plot_source_data_Young, 
          paste0(Fig5E ,"/Fig.5e.csv"), 
          row.names = FALSE)

## Fig5F
# read in cluster_summary and similarity_matrix from Table EV2

clustered_data <- read_excel("Table_EV2.xlsx", sheet = "ORA_Clustered_Data")
cluster_similarity_matrix <- read_excel("Table_EV2.xlsx", sheet = "ORA_Cluster_Similarity_Matrix")

write.csv(clustered_data, 
          paste0(Fig5F ,"/Fig.5f_1.csv"), 
          row.names = FALSE)
write.csv(cluster_similarity_matrix, 
          paste0(Fig5F ,"/Fig.5f_2.csv"), 
          row.names = FALSE)

```


## EV Figure 1
```{r}
# Create needed output folders
FigEV1 <- file.path(Output_folder, "Figure EV1")
if (!dir.exists(FigEV1)) {dir.create(FigEV1)} 

FigEV1A <- file.path(FigEV1, "Figure EV1A")
if (!dir.exists(FigEV1A)) {dir.create(FigEV1A)} 

FigEV1B <- file.path(FigEV1, "Figure EV1B")
if (!dir.exists(FigEV1B)) {dir.create(FigEV1B)} 

FigEV1C <- file.path(FigEV1, "Figure EV1C")
if (!dir.exists(FigEV1C)) {dir.create(FigEV1C)} 


# Fig EV1a
Table_EV2_AlanineMetaboliteIDs <- read_excel("Table_EV2.xlsx", sheet = "AlanineMetaboliteIDs")

write.csv(Table_EV2_AlanineMetaboliteIDs, 
          paste0(FigEV1A ,"/Fig.EV1a.csv"), 
          row.names = FALSE)

# Fig EV1b
Table_EV2_ID_Summary <- read_excel("Table_EV2.xlsx", sheet = "ID_Summary")
Table_EV2_ID_Summary_Deltas <- read_excel("Table_EV2.xlsx", sheet = "ID_Summary_Deltas")
combined <- rbind(Table_EV2_ID_Summary, Table_EV2_ID_Summary_Deltas)

write.csv(combined , 
          paste0(FigEV1B ,"/Fig.EV1b.csv"), 
          row.names = FALSE)


# Fig EV1c
Table_EV2_Alanine_MappingAmbiguities <- read_excel("Table_EV2.xlsx", sheet = "Alanine_MappingAmbiguities")

write.csv(Table_EV2_Alanine_MappingAmbiguities, 
          paste0(FigEV1C ,"/Fig.EV1c.csv"), 
          row.names = FALSE)
```


## EV Figure 2

```{r}
# Create needed output folders
FigEV2 <- file.path(Output_folder, "Figure EV2")
if (!dir.exists(FigEV2)) {dir.create(FigEV2)} 

FigEV2A <- file.path(FigEV2 , "2A")
if (!dir.exists(FigEV2A)) {dir.create(FigEV2A)}

FigEV2B <- file.path(FigEV2 , "2B")
if (!dir.exists(FigEV2B)) {dir.create(FigEV2B)}  

FigEV2C <- file.path(FigEV2 , "2C")
if (!dir.exists(FigEV2C)) {dir.create(FigEV2C)}  

FigEV2D <- file.path(FigEV2 , "2D")
if (!dir.exists(FigEV2D)) {dir.create(FigEV2D)}

FigEV2E <- file.path(FigEV2 , "2E")
if (!dir.exists(FigEV2E)) {dir.create(FigEV2E)}

#-----------------------------------------------------------------------------
# Add Figure Source data and save as .csv files within the respective folders:

## FigEV 2A = Table EV3, sheet 2
PoolEstimationCV <- read_excel("Table_EV3.xlsx", sheet = "PoolEstimationCV")

write.csv(PoolEstimationCV, 
          paste0(FigEV2A ,"/Fig.EV2a.csv"), 
          row.names = FALSE)

## FigEV 2B = Table EV3, sheet 3
MediaBlankCV <- read_excel("Table_EV3.xlsx", sheet = "MediaBlankCV")

write.csv(MediaBlankCV, 
          paste0(FigEV2B ,"/Fig.EV2b.csv"), 
          row.names = FALSE)

## FigEV 2C = PCA plot
# We use the data we created for Fig 3g (see above)
qc_scores_meta$Biological_Replicates <- as.factor(qc_scores_meta$Biological_Replicates)

#Save
write.csv(qc_scores_meta , 
          paste0(FigEV2B ,"/Fig.EV2c.csv"), 
          row.names = FALSE)

## FigEV 2D: Filtering round 1 Hotellins plot
round_num <- 1
hotellings_df <- OutlierRes$DF[[paste0("Hotellings_results_Round", round_num)]]# generated for Fig. 3G (see above)

ggplot(hotellings_df, aes(x = Samples, y = `Group summary statistics`)) +
  geom_point(color = "blue", size = 2)

write.csv(hotellings_df , 
          paste0(FigEV2B ,"/Fig.EV2d.csv"), 
          row.names = FALSE)


## FigEV 2E Screeplot uses PCA results from Filtering round 1:
variance_round1 <- OutlierRes$DF$PCA_variance_Round1 %>%# generated for Fig. 3G (see above)
  mutate(
    PC_num = seq_len(n()),
    VariancePercent = VarianceExplained * 100
  ) %>%
  slice(1:min(20, n()))

ggplot(variance_round1, aes(x = PC_num, y = VariancePercent)) +
  geom_col(width = 1, fill = "grey70", color = "grey40") +
  geom_line(aes(group = 1), color = "black", linewidth = 0.5) +
  geom_point(color = "black", size = 1.5) +
  geom_text(
    aes(label = sprintf("%.2f", round(VariancePercent, 1))),
    vjust = -0.3,
    size = 3
  ) +
  scale_x_continuous(
    breaks = variance_round1$PC_num,
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.08))
  ) +
  labs(
    title = "PCA Explained Variance Plot Filtering Round 1",
    x = "Principal Component",
    y = "Explained variance (%)"
  ) +
  theme_classic()

write.csv(variance_round1 , 
          paste0(FigEV2E ,"/Fig.EV2e.csv"), 
          row.names = FALSE)

```










# Session information

```{r session_info, echo=FALSE}
options(width = 120)
sessionInfo()
```
