---
title: "SourceDataTable"
author:
  - name: Christina Schmidt
    affiliation:
    - Heidelberg University
output:
  html_document:
    self_contained: true
    toc: true
    toc_float: true
    toc_depth: 5
    code_folding: show
vignette: >
  %\VignetteIndexEntry{Standard Metabolomics}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
bibliography: bibliography.bib
editor_options:
  chunk_output_type: console
---

<style>
.vscroll-plot {
    width: 850px;
    height: 500px;
    overflow-y: scroll;
    overflow-x: hidden;
}
</style>

```{r chunk_setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

install the required dependencies and load the libraries:
```{r, message=FALSE, warning=FALSE}
#dependencies that need to be loaded:
library(dplyr)
library(readxl)

```

# Compile the source data tables for the figures

```{r}
#Create output folder
Output_folder <- file.path(getwd(), "SourceData")
if (!dir.exists(Output_folder)) {dir.create(Output_folder)} # Make Results folder
```

## Figure 1
Not applicable as Schematics.

## Figure 2
Not applicable to Fig 2D-E as Schmatics
```{r}
# Create needed output folders
Fig2 <- file.path(Output_folder, "Figure 2")
if (!dir.exists(Fig2)) {dir.create(Fig2)} 

Fig2A <- file.path(Fig2 , "2A")
if (!dir.exists(Fig2A)) {dir.create(Fig2A)}  

Fig2B <- file.path(Fig2 , "2B")
if (!dir.exists(Fig2B)) {dir.create(Fig2B)}  

Fig2C <- file.path(Fig2 , "2C")
if (!dir.exists(Fig2C)) {dir.create(Fig2C)}  


#---------------------------------------------------------------------------------
# Add Figure Source data and save as .csv files within the respective folders:

## Fig2A = ExtendedDataTable 1, sheet 2 "Feature metadata"

## Fig2B = ExtendedDataTable 2, sheet 2 "Alanine IDs"

## Fig2C = ExtendedDataTable 2, sheet 3 "Mapping ambiguities"

```

## Figure 3
Not applicable to Fig 3A as Schmatics
```{r}
# Create needed output folders
Fig3 <- file.path(Output_folder, "Figure 3")
if (!dir.exists(Fig3)) {dir.create(Fig3)} 

Fig3B <- file.path(Fig3 , "3B")
if (!dir.exists(Fig3B)) {dir.create(Fig3B)}  

Fig3C <- file.path(Fig3 , "3C")
if (!dir.exists(Fig3C)) {dir.create(Fig3C)}  

Fig3D <- file.path(Fig3 , "3D")
if (!dir.exists(Fig3D)) {dir.create(Fig3D)}

Fig3E <- file.path(Fig3 , "3E")
if (!dir.exists(Fig3E)) {dir.create(Fig3E)}

Fig3F <- file.path(Fig3 , "3F")
if (!dir.exists(Fig3F)) {dir.create(Fig3F)}

Fig3G <- file.path(Fig3 , "3G")
if (!dir.exists(Fig3G)) {dir.create(Fig3G)}

#---------------------------------------------------------------------------------
# Add Figure Source data and save as .csv files within the respective folders:

## Fig3B = ExtendedDataTable 3, sheet  1 "Raw data." --> needs data wrangling
CellPlot_Intra <- intracell_raw%>%
  column_to_rownames("Code")%>%
  filter(Conditions != "Pool")%>%
   mutate(Celltype = case_when(Conditions=="HK2" ~ 'Healthy',
                               Conditions=="786-O" ~ 'PrimaryTumour',
                               TRUE ~ 'MetastaticTumour'))

CellPlot_Intra$Count <- 1
CellPlot_Intra$CelltypeFactor <- factor(CellPlot_Intra$Celltype, levels = c("Healthy","PrimaryTumour", "MetastaticTumour"))

#--> Add those info to extended data table 3 as sheet 1 about sample infos! Do the same for the other sample data (intra), patients have this already


## Fig3C = ExtendedDataTable 3, sheet ""

## Fig3D

## Fig3E

## Fig3F
```

## Figure 4
```{r}
# Create needed output folders
Fig4 <- file.path(Output_folder, "Figure 4")
if (!dir.exists(Fig4)) {dir.create(Fig4)} 

Fig4A <- file.path(Fig4 , "4A")
if (!dir.exists(Fig4A)) {dir.create(Fig4A)}

Fig4B <- file.path(Fig4 , "4B")
if (!dir.exists(Fig4B)) {dir.create(Fig4B)}  

Fig4C <- file.path(Fig4 , "4C")
if (!dir.exists(Fig4C)) {dir.create(Fig4C)}  

Fig4D <- file.path(Fig4 , "4D")
if (!dir.exists(Fig4D)) {dir.create(Fig4D)}

Fig4E <- file.path(Fig4 , "4E")
if (!dir.exists(Fig4E)) {dir.create(Fig4E)}

Fig4F <- file.path(Fig4 , "4F")
if (!dir.exists(Fig4F)) {dir.create(Fig4F)}

Fig4G <- file.path(Fig4 , "4G")
if (!dir.exists(Fig4G)) {dir.create(Fig4G)}

Fig4H <- file.path(Fig4 , "4H")
if (!dir.exists(Fig4H)) {dir.create(Fig4H)}

#---------------------------------------------------------------------------------
# Add Figure Source data and save as .csv files within the respective folders:

## Fig4A =

## Fig4B = ExtendedDataTable 4, sheet  1 "Raw data." --> needs data wrangling

## Fig4C = ExtendedDataTable 4, sheet ""

## Fig4D

## Fig4E

## Fig4F
```

## Figure 5
```{r}
# Create needed output folders
Fig5 <- file.path(Output_folder, "Figure 5")
if (!dir.exists(Fig5)) {dir.create(Fig5)} 

Fig5A <- file.path(Fig5 , "5A")
if (!dir.exists(Fig5A)) {dir.create(Fig5A)}

Fig5B <- file.path(Fig5 , "5B")
if (!dir.exists(Fig5B)) {dir.create(Fig5B)}  

Fig5C <- file.path(Fig5 , "5C")
if (!dir.exists(Fig5C)) {dir.create(Fig5C)}  

Fig5D <- file.path(Fig5 , "5D")
if (!dir.exists(Fig5D)) {dir.create(Fig5D)}

Fig5E <- file.path(Fig5 , "5E")
if (!dir.exists(Fig5E)) {dir.create(Fig5E)}

Fig5F <- file.path(Fig5 , "5F")
if (!dir.exists(Fig5F)) {dir.create(Fig5F)}

#-----------------------------------------------------------------------------
# Add Figure Source data and save as .csv files within the respective folders:

## Fig5A =

## Fig5B = ExtendedDataTable 5, sheet  1 "Raw data." --> needs data wrangling

## Fig5C = ExtendedDataTable 5, sheet ""

## Fig5D 

## Fig5E

## Fig5F
```



## EV Figure 1
Not applicable as Schematics.

## EV Figure 2

## EV Figure 3










# Session information

```{r session_info, echo=FALSE}
options(width = 120)
sessionInfo()
```
