---
title: "Supplementary_Tables"
author: 
  - name: Jannik Franken
      affiliation:
      - Heidelberg University
date: "14 4 2026"
output: html_document
---

# Supplementary table generation

This script iterates over all output generated and builds the supplementary tables

```{r}
library(MetaProViz)
library(magrittr)
library(dplyr)
library(tidyr)
library(tibble)
library(readxl)
```

```{r}
# Convert All Text to UTF-8
sanitize_df <- function(df) {
  df[] <- lapply(df, function(col) {
    if (is.character(col)) {
      col <- ifelse(is.na(col), NA, col)  # Ensure NAs remain NAs
      col <- iconv(col, from = "UTF-8", to = "UTF-8", sub = "")  # Remove non-UTF-8 characters
      col <- gsub("[[:cntrl:]]", "", col)  # Remove control characters (but not NA)
      col <- substr(col, 1, 30000)  # Truncate long cells
    }
    return(col)
  })
  return(df)
}

read_xlsx_all <- function(url) {
  tmp <- tempfile(fileext = ".xlsx")
  download.file(url, tmp, mode = "wb")
  
  sheets <- excel_sheets(tmp)
  lapply(sheets, function(s) read_excel(tmp, sheet = s)) |>
    setNames(sheets)
}
```

## Table EV1
### Content
- MetSigDB benchmark files

Sheet order:
- Sheet 1: Resource groups.\
- Sheet 2: Metabolite exclusion comparison.\
- Sheet 3: Resource-level benchmark of resources.\
- Sheet 4: Resource-level summary.\
- Sheet 5: Term-size distribution per resource.\
- Sheet 6: MetaProViz comparison to other tools.\

```{r}
# get latest benchmark result folder
latest_folder <- 
  list.files(
    path = "MetSigDB_Benchmark/benchmark_results",
    full.names = TRUE,
    recursive = FALSE
  )

# get latest benchmark file
latest_file <-
  list.files(
    path = latest_folder,
    full.names = TRUE,
    recursive = FALSE,
    include.dirs = FALSE,
    pattern = "\\.xlsx$"
  )

sheets <- excel_sheets(latest_file)
metsigdb_bechmark <-
  lapply(sheets, function(s) read_excel(latest_file, sheet = s)) %>%
    setNames(sheets)

comparison <- read_excel("ComparisonOtherTools/Comparison.xlsx")

# Overview sheets:
df <- data.frame(
  Sheet = paste("Sheet", 1:6),
  Content = c(
    "Resource groups.",
    "Metabolite exclusion comparison.",
    "Resource-level benchmark of resources.",
    "Set-level summary.",
    "Term-size distribution per resource.",
    "MetaProViz comparison to other tools."
  ),
  Description = c(
    "Resource-to-group mapping used in this benchmark run.",
    "No exclusion vs default exclusion counts and deltas by resource and interaction type.",
    "Distinct terms, targets, interactions, and average connectivity metrics.",
    "Per-resource distribution summary of targets per term.",
    "Target membership composition across 1 term, 2 terms, and >2 terms.",
    "Comparison to MetaboAnalyst R 4.0."
  ),
  stringsAsFactors = FALSE
)

# Make lists
MetSigDB_Benchmark <- list(
  "SheetOverview" = df,
  "Resource_Groups" = metsigdb_bechmark[["Resource_Groups"]],
  "Exclusion_Comparison" = metsigdb_bechmark[["Exclusion_Compare"]],
  "Resource_Level" = metsigdb_bechmark[["Resource_Level"]],
  "Set_Level" = metsigdb_bechmark[["Set_Level"]],
  "Term_Size_Distribution" = metsigdb_bechmark[["Degree_Buckets"]],
  "Comparison_to_other_tools" = comparison
)

# Convert All Text to UTF-8
MetSigDB_Benchmark <- lapply(MetSigDB_Benchmark, sanitize_df)

# Save:
writexl::write_xlsx(MetSigDB_Benchmark, "Table_EV1.xlsx" , col_names = TRUE)

```


## Table EV2
### Content
- Previous Extended Data Table 1
- newly added sheets for the new workflow

Previous sheets:
- Sheet 1: ccRCC patients feature metadata.
- Sheet 2: Normalised data.
- Sheet 3: Metadata Analysis Anova.
- Sheet 4: Metadata Analysis Features.
- Sheet 5: Metadata Analysis Top/Bottom Features~PC.
- Sheet 6: Differential analysis TUMOR vs NORMAL.
- Sheet 7: Differential analysis TUMOR vs NORMAL of patient subset Old.
- Sheet 8: Differential analysis TUMOR vs NORMAL of patient subset Young.
- Sheet 9: MCA2Cond biological regulated clustering TvN_Young versus TvN_Old.
- Sheet 10: Summary of MCA2Cond biological regulated clustering.
- Sheet 11: Pathway enrichment analysis using KEGG on Old patient subset.
- Sheet 12: Pathway enrichment analysis using KEGG on Young patient subset.


New sheet order:

Figure 2 workflow tables and EV Figure 1
- Sheet x: Raw ccRCC patients feature metadata (MetaboliteIDs_original).
- Sheet x: ID QC id pair compatibility checks (ID_QC)
- Sheet x: Feature metadata cleaned (MetaboliteIDs_cleaned)
- Sheet x: Feature metadata translated (MetaboliteIDs_Translated)
- Sheet x: Feature metadata traverse (MetaboliteIDs_traverse)
- Sheet x: Feature metadata addEquivalent (MetaboliteIDs_addEquivalent)
- Sheet x: Processed ccRCC patients feature metadata. (Tissue_MetaData_Extended_cleaned)
- Sheet x: ID summary (id_summary)
- Sheet x: ID summary deltas (id_summary_deltas)

Keep these as before:
- Sheet x: Normalised data.
- Sheet x: Metadata Analysis Anova.
- Sheet x: Metadata Analysis Features.
- Sheet x: Metadata Analysis Top/Bottom Features~PC.
- Sheet x: Differential analysis TUMOR vs NORMAL.
- Sheet x: Differential analysis TUMOR vs NORMAL of patient subset Old.
- Sheet x: Differential analysis TUMOR vs NORMAL of patient subset Young.
- Sheet x: MCA2Cond biological regulated clustering TvN_Young versus TvN_Old.
- Sheet x: Summary of MCA2Cond biological regulated clustering.

Figure 5 workflow
- Sheet x: Pathway enrichment analysis using KEGG on Old patient subset. (new analysis, DMA_ORA_res_OLD)
- Sheet x: Pathway enrichment analysis using KEGG on Young patient subset. (new analysis, DMA_ORA_res_YOUNG)
- Sheet x: Clustering results of ORA on Young patient subset, filtered for pathways with p.adj < 0.05. (data_with_clusters)
- Sheet x: Cluster summary of ORA clustering results. (cluster_summary)
- Sheet x: Clustering similarity matrix. (similarity_matrix)

AlanineMetaboliteIDs from previous Extended Data Table 2
- Sheet x: Alanine IDs mapped to pathways from wiki, KEGG and Reactome.
- Sheet x: Mapping ambiguities for `KEGG Glycolysis/Gluconeogenesis` translating from KEGG to HMDB IDs.


```{r}

## load all data for Table EV2
data_list <- 
  read_xlsx_all(
    "https://raw.githubusercontent.com/saezlab/MetaProViz/BioRxiv/Extended%20Data%20Table%201_PatientsTissue_Metabolomics.xlsx"
  )

## Fig2 workflow data

# Original Feature Metadata
MetaboliteIDs_original <- 
  read.csv(
    "MetaProViz_Results/PK/Original/Metabolite_IDs_original.csv",
    sep=";"
  ) %>%
  select(-X)

# ID QC
ID_QC <- 
  read.csv2(
    "MetaProViz_Results/PK/ID_QC/joined_and_manually_cleaned_id_pair_tables.csv", # contains flags from where it was possibly manually curated
    sep=";"
  ) %>%
  select(-X)

# Cleaned Feature Metadata
MetaboliteIDs_cleaned <- 
  read.csv(
    "MetaProViz_Results/PK/Cleaned/MetaboliteIDs_cleaned.csv",
    sep=";"
  ) %>%
  select(-X)

# Translated Feature Metadata
MetaboliteIDs_Translated <- 
  read.csv(
    "MetaProViz_Results/PK/Translated/MetaboliteIDs_Translated.csv",
    sep=";"
  ) %>%
  select(-X)

# Traversed Feature Metadata
MetaboliteIDs_traverse <- 
  read.csv(
    "MetaProViz_Results/PK/Traverse/MetaboliteIDs_traverse.csv",
    sep=";"
  ) %>%
  select(-X)

# addEquivalent Feature Metadata
MetaboliteIDs_addEquivalent <- 
  read.csv(
    "MetaProViz_Results/PK/EquivalentID/MetaboliteIDs_addEquivalent.csv",
    sep=";"
  ) %>%
  select(-X)
 
# id summary tables
id_summary <- 
  read.csv(
    "MetaProViz_Results/PK/id_summary/id_count_df.csv",
    sep=";"
  ) %>%
  select(-X)

id_summary_deltas <- 
  read.csv(
    "MetaProViz_Results/PK/id_summary/delta_df.csv",
    sep=";"
  ) %>%
  select(-X)

# Result of Fig2 workflow "Tissue_MetaData_Extended_cleaned" containing flag column of manual ID QC and manual curation during checkmatch
Tissue_MetaData_Extended_cleaned <-
  read.csv(
    "MetaProViz_Results/PK/End_Result_Fig2/Tissue_MetaData_Extended_cleaned.csv",
    sep=";"
  ) %>%
  select(-X)



## Fig5 workflow data

DMA_ORA_res_OLD <- 
  read.csv(
    "MetaProViz_Results/Enrichment/DM_ORA_res_OLD.csv",
    sep=";"
  ) %>%
  select(-X)

DMA_ORA_res_YOUNG <- 
  read.csv(
    "MetaProViz_Results/Enrichment/DM_ORA_res_YOUNG.csv",
    sep=";"
  ) %>%
  select(-X)

data_with_clusters <-
  read.csv(
    "MetaProViz_Results/Enrichment/data_with_clusters.csv",
    sep=";"
  ) %>%
  select(-X)

cluster_summary <-
  read.csv(
    "MetaProViz_Results/Enrichment/cluster_summary.csv",
    sep=";"
  ) %>%
  select(-X)

similarity_matrix <- 
  as.matrix(
    read.csv2(
      "MetaProViz_Results/Enrichment/similarity_matrix.csv", 
      check.names = FALSE
    )
  )


# AlanineMetaboliteIDs from previous Extended Data Table 2
AlanineMetaboliteIDs <- 
  read_xlsx_all(
    "https://raw.githubusercontent.com/saezlab/MetaProViz/BioRxiv/Extended%20Data%20Table%202_AlanineMetaboliteIDs.xlsx"
  )

AlanineIDs <- AlanineMetaboliteIDs$AlanineIDs
MappingAmbiguities <- AlanineMetaboliteIDs$MappingAmbiguities

# Figure 2 workflow tables
# - Sheet x: Raw ccRCC patients feature metadata (MetaboliteIDs_original).
# - Sheet x: ID QC id pair compatibility checks (ID_QC)
# - Sheet x: Feature metadata cleaned (MetaboliteIDs_cleaned)
# - Sheet x: Feature metadata translated (MetaboliteIDs_Translated)
# - Sheet x: Feature metadata traverse (MetaboliteIDs_traverse)
# - Sheet x: Feature metadata addEquivalent (MetaboliteIDs_addEquivalent)
# - Sheet x: Feature metadata filtered for KEGG and ambiguity-resolved. (Tissue_MetaData_Extended_cleaned)
# - Sheet x: ID summary (id_summary)
# - Sheet x: ID summary deltas (id_summary_deltas)
# 
# Keep these as before:
# - Sheet x: Normalised data.
# - Sheet x: Metadata Analysis Anova.
# - Sheet x: Metadata Analysis Features.
# - Sheet x: Metadata Analysis Top/Bottom Features~PC.
# - Sheet x: Differential analysis TUMOR vs NORMAL.
# - Sheet x: Differential analysis TUMOR vs NORMAL of patient subset Old.
# - Sheet x: Differential analysis TUMOR vs NORMAL of patient subset Young.
# - Sheet x: MCA2Cond biological regulated clustering TvN_Young versus TvN_Old.
# - Sheet x: Summary of MCA2Cond biological regulated clustering.
# 
# Figure 5 workflow
# - Sheet x: Pathway enrichment analysis using KEGG on Old patient subset. (new analysis, DMA_ORA_res_OLD)
# - Sheet x: Pathway enrichment analysis using KEGG on Young patient subset. (new analysis, DMA_ORA_res_YOUNG)
# - Sheet x: Clustering results of ORA on Young patient subset, filtered for pathways with p.adj < 0.05. (data_with_clusters)
# - Sheet x: Cluster summary of ORA clustering results. (cluster_summary)
# - Sheet x: Clustering similarity matrix. (similarity_matrix)

#
# AlanineMetaboliteIDs from previous Extended Data Table 2
# - Sheet x: Alanine IDs mapped to pathways from wiki, KEGG and Reactome.
# - Sheet x: Mapping ambiguities for `KEGG Glycolysis/Gluconeogenesis` translating from KEGG to HMDB IDs.


# Overview sheets:
df <- data.frame(
  Sheet = paste("Sheet", 1:25),
  Content = c(
    "Raw ccRCC patients feature metadata.",
    "ID QC id pair compatibility checks.",
    "Feature metadata cleaned.",
    "Feature metadata translated.",
    "Feature metadata traverse.",
    "Feature metadata addEquivalent.",
    "Feature metadata filtered for KEGG and ambiguity-resolved.",
    "ID summary.",
    "ID summary deltas.",
    "Normalised data.",
    "Metadata Analysis Anova.",
    "Metadata Analysis Features.",
    "Metadata Analysis Top/Bottom Features~PC.",
    "Differential analysis TUMOR vs NORMAL.",
    "Differential analysis TUMOR vs NORMAL of patient subset Old.",
    "Differential analysis TUMOR vs NORMAL of patient subset Young.",
    "MCA2Cond biological regulated clustering TvN_Young versus TvN_Old.",
    "Summary of MCA2Cond biological regulated clustering.",
    "Pathway enrichment analysis using KEGG on Old patient subset.",
    "Pathway enrichment analysis using KEGG on Young patient subset.",
    "Clustering results of ORA on Young patient subset, filtered for pathways with p.adj < 0.05.",
    "Cluster summary of ORA clustering results.",
    "Clustering similarity matrix.",
    "Alanine IDs mapped to pathways from wiki, KEGG and Reactome.",
    "Mapping ambiguities for `KEGG Glycolysis/Gluconeogenesis` translating from KEGG to HMDB IDs."
  ),
  stringsAsFactors = FALSE
)

# Make lists
Table_EV2 <- list(
  "SheetOverview" = df,
  "Raw_Feature_Metadata" = MetaboliteIDs_original,
  "ID_QC" = ID_QC,
  "Feature_Metadata_Cleaned" = MetaboliteIDs_cleaned,
  "Feature_Metadata_Translated" = MetaboliteIDs_Translated,
  "Feature_Metadata_Traverse" = MetaboliteIDs_traverse,
  "Feature_Metadata_EquivalentID" = MetaboliteIDs_addEquivalent,
  "Feature_Metadata_KEGG" = Tissue_MetaData_Extended_cleaned,
  "ID_Summary" = id_summary,
  "ID_Summary_Deltas" = id_summary_deltas,
  "NormalisedData" = data_list[["NormalisedData"]],
  "MetaAnalysis_AOV" = data_list[["MetaAnalysis_AOV"]],
  "MetaAnalysis_Features" = data_list[["MetaAnalysis_Features"]],
  "MetaAnalysis_TopBottom" = data_list[["MetaAnalysis_TopBottom"]],
  "DMA_TUMOR_vs_NORMAL" = data_list[["DMA_TUMOR_vs_NORMAL"]],
  "DMA_TUMOR_vs_NORMAL_Old" = data_list[["DMA_TUMOR_vs_NORMAL_Old"]],
  "DMA_TUMOR_vs_NORMAL_Young" = data_list[["DMA_TUMOR_vs_NORMAL_Young"]],
  "MCA2Cond_Old-Young" = data_list[["MCA2Cond_Old-Young"]],
  "MCA2Cond_Summary" = data_list[["MCA2Cond_Summary"]],
  "KEGG_PathwayEnrichment_Old" = DMA_ORA_res_OLD,
  "KEGG_PathwayEnrichment_Young" = DMA_ORA_res_YOUNG,
  "ORA_Clustered_Data" = data_with_clusters,
  "ORA_Cluster_Summary" = cluster_summary,
  "ORA_Cluster_Similarity_Matrix" = as.data.frame(similarity_matrix),
  "AlanineMetaboliteIDs" = AlanineIDs,
  "Alanine_MappingAmbiguities" = MappingAmbiguities
)

# Convert All Text to UTF-8
Table_EV2 <- lapply(Table_EV2, sanitize_df)

# Save:
writexl::write_xlsx(Table_EV2, "Table_EV2.xlsx" , col_names = TRUE)



```



## Table EV3
### Content
- Previous Extended Data Table 3
- new sheet with merged filtered and flagged features
- new sheet containing the MetalinksDB links

New sheet order:
`Table_EV3.xlsx`:\
- Sheet 1: Raw data.\
- Sheet 2: Pool estimation.\
- Sheet 3: Media blank samples CV.\
- Sheet 4: FilteredFeatures. -> New addition\
- Sheet 5: Processed data.\
- Sheet 6: Differential Results 786-M1A versus HK2.\
- Sheet 7: Differential Results 786-M2A versus HK2.\
- Sheet 8: Differential Results 786-O versus HK2.\
- Sheet 9: Differential Results OSLM1B versus HK2.\
- Sheet 10: Differential Results OSRC2 versus HK2.\
- Sheet 11: Differential Results RFX631 versus HK2.\
- Sheet 12: Consumption-Release feature metadata.\
- Sheet 13: MCA2Cond Clustering rules.\
- Sheet 14: MCA2Cond biological regulated clustering 786M1A versus HK2.\
- Sheet 15: Summary of MCA2Cond biological regulated clustering.\
- Sheet 16: MetalinksDB metabolite-transporter, -receptor and -enzyme links to the measured metabolites. -> new addition\
```{r}

data("mca_core_rules")
mca_results <- read.csv(
    "https://raw.githubusercontent.com/saezlab/MetaProViz/BioRxiv/Cells_CoRe/MetaProViz_Results/MCAcore/mca_2cond_MCA_core_Results_2025-08-01.csv",
    sep=","
  )
mca_summary <- read.csv(
    "https://raw.githubusercontent.com/saezlab/MetaProViz/BioRxiv/Cells_CoRe/MetaProViz_Results/MCAcore/mca_2cond_MCA_core_summary_2025-08-01.csv",
    sep=","
  )

data_list <- 
  read_xlsx_all(
    "https://raw.githubusercontent.com/saezlab/MetaProViz/BioRxiv/Extended%20Data%20Table%203_Cells-Exometabolomics.xlsx"
  )

CoRe_FilteredFeatures <- 
  read.csv2(
    "Cells_CoRe/Merged_FilteredFeatures.csv"
  )

CoRe_MetalinksDB <- 
  read.csv(
    "https://raw.githubusercontent.com/saezlab/MetaProViz/BioRxiv/Cells_CoRe/MetalinksDB_BothDOWN.csv"
  )

# Overview sheets:
df <- data.frame(
  Sheet = paste("Sheet", 1:16),
  Content = c(
    "Raw data.",
    "Pool estimation.",
    "Media blank samples CV.",
    "Filtered and Flanked Features.",
    "Processed data.",
    "Differential Results 786-M1A versus HK2.",
    "Differential Results 786-M2A versus HK2.",
    "Differential Results 786-O versus HK2.",
    "Differential Results OSLM1B versus HK2.",
    "Differential Results OSRC2 versus HK2.",
    "Differential Results RFX631 versus HK2.",
    "Consumption-Release feature metadata.",
    "MCA2Cond Clustering rules.",
    "MCA2Cond biological regulated clustering 786M1A versus HK2.",
    "Summary of MCA2Cond biological regulated clustering.",
    "MetalinksDB metabolite-transporter, -receptor and -enzyme links to the measured metabolites."
  ),
  stringsAsFactors = FALSE
)

# Make lists
Cells_CoRe <- list(
  "SheetOverview" = df,
  "RawData" = data_list[["RawData"]],
  "PoolEstimationCV" = data_list[["PoolEstimationCV"]]%>%arrange(desc(CV)),
  "MediaBlankCV" = data_list[["MediaBlankCV"]]%>%arrange(desc(CV)),
  "FilteredFeatures" = CoRe_FilteredFeatures,
  "ProcessedData" = data_list[["ProcessedData"]],
  "DMA_786-M1A_vs_HK2" = data_list[["DMA_786-M1A_vs_HK2"]],
  "DMA_786-M2A_vs_HK2" = data_list[["DMA_786-M2A_vs_HK2"]],
  "DMA_786-O_vs_HK2" = data_list[["DMA_786-O_vs_HK2" ]],
  "DMA_OSLM1B_vs_HK2" = data_list[["DMA_OSLM1B_vs_HK2"]],
  "DMA_OSRC2_vs_HK2" = data_list[["DMA_OSRC2_vs_HK2"]],
  "DMA_RFX631_vs_HK2" = data_list[[ "DMA_RFX631_vs_HK2"]],
  "DMA_FeatureMetadata" = data_list[["DMA_FeatureMetadata"]],
  "Rules_MCA2Cond" = mca_core_rules,
  "MCA2Cond_786M1A_vs_HK2" = mca_results,
  "MCA2Cond_Summary" = mca_summary,
  "MetalinksDB_metabolite_links" = CoRe_MetalinksDB
)

# Convert All Text to UTF-8
Cells_CoRe <- lapply(Cells_CoRe, sanitize_df)

# Save:
writexl::write_xlsx(Cells_CoRe, "Table_EV3.xlsx" , col_names = TRUE)
```


## Table EV4
### Content
- Previous Extended Data Table 4
- new sheet with merged filtered and flagged features

New sheet order:
`Table_EV4.xlsx`:\
- Sheet 1:  Raw data.\
- Sheet 2:  Pool estimation.\
- Sheet 3:  Processed data.\
- Sheet 4:  FilteredFeatures. -> new addition\
- Sheet 5:  Sum of analytical replicates of processed data.\
- Sheet 6:  Differential Results 786-M1A versus HK2.\
- Sheet 7:  Differential Results 786-M2A versus HK2.\
- Sheet 8:  Differential Results 786-O versus HK2.\
- Sheet 9:  MCA2Cond Clustering rules.\
- Sheet 10: MCA2Cond biological regulated clustering 786-OvsHK2 versus 786-M1AvsHK2.\
- Sheet 11: Summary of MCA2Cond biological regulated clustering.\
```{r}

data_list_intra <- 
  read_xlsx_all(
    "https://raw.githubusercontent.com/saezlab/MetaProViz/BioRxiv/Extended%20Data%20Table%204_Cells_Intracellular.xlsx"
  )

Intra_FilteredFeatures <- 
  read.csv2(
    "Cells_Intracellular/Merged_FilteredFeatures.csv"
  )

# Overview sheets:
df <- data.frame(
  Sheet = paste("Sheet", 1:11),
  Content = c(
    "Raw data.",
    "Pool estimation.",
    "Processed data.",
    "Filtered and Flanked Features.",
    "Sum of analytical replicates of processed data",
    "Differential Results 786-M1A versus HK2",
    "Differential Results 786-M2A versus HK2.",
    "Differential Results 786-O versus HK2.",
    "MCA2Cond Clustering rules.",
    "MCA2Cond biological regulated clustering 786-OvsHK2 versus 786-M1AvsHK2",
    "Summary of MCA2Cond biological regulated clustering."
  ),
  stringsAsFactors = FALSE
)

# Make lists
Cells_Intra <- list(
  "SheetOverview" = df,
  "RawData" = data_list_intra[["RawData"]],
  "PoolEstimationCV" = data_list_intra[["PoolEstimationCV"]]%>%arrange(desc(CV)),
  "ProcessedData" = data_list_intra[["ProcessedData"]],
  "FilteredFeatures" = Intra_FilteredFeatures,
  "ProcessedData_Sum" = data_list_intra[["ProcessedData_Sum"]],
  "DMA_786-M1A_vs_HK2" = data_list_intra[["DMA_786-M1A_vs_HK2"]],
  "DMA_786-M2A_vs_HK2" = data_list_intra[["DMA_786-M2A_vs_HK2"]],
  "DMA_786-O_vs_HK2" = data_list_intra[["DMA_786-O_vs_HK2"]],
  "Rules_MCA2Cond" = data_list_intra[["Rules_MCA2Cond"]],
  "MCA2Cond_786-O_786-M1A" = data_list_intra[["MCA2Cond_786-O_786-M1A"]],
  "MCA2Cond_Summary" = data_list_intra[["MCA2Cond_Summary"]]
)

# Convert All Text to UTF-8
Cells_Intra <- lapply(Cells_Intra, sanitize_df)

# Save:
writexl::write_xlsx(Cells_Intra, "Table_EV4.xlsx" , col_names = TRUE)
```


# Session information

```{r session_info, echo=FALSE}
options(width = 120)
sessionInfo()
```
