---
title: "Supplementary_Tables"
author: 
  - name: Jannik Franken
      affiliation:
      - Heidelberg University
date: "14 4 2026"
output: html_document
---

# Supplementary table generation

This script iterates over all output generated and builds the supplementary tables

```{r}
library(MetaProViz)
library(magrittr)
library(dplyr)
library(tidyr)
library(tibble)
library(readxl)
```

```{r}
# Convert All Text to UTF-8
sanitize_df <- function(df) {
  df[] <- lapply(df, function(col) {
    if (is.character(col)) {
      col <- ifelse(is.na(col), NA, col)  # Ensure NAs remain NAs
      col <- iconv(col, from = "UTF-8", to = "UTF-8", sub = "")  # Remove non-UTF-8 characters
      col <- gsub("[[:cntrl:]]", "", col)  # Remove control characters (but not NA)
      col <- substr(col, 1, 30000)  # Truncate long cells
    }
    return(col)
  })
  return(df)
}

read_xlsx_all <- function(url) {
  tmp <- tempfile(fileext = ".xlsx")
  download.file(url, tmp, mode = "wb")
  
  sheets <- excel_sheets(tmp)
  lapply(sheets, function(s) read_excel(tmp, sheet = s)) |>
    setNames(sheets)
}
```

## Table EV1
### Content
- this 
- that

```{r}


```


## Table EV2
### Content
- this 
- that

```{r}


```



## Table EV3
### Content
- Previous Extended Data Table 3
- new sheet with merged filtered and flagged features
- new sheet containing the MetalinksDB links

New sheet order:
`Table_EV3.xlsx`:\
- Sheet 1: Raw data.\
- Sheet 2: Pool estimation.\
- Sheet 3: Media blank samples CV.\
- Sheet 4: FilteredFeatures. -> New addition\
- Sheet 5: Processed data.\
- Sheet 6: Differential Results 786-M1A versus HK2.\
- Sheet 7: Differential Results 786-M2A versus HK2.\
- Sheet 8: Differential Results 786-O versus HK2.\
- Sheet 9: Differential Results OSLM1B versus HK2.\
- Sheet 10: Differential Results OSRC2 versus HK2.\
- Sheet 11: Differential Results RFX631 versus HK2.\
- Sheet 12: Consumption-Release feature metadata.\
- Sheet 13: MetalinksDB metabolite-transporter, -receptor and -enzyme links to the measured metabolites. -> new addition\
```{r}

data_list <- 
  read_xlsx_all(
    "https://raw.githubusercontent.com/saezlab/MetaProViz/BioRxiv/Extended%20Data%20Table%203_Cells-Exometabolomics.xlsx"
  )

CoRe_FilteredFeatures <- 
  read.csv2(
    "Cells_CoRe/Merged_FilteredFeatures.csv"
  )

CoRe_MetalinksDB <- 
  read.csv(
    "https://raw.githubusercontent.com/saezlab/MetaProViz/BioRxiv/Cells_CoRe/MetalinksDB_BothDOWN.csv"
  )

# Overview sheets:
df <- data.frame(
  Sheet = paste("Sheet", 1:13),
  Content = c(
    "Raw data.",
    "Pool estimation.",
    "Media blank samples CV.",
    "Filtered and Flanked Features.",
    "Processed data.",
    "Differential Results 786-M1A versus HK2.",
    "Differential Results 786-M2A versus HK2.",
    "Differential Results 786-O versus HK2.",
    "Differential Results OSLM1B versus HK2.",
    "Differential Results OSRC2 versus HK2.",
    "Differential Results RFX631 versus HK2.",
    "Consumption-Release feature metadata.",
    "MetalinksDB metabolite-transporter, -receptor and -enzyme links to the measured metabolites."
  ),
  stringsAsFactors = FALSE
)

# Make lists
Cells_CoRe <- list(
  "SheetOverview" = df,
  "RawData" = data_list[["RawData"]],
  "PoolEstimationCV" = data_list[["PoolEstimationCV"]]%>%arrange(desc(CV)),
  "MediaBlankCV" = data_list[["MediaBlankCV"]]%>%arrange(desc(CV)),
  "FilteredFeatures" = CoRe_FilteredFeatures,
  "ProcessedData" = data_list[["ProcessedData"]],
  "DMA_786-M1A_vs_HK2" = data_list[["DMA_786-M1A_vs_HK2"]],
  "DMA_786-M2A_vs_HK2" = data_list[["DMA_786-M2A_vs_HK2"]],
  "DMA_786-O_vs_HK2" = data_list[["DMA_786-O_vs_HK2" ]],
  "DMA_OSLM1B_vs_HK2" = data_list[["DMA_OSLM1B_vs_HK2"]],
  "DMA_OSRC2_vs_HK2" = data_list[["DMA_OSRC2_vs_HK2"]],
  "DMA_RFX631_vs_HK2" = data_list[[ "DMA_RFX631_vs_HK2"]],
  "DMA_FeatureMetadata" = data_list[["DMA_FeatureMetadata"]],
  "MetalinksDB_metabolite_links" = CoRe_MetalinksDB
)

# Convert All Text to UTF-8
Cells_CoRe <- lapply(Cells_CoRe, sanitize_df)

# Save:
writexl::write_xlsx(Cells_CoRe, "Table_EV3.xlsx" , col_names = TRUE)
```



## Table EV4
### Content
- Previous Extended Data Table 4
- new sheet with merged filtered and flagged features

New sheet order:
`Table_EV4.xlsx`:\
- Sheet 1:  Raw data.\
- Sheet 2:  Pool estimation.\
- Sheet 3:  Processed data.\
- Sheet 4:  FilteredFeatures. -> new addition\
- Sheet 5:  Sum of analytical replicates of processed data.\
- Sheet 6:  Differential Results 786-M1A versus HK2.\
- Sheet 7:  Differential Results 786-M2A versus HK2.\
- Sheet 8:  Differential Results 786-O versus HK2.\
- Sheet 9:  MCA2Cond Clustering rules.\
- Sheet 10: MCA2Cond biological regulated clustering 786-OvsHK2 versus versus 786-M1AvsHK2.\
- Sheet 11: Summary of MCA2Cond biological regulated clustering.\
```{r}

data_list_intra <- 
  read_xlsx_all(
    "https://raw.githubusercontent.com/saezlab/MetaProViz/BioRxiv/Extended%20Data%20Table%204_Cells_Intracellular.xlsx"
  )

Intra_FilteredFeatures <- 
  read.csv2(
    "Cells_Intracellular/Merged_FilteredFeatures.csv"
  )

# Overview sheets:
df <- data.frame(
  Sheet = paste("Sheet", 1:11),
  Content = c(
    "Raw data.",
    "Pool estimation.",
    "Processed data.",
    "Filtered and Flanked Features.",
    "Sum of analytical replicates of processed data",
    "Differential Results 786-M1A versus HK2",
    "Differential Results 786-M2A versus HK2.",
    "Differential Results 786-O versus HK2.",
    "MCA2Cond Clustering rules.",
    "MCA2Cond biological regulated clustering 786-OvsHK2 versus versus 786-M1AvsHK2",
    "Summary of MCA2Cond biological regulated clustering."
  ),
  stringsAsFactors = FALSE
)

# Make lists
Cells_Intra <- list(
  "SheetOverview" = df,
  "RawData" = data_list_intra[["RawData"]],
  "PoolEstimationCV" = data_list_intra[["PoolEstimationCV"]]%>%arrange(desc(CV)),
  "ProcessedData" = data_list_intra[["ProcessedData"]],
  "FilteredFeatures" = Intra_FilteredFeatures,
  "ProcessedData_Sum" = data_list_intra[["ProcessedData_Sum"]],
  "DMA_786-M1A_vs_HK2" = data_list_intra[["DMA_786-M1A_vs_HK2"]],
  "DMA_786-M2A_vs_HK2" = data_list_intra[["DMA_786-M2A_vs_HK2"]],
  "DMA_786-O_vs_HK2" = data_list_intra[["DMA_786-O_vs_HK2"]],
  "Rules_MCA2Cond" = data_list_intra[["Rules_MCA2Cond"]],
  "MCA2Cond_786-O_786-M1A" = data_list_intra[["MCA2Cond_786-O_786-M1A"]],
  "MCA2Cond_Summary" = data_list_intra[["MCA2Cond_Summary"]]
)

# Convert All Text to UTF-8
Cells_Intra <- lapply(Cells_Intra, sanitize_df)

# Save:
writexl::write_xlsx(Cells_Intra, "Table_EV4.xlsx" , col_names = TRUE)
```



## Table EV5
### Content
- this 
- that

```{r}


```



# Session information

```{r session_info, echo=FALSE}
options(width = 120)
sessionInfo()
```
