---
#title: "**MetaProViz**"
#author: "Authors: Christina Schmidt, Dimitrios"
#date: "24 November 2022"
output:
   html_document:
    #yml: _site.yml
    css: style.css
    #theme: flatly
    #highlight: tango #`default`, `tango`, `pygments`, `kate`, `monochrome`, `espresso`, `zenburn`, `haddock`, `breezedark`, `textmate`, `arrow`, or `rstudio`
    #self_contained: true
    #toc: true
   # toc_float: true
    #toc_depth: 2
    #code_folding: show
#vignette: >
#  %\VignetteEngine{knitr::rmarkdown}
#  %\VignetteIndexEntry{MetaProViz}
#  %\VignetteEncoding{UTF-8}
#editor_options: 
#  chunk_output_type: console
---
\
<a href="#" class="btn btn-link">[**Browse source code**](https://github.com/ChristinaSchmidt1)</a>
<a href="#" class="btn btn-link">[**View on Bioconductor**](https://github.com/ChristinaSchmidt1)</a>
<a href="#" class="btn btn-link">[**Citing MetaProViz**](https://github.com/ChristinaSchmidt1)</a>
<a href="#" class="btn btn-link">[**FrezzaLab**](https://twitter.com/FrezzaLab)</a>
<a href="#" class="btn btn-link">[**SaezLab**](https://twitter.com/saezlab)</a>


```{r chunk_setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

<div class="progress progress-striped active">
  <div class="progress-bar progress-bar-success" style="width: 100%"></div>
</div>

## CoRe Metabolomics (Cell Culture Media)
Description

***
### Loading the example data
***
Here we choose an example datasets, which is publicly available on [metabolomics workbench project PR001418](https://www.metabolomicsworkbench.org/data/DRCCMetadata.php?Mode=Project&ProjectID=PR001418) including metabolic profiles of human renal epithelial cells HK2 and cell renal cell carcinoma (ccRCC) cell lines cultured in Plasmax cell culture media.\
Here we use the integrated raw peak data as example data using the trival metabolite name in combination with the KEGG ID as the metabolite identifiers.\
As part of the package you can load:\
\
**1. CoRe experiment (Cell Culture Media)** \
The raw data is available via [metabolomics workbench study ST002226](https://www.metabolomicsworkbench.org/data/DRCCMetadata.php?Mode=Study&StudyID=ST002226&StudyType=MS&ResultType=1) were exometabolomics of HK2 and ccRCC cell lines 786-O, 786-M1A, 786-M2A, OS-RC-2, OS-LM1 and RFX-631 were performed.\
\
**2. Intracellular experiment (Intra)** \
The raw data are available via [metabolomics workbench study ST002224](https://www.metabolomicsworkbench.org/data/DRCCMetadata.php?Mode=Study&StudyID=ST002224&StudyType=MS&ResultType=1) were intracellular metabolomics of HK2 and ccRCC cell lines 786-O, 786-M1A and 786-M2A were performed.\
\
**3. Pathway information mapping the trivial metabolite names to KEGG IDs and pathways (Pathways)** \
\
\
You can load the example data into your global environment using toy_data():
```{r, eval=FALSE}
devtools::load_all("C:/Users/chris/Documents/GitHub/MetaProViz")#only loaded until package is online

MetaProViz::toy_data()

# Check how our data looks like:
head(Media)
```

***
### Run MetaProViz functions
***
#### 1. Pre-processing
Discuss pitfalls of missing value imputation as discussed with Aurelien and give a clear statement! Cite the papers I have mentioned on my trello card
"Benchmark": 
- Show MVI results with different percentages (not just 50% as per half minimum!)
- Show that per feature imputation is advantageous over the "standard" imputation
- Cite the benchmark of the different methods and explain why we did not implement the others (e.g. quantile is only good for normal distribution)

```{r}


#MetaProViz::Preprocessing


```

#### 2. Differntial Metabolite Analysis (DMA)
Here explain the way we deal with 0/NA --> again refer back to above!

Here explain how we adapted the Log2FC calculation
- Benchmark: e.g. how many metabolites do we identify with the new Log2FC calculation opposed to using the absolute values (perfect example would be a metabolite that has -10 in C1 qand +11 in C2, and hence a small Log2FC)

```{r}

#MetaProViz::Preprocessing


```
Similar to DESEq2 explain what it means if a Log2FC = Inf & p.val=NA versus Log2FC= Value & p.val=NA




#### 3. Metabolite Clustering Analysis (MCA)

Create Figure (alluvial plot) explaining the general concept --> usable for manuscript as well

#### 4. Visualisation (Viz)
No replicates


\
<div class="progress progress-striped active">
  <div class="progress-bar progress-bar-success" style="width: 100%"></div>
</div>

# Session information
```{r session_info, echo=FALSE}
options(width = 120)
sessioninfo::session_info()
```
