---
title: "TutorialR"
author: "Christina Schmidt, Dimitrios,..."
date: "24 November 2022"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
editor_options: 
  chunk_output_type: console
---

# Usage MetaProViz
***MetaProViz*** (Authors. 2023. Journal. DOI: ) enables the user to pre-process metabolomics data including consumption-release (CoRe) data and contains different visualisation methods to extract biological interpretable graphs.\
This is the tutorial showcasing the different functionalities MetaProViz offers.\
For this we downloaded two publicly available datasets and use them as the example data for MetaProViz.

# Installation
[MetaProViz](Link to bioconductor) is an R package distributed as part of the Bioconductor project. To install the package, start R and enter:\
\
```{r bioconductor_install, eval=FALSE}
install.packages("BiocManager")
BiocManager::install("MetaProViz")
```
\
Alternatively, you can install the latest development version from [GitHub](https://github.com/ChristinaSchmidt1/MetaProViz) with:

```{r github_install, eval=FALSE}
BiocManager::install("ChristinaSchmidt1/MetaProViz")
```

## Libraries
[MetaProViz](Link to bioconductor) can be imported as:

```{r load_library, message=FALSE}
library(MetaProViz)

# Extra dependencies:
library(dplyr)
library(tidyverse)
```

# Input data
Example datasets are publicly available on [metabolomics workbench project PR001418](https://www.metabolomicsworkbench.org/data/DRCCMetadata.php?Mode=Project&ProjectID=PR001418) including metabolic profiles of human renal epithelial cells HK2 and cell renal cell carcinoma (ccRCC) cell lines cultured in Plasmax media.\
Here we use the integrated raw peak data as example data using the trival metabolite name in combination with the KEGG ID as the metabolite identifiers. As part of the package you can load:\
1. Intracellular experiment (Intra)
  + The raw data are available via [metabolomics workbench study ST002224](https://www.metabolomicsworkbench.org/data/DRCCMetadata.php?Mode=Study&StudyID=ST002224&StudyType=MS&ResultType=1) were intracellular metabolomics of HK2 and ccRCC cell lines 786-O, 786-M1A and 786-M2A were performed.
2. CoRe experiment (Media)
  + The raw data is available via [metabolomics workbench study ST002226](https://www.metabolomicsworkbench.org/data/DRCCMetadata.php?Mode=Study&StudyID=ST002226&StudyType=MS&ResultType=1) were exometabolomics of HK2 and ccRCC cell lines 786-O, 786-M1A, 786-M2A, OS-RC-2, OS-LM1 and RFX-631 were performed.
3. Pathway information mapping the trivial metabolite names to KEGG IDs and pathways (Pathways)
\
You can load the example data into your global environment using toy_data():
```{r}

MetaProViz::toy_data()

```


# Run MetaProViz functions

## 1. Pre-processing
### Intracellular experiment

### CoRe experiment

1. MetaProViz::Preprocessing()


```{r}

```



## 2. Biological clustering


## 3. Visualisation
No replicates




# Session information
```{r}
sessionInfo()
```


# Bibliography
```{r vignetteBiblio, results = "asis", echo = FALSE, warning = FALSE, message = FALSE}
## Print bibliography
PrintBibliography(bib, .opts = list(hyperlink = "to.doc", style = "html"))
```

