---
title: "ID Processing Workflow"
author:
  - name: Jannik Franken
    affiliation:
    - Heidelberg University
output:
  BiocStyle::html_document:
    self_contained: true
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: show
vignette: >
  %\VignetteIndexEntry{ID Processing Workflow}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
bibliography: bibliography.bib
resource_files:
  - figures/id-processing-workflow-scheme.svg
editor_options:
  chunk_output_type: console
---
```{r chunk_setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)

# Preview tables: wide tables scroll horizontally instead of stretching the page
preview_table <- function(x, ...) {
    kableExtra::scroll_box(
        kableExtra::kable_classic(
            kableExtra::kbl(x, ...),
            full_width = FALSE, html_font = "Cambria", font_size = 12
        ),
        width = "100%"
    )
}
```

This vignette demonstrates the `id_processing()` workflow for feature metadata in **MetaProViz**. The function can quantify the current ID space, perform automatic seed-ID compatibility handling, and expand the ID space through graph traversal.

The examples use the `tissue_meta` dataset that ships with MetaProViz. It contains the feature metadata of the ccRCC tissue metabolomics study by [Hakimi et al.](https://doi.org/10.1016/j.ccell.2015.12.004), where each metabolite is annotated with IDs from several databases, pathways and other feature information. Compatibility between the seed IDs is checked and resolved automatically by `seed_id_compatibility_check()`, so the IDs need no manual cleaning first.

Fig. 1 gives an overview of the steps `id_processing()` runs. Each of the workflows below enables a different combination of these steps: inspection only, compatibility check, translation, or traversal.

<center>

![Fig. 1: Overview of the `id_processing()` workflow.](figures/id-processing-workflow-scheme.svg){width="100%"}

</center>

<br>

```{r load-packages, message=FALSE, warning=FALSE}
library(MetaProViz)
library(dplyr)
library(tibble)

data(tissue_meta)
```

The workflow-managed identifier namespaces are currently `HMDB`, `KEGG`, `CHEBI`, and `PUBCHEM`. These columns are treated as the canonical ID-space columns, while other metadata columns such as names, pathway classes, or platform information are preserved across all workflow stages. `tissue_meta` has no `CHEBI` column; `id_processing()` adds it as an empty column, which the traversal later fills.

Some features have more than one ID per namespace, stored in a single cell. `id_processing()` splits these cells using one `delimiter` for all ID columns, so all ID columns must use the same separator. In `tissue_meta` this is not the case: multiple `HMDB` and `KEGG` IDs are separated by commas, whereas multiple `PUBCHEM` IDs are separated by semicolons:

```{r delimiter-check}
multi_id <- tissue_meta %>%
    select(Metabolite, HMDB, KEGG, PUBCHEM)

bind_rows(
    multi_id %>% filter(grepl(",", HMDB) | grepl(",", KEGG)) %>% head(4),
    multi_id %>% filter(grepl(";", PUBCHEM)) %>% head(2)
) %>%
    distinct() %>%
    preview_table(caption = "Features with multiple IDs in one cell.")
```

We therefore harmonise the separators to semicolons before running the workflow, and set `delimiter = ";"` in all calls below. We also remove the unnamed features (e.g. `X - 12345`), which have no IDs:

```{r prepare-metadata}
Tissue_MetaData <- tissue_meta %>%
    filter(!grepl("^X\\s*-\\s*\\d+$", Metabolite)) %>%
    mutate(across(c(HMDB, KEGG), ~ gsub(",\\s*", "; ", .x)))
```

# Inspection-only workflow

An inspection-only run quantifies the current identifier coverage and overlap structure without changing the feature metadata. The call and its result-inspection code are shown below for reference, but are not evaluated in this vignette.

```{r id-processing-default, eval=FALSE}
id_default <- id_processing(
    data = Tissue_MetaData,
    id_types = c("HMDB", "KEGG", "CHEBI", "PUBCHEM"),
    delimiter = ";",
    run_compatibility_check = FALSE,
    handle_partially_compatible = TRUE,
    handle_completely_incompatible = TRUE,
    completely_incompatible_priority = c("HMDB", "CHEBI", "PUBCHEM", "KEGG"),
    run_translation = FALSE,
    translation_from = NULL,
    translation_to = NULL,
    translation_summary = FALSE,
    run_traversal = FALSE,
    edge_table = NULL,
    compare_name_col = "Metabolite",
    save_plot = NULL,
    save_table = NULL,
    print_plot = FALSE,
    verbose = FALSE,
    path = NULL
)

names(id_default)
names(id_default$Data)
names(id_default$Plot)
names(id_default$Workflow)
preview_table(utils::head(id_default$Data$id_count_summary, 10L))
preview_table(utils::head(id_default$Data$compare_pk_summary_by_stage$input, 10L))
preview_table(utils::head(id_default$Data$count_id_tables_by_stage$input$HMDB, 10L))
```

# Compatibility-check workflow

The compatibility-only workflow can assess and handle partially or completely incompatible IDs without traversal. It is shown, but not evaluated, here.

```{r id-processing-qc, eval=FALSE}
id_qc <- id_processing(
    data = Tissue_MetaData,
    id_types = c("HMDB", "KEGG", "CHEBI", "PUBCHEM"),
    delimiter = ";",
    run_compatibility_check = TRUE,
    handle_partially_compatible = TRUE,
    handle_completely_incompatible = TRUE,
    completely_incompatible_priority = c("HMDB", "CHEBI", "PUBCHEM", "KEGG"),
    run_translation = FALSE,
    translation_from = NULL,
    translation_to = NULL,
    translation_summary = FALSE,
    run_traversal = FALSE,
    edge_table = NULL,
    compare_name_col = "Metabolite",
    save_plot = NULL,
    save_table = NULL,
    print_plot = FALSE,
    verbose = FALSE,
    path = NULL
)

preview_table(utils::head(id_qc$Data$input[, c("Metabolite", "HMDB", "KEGG", "CHEBI", "PUBCHEM")], 10L))
preview_table(utils::head(id_qc$Data$after_compatibility[, c("Metabolite", "HMDB", "KEGG", "CHEBI", "PUBCHEM")], 10L))
preview_table(utils::head(id_qc$Data$id_count_summary, 10L))
id_qc$Workflow$steps_run
names(id_qc$Data$compatibility)
id_qc$Workflow$result_overview_text
```

# Translation workflow

Translation can map one identifier namespace to one or more others. It is also shown but not evaluated; translation and traversal are mutually exclusive within one call.

```{r id-processing-translation, eval=FALSE}
id_translation <- id_processing(
    data = Tissue_MetaData,
    id_types = c("HMDB", "KEGG", "CHEBI", "PUBCHEM"),
    delimiter = ";",
    run_compatibility_check = TRUE,
    handle_partially_compatible = TRUE,
    handle_completely_incompatible = TRUE,
    completely_incompatible_priority = c("HMDB", "CHEBI", "PUBCHEM", "KEGG"),
    run_translation = TRUE,
    translation_from = "PUBCHEM",
    translation_to = c("HMDB", "KEGG"),
    translation_summary = FALSE,
    run_traversal = FALSE,
    edge_table = NULL,
    compare_name_col = "Metabolite",
    save_plot = NULL,
    save_table = NULL,
    print_plot = FALSE,
    verbose = FALSE,
    path = NULL
)

preview_table(utils::head(id_translation$Data$after_translation[, c("Metabolite", "PUBCHEM", "HMDB", "KEGG")], 10L))
preview_table(id_translation$Data$id_count_summary, caption = "Translation ID-count summary")
names(id_translation$Data$translation)
preview_table(utils::head(id_translation$Data$translation$input, 10L))
preview_table(utils::head(id_translation$Data$translation$translated_ids, 10L))
```

# Traversal workflow

This vignette runs the traversal workflow because it includes the initial ID-space inspection and compatibility handling before expanding the seed ID space through the RaMP mapping graph. The default inspection-only workflow and the translation workflow remain available through `id_processing()` when those focused analyses are needed, but are not run here. Translation and traversal are mutually exclusive within one call.

```{r id-processing-traversal, eval=TRUE}
id_traversal <- id_processing(
    data = Tissue_MetaData,
    id_types = c("HMDB", "KEGG", "CHEBI", "PUBCHEM"),
    delimiter = ";",
    run_compatibility_check = TRUE,
    handle_partially_compatible = TRUE,
    handle_completely_incompatible = TRUE,
    completely_incompatible_priority = c("HMDB", "CHEBI", "PUBCHEM", "KEGG"),
    run_translation = FALSE,
    translation_from = NULL,
    translation_to = NULL,
    translation_summary = FALSE,
    run_traversal = TRUE,
    edge_table = NULL,
    compare_name_col = "Metabolite",
    save_plot = NULL,
    save_table = NULL,
    print_plot = FALSE,
    verbose = FALSE,
    path = NULL
)
```

## Inspecting traversal results

`id_processing()` returns a list with three parts:

- **`Data`**: the result tables.
    - `input`, `after_compatibility` and `after_traversal`: the feature metadata after each stage. They keep all rows and columns of the input (here plus the `CHEBI` column, which was added because `tissue_meta` has none); only the ID columns change.
    - `id_count_summary`: one row per stage and ID type with the number of features without, with one and with several IDs.
    - `compare_pk_summary_by_stage`: for each stage, one row per feature with a flag for every ID type it has. These are the data behind the overlap plots below.
    - `count_id_tables_by_stage`: for each stage and ID type, the feature metadata with the number of IDs per feature (`entry_count`, `id_label`). These are the data behind the IDs-per-feature plots below.
    - `compatibility`: the quality control (QC) of the compatibility check. `feature_summary` has one row per feature, `pairs_before_handling` and `pairs_after_handling` one row per pair of seed IDs of a feature.
    - `traversal`: `prior_knowledge_edges`, the ID mapping graph that was used for the traversal (one row per mapping between two IDs).
- **`Plot`**: the overlap (`compare_pk()`) and IDs-per-feature (`count_id()`) plots of each stage.
- **`Workflow`**: the settings, the steps that were run, and a message per stage.

In most cases, the feature metadata after traversal is what you continue with, and the other tables explain how it was created. As examples, we compare three of these tables between the input and the end of the workflow (after traversal): the feature metadata, the ID coverage per feature, and the ID counts per ID type.

### Feature metadata

We pick four features that show the typical outcomes:

```{r id-processing-example-features}
example_features <- c(
    "1,2-propanediol",
    "1-arachidonoylglycerophosphoethanolamine*",
    "1,3-dihydroxyacetone",
    "1-oleoylglycerophosphoinositol*"
)
id_columns <- c("Metabolite", "HMDB", "KEGG", "CHEBI", "PUBCHEM")

preview_table(
    id_traversal$Data$input %>%
        filter(Metabolite %in% example_features) %>%
        select(all_of(id_columns)),
    caption = "Input: IDs of the example features."
)
preview_table(
    id_traversal$Data$after_traversal %>%
        filter(Metabolite %in% example_features) %>%
        select(all_of(id_columns)),
    caption = "After traversal: IDs of the same features."
)
```

The traversal follows the mappings between ID types and adds every ID it reaches from the seed IDs of a feature. All IDs are also written in a standard format, e.g. HMDB IDs with seven digits (`HMDB01881` becomes `HMDB0001881`), ChEBI IDs with the `CHEBI:` prefix and PubChem IDs with the `CID` prefix.

- *1-arachidonoylglycerophosphoethanolamine\** only had an HMDB ID; the traversal adds a ChEBI and a PubChem ID.
- *1,3-dihydroxyacetone* had HMDB, KEGG and PubChem IDs that all describe the same molecule, and it gains a ChEBI ID.
- *1-oleoylglycerophosphoinositol\** has no ID at all, so there is nothing to start the traversal from, and it stays without IDs.
- *1,2-propanediol* gets a different KEGG ID than it had in the input. This is the result of the compatibility check, which is the QC step before the traversal. Its HMDB and KEGG IDs could not be linked to each other, so they likely describe different molecules:

```{r id-processing-qc-example}
preview_table(
    id_traversal$Data$compatibility$pairs_before_handling %>%
        filter(Metabolite == "1,2-propanediol") %>%
        select(Metabolite, seed1_type, seed1_id, seed2_type, seed2_id,
               pair_compatible, compatibility_path, n_seed_ids, all_seed_ids_compatible),
    caption = "Compatibility check of the seed IDs of 1,2-propanediol."
)
```

Each row is one pair of seed IDs of a feature. `pair_compatible` tells whether the two IDs could be linked through the mapping graph, and `compatibility_path` how: `direct` (one ID maps to the other), `secondary` (via a third ID) or `no_match`. `all_seed_ids_compatible` is `TRUE` if all pairs of the feature are compatible. Features where no pair is compatible are *completely incompatible*. For those, `id_processing()` keeps only the ID type that comes first in `completely_incompatible_priority` (here HMDB) and removes the others. The traversal then starts from the HMDB ID alone and adds the KEGG ID that belongs to it.

### ID coverage per feature

The table returned by `compare_pk()` flags for each feature which ID types it has (`1`) or lacks (`0`). The column `None` is `1` for features without any ID:

```{r id-processing-compare-pk-example}
preview_table(
    id_traversal$Data$compare_pk_summary_by_stage$input %>%
        filter(Metabolite %in% example_features),
    caption = "Input: ID types per feature."
)
preview_table(
    id_traversal$Data$compare_pk_summary_by_stage$after_traversal %>%
        filter(Metabolite %in% example_features),
    caption = "After traversal: ID types per feature."
)
```

Here the gain is easier to see than in the ID columns: two of the example features now have all four ID types and one has three, so they can be linked to resources that use any of them.

### ID counts per ID type

`id_count_summary` adds this up over all features. For every ID type, `n_no_id`, `n_single_id` and `n_multiple_ids` count the features without, with one and with several IDs, and `n_total_ids` counts all IDs. The table also has `delta_*` columns with the change compared to the previous stage and to the input, which we leave out here:

```{r id-processing-id-count-example}
preview_table(
    id_traversal$Data$id_count_summary %>%
        filter(stage %in% c("input", "after_traversal")) %>%
        select(namespace, stage, n_features, n_no_id, n_single_id, n_multiple_ids, n_total_ids) %>%
        arrange(namespace, desc(stage == "input")),
    caption = "Input and after traversal: number of features without, with one and with several IDs per ID type."
)
```

For every ID type, the rows *input* and *after_traversal* are directly below each other. The traversal reduces the number of features without ID for every ID type and adds many IDs, most of all for ChEBI, which was not in the input at all. At the same time, more features have several IDs of one type, e.g. IDs of stereoisomers or other closely related molecules that the mapping graph links to the same feature. The next section shows the same numbers for every stage as plots.

# Traversal workflow: comparison across stages

The following plots compare the input, compatibility-handled, and traversal-expanded ID spaces.

## Identifier overlap

For each stage, `compare_pk()` shows the identifier overlap between the four managed namespaces:

```{r traversal-stage-compare-pk, results='asis', fig.width=8, fig.height=4.5, out.width='85%'}
# "after_compatibility" -> "After compatibility"
stage_label <- function(stage) {
    stage <- gsub("_", " ", stage)
    paste0(toupper(substr(stage, 1, 1)), substring(stage, 2))
}

for (stage in names(id_traversal$Plot$compare_pk_by_stage)) {
    cat("\n\n### ", stage_label(stage), "\n\n", sep = "")
    print(id_traversal$Plot$compare_pk_by_stage[[stage]])
}
```

## IDs per feature

`count_id()` shows the number of IDs per feature. `id_processing()` returns the plain `count_id()` plots, so we call `count_id()` on the metadata of each stage to show its standard MetaProViz figure (`Plot_Sized`). For each ID type, the three stages are shown next to each other: input, after compatibility check and after traversal.

```{r traversal-stage-count-id, results='asis', fig.width=17, fig.height=4.72, out.width='100%'}
id_counts <- id_traversal$Data$id_count_summary

# Short description of the ID counts of one ID type across the stages
count_text <- function(id_type) {
    x <- id_counts %>% filter(namespace == id_type)
    n <- function(stage, column) x[[column]][x$stage == stage]
    with_id <- function(stage) n(stage, "n_single_id") + n(stage, "n_multiple_ids")
    describe <- function(stage) {
        sprintf(
            "%i without, %i with one and %i with several %s IDs",
            n(stage, "n_no_id"), n(stage, "n_single_id"), n(stage, "n_multiple_ids"), id_type
        )
    }

    text <- if (with_id("input") == 0L) {
        sprintf(
            "The input has no %s IDs, so the compatibility check has nothing to check and does not change them either. All %s IDs are added by the traversal: afterwards there are %s.",
            id_type, id_type, describe("after_traversal")
        )
    } else {
        removed <- with_id("input") - with_id("after_compatibility")
        sprintf(
            "In the input there are %s. %s After traversal there are %s.",
            describe("input"),
            if (removed > 0L) {
                sprintf(
                    "The compatibility check removes the %s IDs of %i feature(s), because they did not match the other IDs of the feature.",
                    id_type, removed
                )
            } else {
                "The compatibility check does not remove any of them."
            },
            describe("after_traversal")
        )
    }

    sprintf(
        "%s In total, %i of %i features have at least one %s ID after traversal, compared to %i in the input.\n\n",
        text, with_id("after_traversal"), n("input", "n_features"), id_type, with_id("input")
    )
}

for (id_type in c("HMDB", "KEGG", "CHEBI", "PUBCHEM")) {
    cat("\n\n### ", id_type, "\n\n", sep = "")
    stage_plots <- lapply(names(id_traversal$Plot$count_id_by_stage), function(stage) {
        count_id(
            data = id_traversal$Data[[stage]],
            column = id_type,
            delimiter = ";",
            title_prefix = sprintf("%s IDs per feature: %s", id_type, tolower(stage_label(stage))),
            save_plot = NULL,
            save_table = NULL,
            print_plot = FALSE
        )$Plot_Sized
    })
    gridExtra::grid.arrange(grobs = stage_plots, ncol = 3)
    cat("\n\n", count_text(id_type), sep = "")
}
```

# Summary

```{r id-processing-summary}
coverage <- id_counts %>%
    filter(stage %in% c("input", "after_traversal")) %>%
    mutate(n_with_id = n_single_id + n_multiple_ids) %>%
    select(namespace, stage, n_with_id, n_total_ids) %>%
    tidyr::pivot_wider(names_from = stage, values_from = c(n_with_id, n_total_ids)) %>%
    transmute(
        `ID type` = namespace,
        `Features with ID (input)` = n_with_id_input,
        `Features with ID (after traversal)` = n_with_id_after_traversal,
        `IDs (input)` = n_total_ids_input,
        `IDs (after traversal)` = n_total_ids_after_traversal
    )

no_id_input <- sum(id_traversal$Data$compare_pk_summary_by_stage$input$None)
no_id_after <- sum(id_traversal$Data$compare_pk_summary_by_stage$after_traversal$None)

preview_table(coverage, caption = "ID coverage of the input and after traversal.")
```

Starting from `r nrow(Tissue_MetaData)` features, the workflow increased the number of IDs over all four ID types from `r sum(coverage[["IDs (input)"]])` to `r sum(coverage[["IDs (after traversal)"]])`. The largest gains are for ChEBI, which was not in the input, and for HMDB and PubChem. KEGG gains the least. `r if (no_id_input == no_id_after) sprintf('The number of features without any ID stays at %i', no_id_input) else sprintf('The number of features without any ID only changes from %i to %i', no_id_input, no_id_after)`: the traversal can only start from an existing ID, so features without any ID need manual annotation.

This matters when the data are linked to prior knowledge. Every resource uses its own ID type, e.g. KEGG pathways use KEGG IDs and many metabolite-protein resources use HMDB or ChEBI IDs. A feature can only be found in a resource if it has an ID of the right type, so the more features carry each ID type, the more features can be mapped, and the more complete the pathway or metabolite sets that are tested in an enrichment analysis. At the same time, features with several IDs of one type can map to several entries of the same resource and inflate an enrichment analysis. How to check and resolve this is shown in the [Prior Knowledge vignette](https://saezlab.github.io/MetaProViz/articles/pkgdown/prior-knowledge.html).

# Session information

```{r session-info}
sessionInfo()
```
