---
title: "Feature Metadata Analysis"
author:
  - name: Christina Schmidt
    affiliation:
    - Heidelberg University
output:
  html_document:
    self_contained: true
    toc: true
    toc_float: true
    toc_depth: 5
    code_folding: show
vignette: >
  %\VignetteIndexEntry{Standard Metabolomics}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
bibliography: bibliography.bib
editor_options:
  chunk_output_type: console
---

<style>
.vscroll-plot {
    width: 850px;
    height: 500px;
    overflow-y: scroll;
    overflow-x: hidden;
}
</style>

```{r chunk_setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

```{r message=FALSE, warning=FALSE}
# 1. Install Rtools if you haven’t done this yet, using the appropriate version (e.g.windows or macOS).
# 2. Install the latest development version from GitHub using devtools
# devtools::install_github("https://github.com/saezlab/MetaProViz")

#remotes::install_github('saezlab/OmnipathR@devel')
#remotes::install_github("saezlab/MetaProViz@devel")
library(MetaProViz)

# dependencies that need to be loaded:
library(magrittr)
library(dplyr)
library(rlang)
library(tidyr)
library(tibble)
library(stringr)
```

```{r}
# Make folder for results:
if (!dir.exists("MetaProViz_Results/PK")) {
  dir.create("MetaProViz_Results/PK", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/Original")) {
  dir.create("MetaProViz_Results/PK/Original", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PlusTranslatedIDs_FromCAS")) {
  dir.create("MetaProViz_Results/PK/PlusTranslatedIDs_FromCAS", recursive = TRUE)
}

##------------------------------------------------------------------
# Individual DBs
## HMDB
if (!dir.exists("MetaProViz_Results/PK/PK_HMDB")) {
  dir.create("MetaProViz_Results/PK/PK_HMDB", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PK_HMDB/Original")) {
  dir.create("MetaProViz_Results/PK/PK_HMDB/Original", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PK_HMDB/Translated")) {
  dir.create("MetaProViz_Results/PK/PK_HMDB/Translated", recursive = TRUE)
}

## KEGG
if (!dir.exists("MetaProViz_Results/PK/PK_KEGG")) {
  dir.create("MetaProViz_Results/PK/PK_KEGG", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PK_KEGG/Original")) {
  dir.create("MetaProViz_Results/PK/PK_KEGG/Original", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PK_KEGG/Translated")) {
  dir.create("MetaProViz_Results/PK/PK_KEGG/Translated", recursive = TRUE)
}

## PUBCHEM
if (!dir.exists("MetaProViz_Results/PK/PK_PUBCHEM")) {
  dir.create("MetaProViz_Results/PK/PK_PUBCHEM", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PK_PUBCHEM/Original")) {
  dir.create("MetaProViz_Results/PK/PK_PUBCHEM/Original", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PK_PUBCHEM/Translated")) {
  dir.create("MetaProViz_Results/PK/PK_PUBCHEM/Translated", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/Translated")) {
  dir.create("MetaProViz_Results/PK/Translated", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/Results_Traverse")) {
  dir.create("MetaProViz_Results/PK/Results_Traverse", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PK_HMDB/Traverse")) {
  dir.create("MetaProViz_Results/PK/PK_HMDB/Traverse", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PK_KEGG/Traverse")) {
  dir.create("MetaProViz_Results/PK/PK_KEGG/Traverse", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PK_PUBCHEM/Traverse")) {
  dir.create("MetaProViz_Results/PK/PK_PUBCHEM/Traverse", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/Traverse")) {
  dir.create("MetaProViz_Results/PK/Traverse", recursive = TRUE)
}
```

# 1. Loading the Feature metadata
Here we choose an example datasets, which is publicly available in the [paper](https://www.cell.com/cancer-cell/comments/S1535-6108(15)00468-7#supplementaryMaterial) "An Integrated Metabolic Atlas of Clear Cell Renal Cell Carcinoma", which includes metabolomic profiling on 138 matched clear cell renal cell carcinoma (ccRCC)/normal tissue pairs [@Hakimi2016]. Metabolomics was done using The company Metabolon, so this is untargeted metabolomics.\
We have updated the PubChem IDs as there were PubChem entries in the excel file that are not PubChem IDs. It is likely that this is an error caused by excel conversion as it seems in some cases the first part of the ID corresponds to one of the other IDs, so maybe these where multiple potential PubChem IDs.To see details on how this was done check out our example data branch: https://github.com/saezlab/MetaProViz/blob/ExampleData/ccRCC_Hakimi-et-all.html

Load the feature metadata:
```{r}
data(tissue_meta)

# Global separator convention for all metabolite ID columns
id_separator <- ", "

# Normalize mixed separators (",", ";") to one global format
normalize_id_cell <- function(x, out_sep = id_separator) {
  ifelse(
    is.na(x),
    NA_character_,
    purrr::map_chr(stringr::str_split(x, "[;,]"), ~ {
      vals <- stringr::str_trim(.x)
      vals <- vals[vals != ""]
      if (length(vals) == 0) NA_character_ else paste(vals, collapse = out_sep)
    })
  )
}

MetaboliteIDs <- tissue_meta %>%
  dplyr::filter(!stringr::str_detect(Metabolite, "^X\\s*-\\s*\\d+$"))%>%# only keep entries with trivial names
   dplyr::mutate(
    SUPER_PATHWAY = dplyr::coalesce(SUPER_PATHWAY, "None"),
    SUB_PATHWAY   = dplyr::coalesce(SUB_PATHWAY, "None"),
    CAS = normalize_id_cell(CAS),
    HMDB = normalize_id_cell(HMDB),
    KEGG = normalize_id_cell(KEGG),
    PUBCHEM = normalize_id_cell(PUBCHEM)
  )
```

# 2. MetaProViz Metabolite ID workflow
## Original Feature Space Exploration
1. Present metabolite IDs:
```{r, warning=FALSE}
ccRCC_CompareIDs <- MetaProViz::compare_pk(data = list(Biocft = MetaboliteIDs |> dplyr::rename("Class"="SUPER_PATHWAY")),
                                           name_col = "Metabolite",
                                           metadata_info = list(Biocft = c("KEGG", "HMDB", "PUBCHEM")),
                                           plot_name = "Overlap of ID types in ccRCC data",
                                           path= "MetaProViz_Results/PK/Original")
```
\
Here we notice that 76 features have no metabolite ID assigned, yet have a trivial name and a metabolite class assigned. For 135 metabolites we only have a pubchem ID, yet no HMDB or KEGG ID. 

2. Metabolite IDs per feature:
2.0. Harmonize ID separators:
We use one separator convention (`; `) across all ID columns (CAS, HMDB, KEGG, PUBCHEM) from this point onward.

2.1. HMDB:
```{r}
Plot1_HMDB <- MetaProViz::count_id(MetaboliteIDs, 
                                   delimiter = id_separator,
                                   "HMDB",
                                   fill_colors = c(`No ID` = "#006384", `Single ID` = "#9F0162", `Multiple IDs` = "#BEBEBE"),
                                   path = "MetaProViz_Results/PK/PK_HMDB/Original")
```

2.2. KEGG:
```{r}
Plot1_KEGG <- MetaProViz::count_id(MetaboliteIDs, 
                                   delimiter = id_separator,
                                   fill_colors = c(`No ID` = "#006384", `Single ID` = "#9F0162", `Multiple IDs` = "#BEBEBE"),
                                   "KEGG",
                                   path= "MetaProViz_Results/PK/PK_KEGG/Original")
```

2.3.: PubChem:
```{r}
Plot1_PubChem <- MetaProViz::count_id(MetaboliteIDs,
                                   delimiter = id_separator,
                                   fill_colors = c(`No ID` = "#006384", `Single ID` = "#9F0162", `Multiple IDs` = "#BEBEBE"),
                                   "PUBCHEM",
                                   path= "MetaProViz_Results/PK/PK_PUBCHEM/Original")
```

3.ID compatibility check
@Jannik
```{r}


```


## Translate Cas to metabolite ID
Features with no HMDB, KEGG or PubChem ID
```{r}
# Extract features with No HMDB, KEGG or PubChem ID:
features_no_ids <- MetaboliteIDs %>%
  dplyr::filter(
    (is.na(HMDB)    | trimws(HMDB) %in% c("", "NA")) &
    (is.na(KEGG)    | trimws(KEGG) %in% c("", "NA")) &
    (is.na(PUBCHEM) | trimws(PUBCHEM) %in% c("", "NA"))
  )

#Of those some may have a CAS number:
features_no_ids_with_cas <- features_no_ids %>%
  dplyr::filter(!(is.na(CAS) | trimws(CAS) %in% c("", "NA")))

#semicolon at the end needs ti be removed

print(paste0("We have ", nrow(features_no_ids_with_cas), " features with CAS number but no other metabolite ID and ", nrow(features_no_ids)- nrow(features_no_ids_with_cas), " features with no CAS number and no other metabolite ID."))
```

There are some cases in which we have a CAS number, yet no HMDB, PubChem or KEGG ID. We can use the CAS number to check if we can find a metabolite ID for these features.\
```{r}
features_ids_from_cas <- MetaProViz::translate_id(data = features_no_ids_with_cas, 
                                                  metadata_info = list(InputID = "CAS", grouping_variable = "SUPER_PATHWAY"), 
                                                  from = "cas", 
                                                  to = c("hmdb", "kegg", "pubchem"), 
                                                  path="MetaProViz_Results/PK/PlusTranslatedIDs_FromCAS")

# Keep only rows where at least one new ID was actually found
cas_ids_found <- features_ids_from_cas[["TranslatedDF"]] %>%
  dplyr::filter(
    !(is.na(hmdb)    | hmdb    == "") |
    !(is.na(kegg)    | kegg    == "") |
    !(is.na(pubchem) | pubchem == "")
  ) %>%
  # Replace the (empty) original ID columns with the translated ones
  dplyr::mutate(
    HMDB = normalize_id_cell(hmdb),
    KEGG = normalize_id_cell(kegg),
    PUBCHEM = normalize_id_cell(pubchem)
  ) %>%
  dplyr::select(-hmdb, -kegg, -pubchem)

# Merge back: update matching rows in MetaboliteIDs by COMP_ID
MetaboliteID_Translated <- MetaboliteIDs %>%
  dplyr::rows_update(cas_ids_found, by = "COMP_ID")
```
\
We will repeat the `count_id` and `compare_pk` plot after every change to the number of metabolite IDs:
```{r}
ccRCC_CompareIDs_Translated <- MetaProViz::compare_pk(data = list(Biocft = MetaboliteID_Translated |> dplyr::rename("Class"="SUPER_PATHWAY")),
                                           name_col = "Metabolite",
                                           metadata_info = list(Biocft = c("KEGG", "HMDB", "PUBCHEM")),
                                           plot_name = "Overlap of ID types in ccRCC data",
                                           path= "MetaProViz_Results/PK/Translated")

Plot2_HMDB <- MetaProViz::count_id(MetaboliteID_Translated,
                                   delimiter = id_separator,
                                   "HMDB",
                                   fill_colors = c(`No ID` = "#006384", `Single ID` = "#9F0162", `Multiple IDs` = "#BEBEBE"),
                                   binwidth = 1,
                                   path = "MetaProViz_Results/PK/PK_HMDB/Translated")

Plot2_KEGG <- MetaProViz::count_id(MetaboliteID_Translated,
                                   delimiter = id_separator,
                                   fill_colors = c(`No ID` = "#006384", `Single ID` = "#9F0162", `Multiple IDs` = "#BEBEBE"),
                                   "KEGG",
                                   path= "MetaProViz_Results/PK/PK_KEGG/Translated")

Plot2_PubChem <- MetaProViz::count_id(MetaboliteID_Translated,
                                   delimiter = id_separator,
                                   fill_colors = c(`No ID` = "#006384", `Single ID` = "#9F0162", `Multiple IDs` = "#BEBEBE"),
                                   "PUBCHEM",
                                   path= "MetaProViz_Results/PK/PK_PUBCHEM/Translated")
```
\
We will start to compile a summary table that will include the number of IDs  (1,2,n) after each step taken to increase the feature space. So now we will add the The CAS IDs
```{r}
# Helper: count features with at least one ID in one DB from count_id output
count_present <- function(count_id_obj) {
  tbl <- count_id_obj[["Table"]]
  sum(tbl$id_label %in% c("Single ID", "Multiple IDs"), na.rm = TRUE)
}

# Helper: count features with no HMDB, KEGG and PUBCHEM ID at all
count_none_all <- function(df) {
  sum(
    (is.na(df$HMDB)    | trimws(df$HMDB) %in% c("", "NA")) &
    (is.na(df$KEGG)    | trimws(df$KEGG) %in% c("", "NA")) &
    (is.na(df$PUBCHEM) | trimws(df$PUBCHEM) %in% c("", "NA")),
    na.rm = TRUE
  )
}

CAS_ID_Summary <- tibble::tibble(
  HMDB_Original = count_present(Plot1_HMDB),
  HMDB_Translated = count_present(Plot2_HMDB),
  HMDB_deltaOT = HMDB_Translated - HMDB_Original,
  KEGG_Original = count_present(Plot1_KEGG),
  KEGG_Translated = count_present(Plot2_KEGG),
  KEGG_deltaOT = KEGG_Translated - KEGG_Original,
  PUBCHEM_Original = count_present(Plot1_PubChem),
  PUBCHEM_Translated = count_present(Plot2_PubChem),
  PUBCHEM_deltaOT = PUBCHEM_Translated - PUBCHEM_Original,
  None_Original = count_none_all(MetaboliteIDs),
  None_Translated = count_none_all(MetaboliteID_Translated),
  None_deltaOT = None_Translated - None_Original
)

CAS_ID_Summary

# Optional detailed distribution (No ID / Single ID / Multiple IDs) by DB
CAS_ID_Label_Summary <- dplyr::bind_rows(
  Plot1_HMDB[["Table"]] %>% dplyr::count(id_label, name = "HMDB_Original"),
  Plot2_HMDB[["Table"]] %>% dplyr::count(id_label, name = "HMDB_Translated")
) %>%
  dplyr::group_by(id_label) %>%
  dplyr::summarise(
    HMDB_Original = sum(dplyr::coalesce(HMDB_Original, 0L)),
    HMDB_Translated = sum(dplyr::coalesce(HMDB_Translated, 0L)),
    .groups = "drop"
  ) %>%
  dplyr::left_join(
    dplyr::bind_rows(
      Plot1_KEGG[["Table"]] %>% dplyr::count(id_label, name = "KEGG_Original"),
      Plot2_KEGG[["Table"]] %>% dplyr::count(id_label, name = "KEGG_Translated")
    ) %>%
      dplyr::group_by(id_label) %>%
      dplyr::summarise(
        KEGG_Original = sum(dplyr::coalesce(KEGG_Original, 0L)),
        KEGG_Translated = sum(dplyr::coalesce(KEGG_Translated, 0L)),
        .groups = "drop"
      ),
    by = "id_label"
  ) %>%
  dplyr::left_join(
    dplyr::bind_rows(
      Plot1_PubChem[["Table"]] %>% dplyr::count(id_label, name = "PUBCHEM_Original"),
      Plot2_PubChem[["Table"]] %>% dplyr::count(id_label, name = "PUBCHEM_Translated")
    ) %>%
      dplyr::group_by(id_label) %>%
      dplyr::summarise(
        PUBCHEM_Original = sum(dplyr::coalesce(PUBCHEM_Original, 0L)),
        PUBCHEM_Translated = sum(dplyr::coalesce(PUBCHEM_Translated, 0L)),
        .groups = "drop"
      ),
    by = "id_label"
  ) %>%
  dplyr::mutate(
    HMDB_deltaOT = HMDB_Translated - HMDB_Original,
    KEGG_deltaOT = KEGG_Translated - KEGG_Original,
    PUBCHEM_deltaOT = PUBCHEM_Translated - PUBCHEM_Original
  ) %>%
  dplyr::arrange(factor(id_label, levels = c("No ID", "Single ID", "Multiple IDs")))

CAS_ID_Label_Summary

```

## Traverse ID
Next, we fill the gaps by expanding the metabolite identifiers across HMDB, KEGG and PubChem by building and traversing a metabolite ID graph (see methods for details).  This fills the gaps by iteratively going through the cross-database mappings and collecting all connected IDs whilst additionally adding ChEBI IDs to the features.

@Jannik
```{r}
# Harmonize ID separators for traversal (function expects one delimiter)
Input_Traverse <- MetaboliteID_Translated %>%
  dplyr::mutate(
    HMDB = normalize_id_cell(HMDB),
    KEGG = normalize_id_cell(KEGG),
    PUBCHEM = normalize_id_cell(PUBCHEM)
  )

Results_Traverse <- MetaProViz::traverse_ids(
  data = Input_Traverse,
  id_types = c("HMDB", "KEGG", "CHEBI", "PUBCHEM"),
  delimiter = ";",
  path = "MetaProViz_Results/PK/Results_Traverse"
)


```
\
First we need to take some decisions regarding detected incompatible seed IDs in 74 rows. Not all seed IDs in these rows are mutually reachable through IDEdges, suggesting they may map to different molecules. This can overexpanded the ID space during traversal. Please manually remove or correct incompatible seed IDs and rerun traverse_ids() freshly on a clean input table.
```{r}
#How can I see what are compatible once for a feature? If e.g. 2 of 3 are, How can I ignore the one an exclude the new IDs coming from this?



#Final Output:
MetaboliteID_Traverse <- Results_Traverse[["ExpandedDF"]]
```


\
We will repeat the `count_id` and `compare_pk` plot after every change to the number of metabolite IDs:
```{r}
ccRCC_CompareIDs_Traverse <- MetaProViz::compare_pk(data = list(Biocft = MetaboliteID_Traverse |> dplyr::rename("Class"="SUPER_PATHWAY")),
                                           name_col = "Metabolite",
                                           metadata_info = list(Biocft = c("KEGG", "HMDB", "PUBCHEM", "CHEBI")),
                                           plot_name = "Overlap of ID types in ccRCC data",
                                           path= "MetaProViz_Results/PK/Traverse")

Plot3_HMDB <- MetaProViz::count_id(MetaboliteID_Traverse,
                                   delimiter = id_separator,
                                   "HMDB",
                                   fill_colors = c(`No ID` = "#006384", `Single ID` = "#9F0162", `Multiple IDs` = "#BEBEBE"),
                                   binwidth = 1,
                                   path = "MetaProViz_Results/PK/PK_HMDB/Traverse")

Plot3_KEGG <- MetaProViz::count_id(MetaboliteID_Traverse,
                                   delimiter = id_separator,
                                   fill_colors = c(`No ID` = "#006384", `Single ID` = "#9F0162", `Multiple IDs` = "#BEBEBE"),
                                   "KEGG",
                                   path= "MetaProViz_Results/PK/PK_KEGG/Traverse")

Plot3_PubChem <- MetaProViz::count_id(MetaboliteID_Traverse,
                                   delimiter = id_separator,
                                   fill_colors = c(`No ID` = "#006384", `Single ID` = "#9F0162", `Multiple IDs` = "#BEBEBE"),
                                   "PUBCHEM",
                                   path= "MetaProViz_Results/PK/PK_PUBCHEM/Traverse")


```
\
We will start to compile a summary table that will include the number of IDs  (1,2,n) after each step taken to increase the feature space. So now we will add the new IDs from Traverse ID
```{r}


```


## Eqivalent ID
@Jannik --> Do we already have them from tarverse?

```{r}





# Add equivalent IDs:
Res_Equivalent <- MetaProViz::equivalent_id(data= Input_HMDB,
                                                           metadata_info = c(InputID="HMDB"),# ID in the measured data, here we use the HMDB ID
                                                           from = "hmdb",
                                                     path= "MetaProViz_Results/PK/PK_HMDB/PlusEquivalentIDs")


```

\
We will repeat the `count_id` and `compare_pk` plot after every change to the number of metabolite IDs:
```{r}


```
\
We will start to compile a summary table that will include the number of IDs  (1,2,n) after each step taken to increase the feature space. So now we will add the The CAS IDs
```{r}


```




















# 3. Integrate feature IDs with prior knowledge

Update below (is from revision branch). Here we also need to create a summary table according to reviewers request!




Given that in some cases we have multiple KEGG IDs for a measured feature, we will check if this causes mapping to multiple, different entries in the KEGG pathway-metabolite sets:
```{r}
#Load Kegg pathways:
KEGG_Pathways <- metsigdb_kegg()

#check mapping with metadata

ccRCC_to_KEGGPathways <- MetaProViz::checkmatch_pk_to_data(data = Tissue_MetaData_Extended,
                                                           input_pk = KEGG_Pathways,
                                                           metadata_info = c(InputID = "KEGG", PriorID = "MetaboliteID", grouping_variable = "term"))

problems_terms <- ccRCC_to_KEGGPathways[["GroupingVariable_summary"]]%>%
  filter(!Group_Conflict_Notes== "None")
```
```{r, echo=FALSE}
problems_terms[,c(1,5:8)]%>%
  kableExtra::kbl(caption = "Terms in KEGG pathways where the same measured feature mapps with more than one ID, which will inflate the enrichment analysis.") %>%
  kableExtra::kable_classic(full_width = F, html_font = "Cambria", font_size = 12)
```
\
Dependent on the biological question and the organism and prior knowledge, one can either maintain the metabolite ID of the more likely metabolite (e.g. in human its moire likely that we have L-aminoacid than D-aminoacid) or the metabolite ID that is represented in more/less pathways (specificity). If we are looking into the cases where we do have multiple IDs, for cases with no match to the prior knowledge we can just maintain one ID, whilst for cases with exactly one match we should maintain the ID that is found in the prior knowledge. 
```{r}
# Select the cases where a feature has multiple IDs
multipleIDs <- ccRCC_to_KEGGPathways[["data_summary"]]%>%
  filter(original_count>1)
```
```{r, echo=FALSE}
multipleIDs[,c(1,12, 5:6, 3, 9, 11, 13)]%>%
  kableExtra::kbl(caption = "Terms in KEGG pathways where the same measured feature mapps with more than one ID, which may inflate the enrichment analysis.") %>%
  kableExtra::kable_classic(full_width = F, html_font = "Cambria", font_size = 12)
```
\
In case of `ActionRequired=="Check"`, we can look into the column `Action_Specific` which contains additional information. In case of the entry `KeepEachID`, multiple matches to the prior knowledge were found, yet the features are in different pathways (=GroupingVariable). Yet, in case of `KeepOneID`, the different IDs map to the same pathway in the prior knowledge for at least one case and therefore keeping both would inflate the enrichment analysis.\
```{r}
SelectedIDs <- ccRCC_to_KEGGPathways[["data_summary"]]%>%
  #Expand rows where Action == KeepEachID by splitting `matches`
  dplyr::mutate(matches_split = if_else(Action_Specific == "KeepEachID", matches, NA_character_)) %>%
  separate_rows(matches_split, sep = ",\\s*") %>%
  mutate(InputID_select = if_else(Action_Specific  == "KeepEachID", matches_split, InputID_select)) %>%
  select(-matches_split) %>%
  #Select one ID for AcionSpecific==KeepOneID
  dplyr::mutate(InputID_select = case_when(
    Action_Specific == "KeepOneID" & matches == "C03460, C03722" ~ "C03722", # 2-Methylprop-2-enoyl-CoA versus Quinolinate. No evidence, hence we keep the one present in more pathways ( C03722=7 pathways, C03460=2 pathway)
    Action_Specific == "KeepOneID" & matches ==  "C00221, C00031" ~ "C00031", # These are D- and L-Glucose. We have human samples, so in this conflict we will maintain L-Glucose
    Action_Specific == "KeepOneID" & matches ==  "C17737, C00695" ~ "C00695", # Allocholic acid versus Cholic acid. No evidence, hence we keep the one present in more pathways (C00695 = 4 pathways, C17737 = 1 pathway)
    Action_Specific == "KeepOneID" ~ InputID_select,  # Keep NA where not matched manually
    TRUE ~ InputID_select
  )) 
```
\
Lastly, we need to add the column including our selected IDs to the metadata table:
```{r}
Tissue_MetaData_Extended <- merge(x= SelectedIDs %>%
                                        dplyr::select(KEGG, InputID_select),
                                  y= Tissue_MetaData_Extended,
                                  by= "KEGG",
                                  all.y=TRUE)
```


This final table will be used as input for the enrichment analysis





# Session information

```{r session_info, echo=FALSE}
options(width = 120)
sessionInfo()
```

# Bibliography
