---
title: "Feature Metadata Analysis"
author:
  - name: Christina Schmidt
    affiliation:
    - Heidelberg University
output:
  html_document:
    self_contained: true
    toc: true
    toc_float: true
    toc_depth: 5
    code_folding: show
vignette: >
  %\VignetteIndexEntry{Standard Metabolomics}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
bibliography: bibliography.bib
editor_options:
  chunk_output_type: console
---

<style>
.vscroll-plot {
    width: 850px;
    height: 500px;
    overflow-y: scroll;
    overflow-x: hidden;
}
</style>

```{r chunk_setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

```{r message=FALSE, warning=FALSE}
# 1. Install Rtools if you haven’t done this yet, using the appropriate version (e.g.windows or macOS).
# 2. Install the latest development version from GitHub using devtools
# devtools::install_github("https://github.com/saezlab/MetaProViz")

#remotes::install_github('saezlab/OmnipathR@devel')
#remotes::install_github("saezlab/MetaProViz@devel")
library(MetaProViz)

# dependencies that need to be loaded:
library(magrittr)
library(dplyr)
library(rlang)
library(tidyr)
library(tibble)
library(stringr)
```

```{r}
# Make folder for results:
if (!dir.exists("MetaProViz_Results/PK")) {
  dir.create("MetaProViz_Results/PK", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/Original")) {
  dir.create("MetaProViz_Results/PK/Original", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PlusTranslatedIDs_FromCAS")) {
  dir.create("MetaProViz_Results/PK/PlusTranslatedIDs_FromCAS", recursive = TRUE)
}

##------------------------------------------------------------------
# Individual DBs
## HMDB
if (!dir.exists("MetaProViz_Results/PK/PK_HMDB")) {
  dir.create("MetaProViz_Results/PK/PK_HMDB", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PK_HMDB/Original")) {
  dir.create("MetaProViz_Results/PK/PK_HMDB/Original", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PK_HMDB/Translated")) {
  dir.create("MetaProViz_Results/PK/PK_HMDB/Translated", recursive = TRUE)
}

## KEGG
if (!dir.exists("MetaProViz_Results/PK/PK_KEGG")) {
  dir.create("MetaProViz_Results/PK/PK_KEGG", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PK_KEGG/Original")) {
  dir.create("MetaProViz_Results/PK/PK_KEGG/Original", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PK_KEGG/Translated")) {
  dir.create("MetaProViz_Results/PK/PK_KEGG/Translated", recursive = TRUE)
}

## PUBCHEM
if (!dir.exists("MetaProViz_Results/PK/PK_PUBCHEM")) {
  dir.create("MetaProViz_Results/PK/PK_PUBCHEM", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PK_PUBCHEM/Original")) {
  dir.create("MetaProViz_Results/PK/PK_PUBCHEM/Original", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PK_PUBCHEM/Translated")) {
  dir.create("MetaProViz_Results/PK/PK_PUBCHEM/Translated", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/Translated")) {
  dir.create("MetaProViz_Results/PK/Translated", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/Results_Traverse")) {
  dir.create("MetaProViz_Results/PK/Results_Traverse", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PK_HMDB/Traverse")) {
  dir.create("MetaProViz_Results/PK/PK_HMDB/Traverse", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PK_KEGG/Traverse")) {
  dir.create("MetaProViz_Results/PK/PK_KEGG/Traverse", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PK_PUBCHEM/Traverse")) {
  dir.create("MetaProViz_Results/PK/PK_PUBCHEM/Traverse", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/Traverse")) {
  dir.create("MetaProViz_Results/PK/Traverse", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/ID_QC")) {
  dir.create("MetaProViz_Results/PK/ID_QC", recursive = TRUE)
}
```

# 1. Loading the Feature metadata
Here we choose an example datasets, which is publicly available in the [paper](https://www.cell.com/cancer-cell/comments/S1535-6108(15)00468-7#supplementaryMaterial) "An Integrated Metabolic Atlas of Clear Cell Renal Cell Carcinoma", which includes metabolomic profiling on 138 matched clear cell renal cell carcinoma (ccRCC)/normal tissue pairs [@Hakimi2016]. Metabolomics was done using The company Metabolon, so this is untargeted metabolomics.\
We have updated the PubChem IDs as there were PubChem entries in the excel file that are not PubChem IDs. It is likely that this is an error caused by excel conversion as it seems in some cases the first part of the ID corresponds to one of the other IDs, so maybe these where multiple potential PubChem IDs.To see details on how this was done check out our example data branch: https://github.com/saezlab/MetaProViz/blob/ExampleData/ccRCC_Hakimi-et-all.html

Load the feature metadata:
```{r}
data(tissue_meta)

# Global separator convention for all metabolite ID columns
id_separator <- ", "

# Normalize mixed separators (",", ";") to one global format
normalize_id_cell <- function(x, out_sep = id_separator) {
  ifelse(
    is.na(x),
    NA_character_,
    purrr::map_chr(stringr::str_split(x, "[;,]"), ~ {
      vals <- stringr::str_trim(.x)
      vals <- vals[vals != ""]
      if (length(vals) == 0) NA_character_ else paste(vals, collapse = out_sep)
    })
  )
}

MetaboliteIDs <- tissue_meta %>%
  dplyr::filter(!stringr::str_detect(Metabolite, "^X\\s*-\\s*\\d+$"))%>%# only keep entries with trivial names
   dplyr::mutate(
    SUPER_PATHWAY = dplyr::coalesce(SUPER_PATHWAY, "None"),
    SUB_PATHWAY   = dplyr::coalesce(SUB_PATHWAY, "None"),
    CAS = normalize_id_cell(CAS),
    HMDB = normalize_id_cell(HMDB),
    KEGG = normalize_id_cell(KEGG),
    PUBCHEM = normalize_id_cell(PUBCHEM)
  )
```

# 2. MetaProViz Metabolite ID workflow
## Original Feature Space Exploration
1. Present metabolite IDs:
```{r, warning=FALSE}
ccRCC_CompareIDs <- MetaProViz::compare_pk(data = list(Biocft = MetaboliteIDs |> dplyr::rename("Class"="SUPER_PATHWAY")),
                                           name_col = "Metabolite",
                                           metadata_info = list(Biocft = c("KEGG", "HMDB", "PUBCHEM")),
                                           plot_name = "Overlap of ID types in ccRCC data",
                                           path= "MetaProViz_Results/PK/Original")
```
\
Here we notice that 76 features have no metabolite ID assigned, yet have a trivial name and a metabolite class assigned. For 135 metabolites we only have a pubchem ID, yet no HMDB or KEGG ID. 

2. Metabolite IDs per feature:
2.0. Harmonize ID separators:
We use one separator convention (`; `) across all ID columns (CAS, HMDB, KEGG, PUBCHEM) from this point onward.

2.1. HMDB:
```{r}
Plot1_HMDB <- MetaProViz::count_id(MetaboliteIDs, 
                                   delimiter = id_separator,
                                   "HMDB",
                                   fill_colors = c(`No ID` = "#006384", `Single ID` = "#9F0162", `Multiple IDs` = "#BEBEBE"),
                                   path = "MetaProViz_Results/PK/PK_HMDB/Original")
```

2.2. KEGG:
```{r}
Plot1_KEGG <- MetaProViz::count_id(MetaboliteIDs, 
                                   delimiter = id_separator,
                                   fill_colors = c(`No ID` = "#006384", `Single ID` = "#9F0162", `Multiple IDs` = "#BEBEBE"),
                                   "KEGG",
                                   path= "MetaProViz_Results/PK/PK_KEGG/Original")
```

2.3.: PubChem:
```{r}
Plot1_PubChem <- MetaProViz::count_id(MetaboliteIDs,
                                   delimiter = id_separator,
                                   fill_colors = c(`No ID` = "#006384", `Single ID` = "#9F0162", `Multiple IDs` = "#BEBEBE"),
                                   "PUBCHEM",
                                   path= "MetaProViz_Results/PK/PK_PUBCHEM/Original")
```

3.ID compatibility check
@Jannik
```{r}

MetaboliteIDs_compatibility_check <- seed_id_compatibility_check(
    data = MetaboliteIDs,
    id_types = c("HMDB", "KEGG", "CHEBI", "PUBCHEM"),
    delimiter = ","
)

##  Note: the compatibility check does not assess CAS ID compatibility
## Note: The input df does not contain ChEBI IDs. For the compatibility check, 
##       ChEBI IDs are internally used as possible stepstones for ID compatibility

## Look at the compatibility check results
MetaboliteIDs_pair_compatibility <- MetaboliteIDs_compatibility_check$ID_pair_compatibility
MetaboliteIDs_data_with_compatibility <- MetaboliteIDs_compatibility_check$data_with_compatibility

## how many features of the "MetaboliteIDs" df have a completely compatible list of IDs? (only considering PUBCHEM, KEGG & HMDB)
length(which(MetaboliteIDs_data_with_compatibility$all_seed_ids_compatible == TRUE))
# 503

## how many features are not completely compatible (can still have partial compatibility)?
length(which(MetaboliteIDs_data_with_compatibility$all_seed_ids_compatible == FALSE))
# 74

## for all ID pairs across features,
## how many direct matches were there?
length(which(MetaboliteIDs_pair_compatibility$compatibility_path == "direct"))
#790

## how many secondary matches were there?
length(which(MetaboliteIDs_pair_compatibility$compatibility_path == "secondary"))
# 47

## how many were no_match? -> these contain rows, which only had CAS or only one of HMDB, KEGG, PUBCHEM but not two
length(which(MetaboliteIDs_pair_compatibility$compatibility_path == "no_match"))
# 351


```


```{r}
## assess ID compatibility and remove incompatible IDs


# Option 1: 
#   go through the pair_compatibility permutation table
#   exclude all incompatible ID pairs
#       - sometimes there is less than two HMDB, KEGG, PUBCHEM seed IDs.
#         Maybe there is only a CAS or only one of the other ID types.
#         In this case, we don't want to exclude them, because we can not really
#         assess compatibility. Therefore, "all_seed_ids_compatible" is TRUE
#         -> keep all where "all_seed_ids_compatible" == TRUE
#       - in the Results_Traverse$ID_pair_compatibility (traverse_ID_pair_compatibility),
#         we have the column "original_row_id" to re-group seed_id-pairs later
#       - we need to exclude seed_id combinations without a match for all rows where
#         "all_seed_ids_compatible" == FALSE
#       - manually decide for features with multiple, all-incompatible IDs (this is the most important point here)

## a good example is original_row_id == 99 of the original df, the metabolite 4-hydroxyglutamate

# > MetaboliteIDs_pair_compatibility[MetaboliteIDs_pair_compatibility$"original_row_id" == "99", ]

#   original_row_id Metabolite   CAS   SUPER_PATHWAY  ...   PUBCHEM KEGG  HMDB  CHEBI row_id seed1_type seed1_id seed2_type seed2_id pair_compatible compatibility_path n_seed_ids all_seed_ids_compatible
#   <chr>           <chr>        <chr> <chr>          ...   <chr>   <chr> <chr> <chr>  <int> <chr>      <chr>    <chr>      <chr>    <lgl>           <chr>                   <int> <lgl>                  
# 1 99              4-hydroxygl~ 2485~ Amino acid     ...   546007~ C030~ HMDB~ NA        99 HMDB       HMDB000~ KEGG       C03079   FALSE           no_match                    4 FALSE
# 2 99              4-hydroxygl~ 2485~ Amino acid     ...   546007~ C030~ HMDB~ NA        99 HMDB       HMDB000~ PUBCHEM    5460078  FALSE           no_match                    4 FALSE
# 3 99              4-hydroxygl~ 2485~ Amino acid     ...   546007~ C030~ HMDB~ NA        99 HMDB       HMDB000~ PUBCHEM    5460800  FALSE           no_match                    4 FALSE
# 4 99              4-hydroxygl~ 2485~ Amino acid     ...   546007~ C030~ HMDB~ NA        99 KEGG       C03079   PUBCHEM    5460078  TRUE            direct                      4 FALSE 
# 5 99              4-hydroxygl~ 2485~ Amino acid     ...   546007~ C030~ HMDB~ NA        99 KEGG       C03079   PUBCHEM    5460800  TRUE            direct                      4 FALSE 
# 6 99              4-hydroxygl~ 2485~ Amino acid     ...   546007~ C030~ HMDB~ NA        99 PUBCHEM    5460078  PUBCHEM    5460800  TRUE            secondary                   4 FALSE 

## here, we have "all_seed_ids_compatible" == FALSE, but we want to keep some of the combinations
## we also have no_match direct and secondary matches, which makes it a good example
## we want to keep all direct and secondary matches and discard no_match

MetaboliteIDs_pair_compatibility_filtered <- MetaboliteIDs_pair_compatibility %>% 
  filter( 
    (all_seed_ids_compatible == TRUE) | 
    (all_seed_ids_compatible == FALSE & compatibility_path == "direct") | 
    (all_seed_ids_compatible == FALSE & compatibility_path == "secondary") 
  )


# let's assess, if we lost any whole features due to this filtering
## these are these features:
length(unique(MetaboliteIDs_pair_compatibility$Metabolite)) #577
length(unique(MetaboliteIDs_pair_compatibility_filtered$Metabolite)) #544

# that means, 33 whole features would get lost if we filter strictly like above

# let's investigate these "lost" features and see why they are filtered out

lost_Metabolites_after_filtering <- list( 
  setdiff(
      unique(MetaboliteIDs_pair_compatibility$Metabolite),
      unique(MetaboliteIDs_pair_compatibility_filtered$Metabolite)
  )
)

# now look at all seed_id compatibility checks of these lost features
lost_id_pair_rows <- MetaboliteIDs_pair_compatibility[ MetaboliteIDs_pair_compatibility$Metabolite %in% lost_Metabolites_after_filtering[[1]]  , ]

# length(unique(lost_id_pair_rows$Metabolite))
# > 33

## as we can see, these are exactly the 33 rows missing and they have one thing in common:
## They have either 2 or 3 seed_ids AND -> all their seed_id pairs mismatched,
## i.e. they had no single seed_id pair which was compatible

## for these cases, we should not just delete them but handle them differently

lost_id_pair_rows_with_cas <- lost_id_pair_rows[is.na(lost_id_pair_rows$CAS) == FALSE, ]
lost_id_pair_rows_no_cas <- lost_id_pair_rows[is.na(lost_id_pair_rows$CAS) == TRUE, ]

## add the original CAS name to the "lost_id_pair_rows_with_cas" table to check 
## if the originally assigned CAS IDs match the tivial names

# get CAS number to trivial Name mappings from CAS through OmnipathR
RaMP_table_source <- OmnipathR::ramp_table("source")
RaMP_table_CAS <- RaMP_table_source %>%
  filter(IDtype == "CAS") %>%
  mutate(
    sourceId = sub("^CAS:", "", sourceId)
  ) %>%
  select(
    sourceId,
    commonName
  )

# join correct CAS commonName to lost_id_pair_rows_with_cas
lost_id_pair_rows_with_cas <- lost_id_pair_rows_with_cas %>%
  left_join(
    RaMP_table_CAS %>%
      rename(CAS_commonName = commonName),
    by = c("CAS" = "sourceId")
  ) %>%
  relocate(CAS_commonName, .after = Metabolite)
  
## still, some CAS numbers do not match.

## save the two tables:
"MetaProViz_Results/PK/ID_QC"

write.csv(
  lost_id_pair_rows_with_cas,
  "MetaProViz_Results/PK/ID_QC/lost_id_pair_rows_with_cas.csv"
)
write.csv(
  lost_id_pair_rows_no_cas,
  "MetaProViz_Results/PK/ID_QC/lost_id_pair_rows_no_cas.csv"
) 


#@Jannik
## next steps:
# - print the tables with and with no cas using kable extra in this Rmd
# - @Christina manually resolves the IDs
# - get back manually resolved IDs
# - wrangle them back into a joint table for further analysis
#     - join the table back together
#     - keep "all_seed_ids_compatible"
#     - flag:
#         - all incompatible but CAS solved
#         - all incompatible but manually solved
#         - at least one seed_id pair compatible, incompatible pairs removed, feature kept

# - make correct print messages for the knitted HTMLs
# - build summary_counting table in the beginning
#     - contains original, clean, translated(cas), tranverse, equivalent_id
#       deltas, counts for noID, singleID, multipleID, totalIDs




## some thoughts about manually resolving the IDs:

# we will now manually investigate all of these cases and decide on IDs to keep
# looking at lost_id_pair_rows, we decide on an ID to keep and edit the


# How to resolve manually:
# is the Metabolite name correct in comparison to the ID name?
# is the CAS number compatible to one of the HMDB, KEGG or PUBCHEM IDs?
# for features, which have all-incompatible HMDB-KEGG-PUBCHEM seed_ids, keep those with CAS






### continue workflow HERE:

## re-join the filtered and manually fit list into a new DF to proceed working

## NOTE: keep "all_seed_ids_compatible"
## also FLAG: all incompatible but CAS solved
##            all incompatible but manually solved
##            at least one seed_id pair compatible, incompatible pairs removed, feature kept

MetaboliteIDs_filtered <- MetaboliteIDs_pair_compatibility_filtered_manually %>%
  group_by(original_row_id) %>%
  summarise(
    Metabolite    = first(Metabolite),
    CAS           = first(CAS),
    SUPER_PATHWAY = first(SUPER_PATHWAY),
    SUB_PATHWAY   = first(SUB_PATHWAY),
    COMP_ID       = first(COMP_ID),
    PLATFORM      = first(PLATFORM),
    RI            = first(RI),
    MASS          = first(MASS),
    
    PUBCHEM = {
      x <- trimws(as.character(PUBCHEM))
      x[x == ""] <- NA_character_
      x <- unique(na.omit(x))
      if (length(x) == 0) NA_character_ else paste(x, collapse = ", ")
    },
    
    KEGG = {
      x <- trimws(as.character(KEGG))
      x[x == ""] <- NA_character_
      x <- unique(na.omit(x))
      if (length(x) == 0) NA_character_ else paste(x, collapse = ", ")
    },
    
    HMDB = {
      x <- trimws(as.character(HMDB))
      x[x == ""] <- NA_character_
      x <- unique(na.omit(x))
      if (length(x) == 0) NA_character_ else paste(x, collapse = ", ")
    },
    
    CHEBI = {
      x <- trimws(as.character(CHEBI))
      x[x == ""] <- NA_character_
      x <- unique(na.omit(x))
      if (length(x) == 0) NA_character_ else paste(x, collapse = ", ")
    },
    
    .groups = "drop"
  ) %>%
  select(
    Metabolite,
    CAS,
    SUPER_PATHWAY,
    SUB_PATHWAY,
    COMP_ID,
    PLATFORM,
    RI,
    MASS,
    PUBCHEM,
    KEGG,
    HMDB,
    CHEBI
  )



```

```{r}
# run plots again




#@Jannik


# then initialize counting table with No id, singleID, multipleID, totalIDs
# then after every step, run count_id functions again and expand the counting table with the new steps

```


## Translate Cas to metabolite ID
Features with no HMDB, KEGG or PubChem ID
```{r}
# Extract features with No HMDB, KEGG or PubChem ID:
features_no_ids <- MetaboliteIDs %>%
  dplyr::filter(
    (is.na(HMDB)    | trimws(HMDB) %in% c("", "NA")) &
    (is.na(KEGG)    | trimws(KEGG) %in% c("", "NA")) &
    (is.na(PUBCHEM) | trimws(PUBCHEM) %in% c("", "NA"))
  )

#Of those some may have a CAS number:
features_no_ids_with_cas <- features_no_ids %>%
  dplyr::filter(!(is.na(CAS) | trimws(CAS) %in% c("", "NA")))

#semicolon at the end needs to be removed

print(paste0("We have ", nrow(features_no_ids_with_cas), " features with CAS number but no other metabolite ID and ", nrow(features_no_ids)- nrow(features_no_ids_with_cas), " features with no CAS number and no other metabolite ID."))
```

There are some cases in which we have a CAS number, yet no HMDB, PubChem or KEGG ID. We can use the CAS number to check if we can find a metabolite ID for these features.\
```{r}
features_ids_from_cas <- MetaProViz::translate_id(data = features_no_ids_with_cas, 
                                                  metadata_info = list(InputID = "CAS", grouping_variable = "SUPER_PATHWAY"), 
                                                  from = "cas", 
                                                  to = c("hmdb", "kegg", "pubchem"), 
                                                  path="MetaProViz_Results/PK/PlusTranslatedIDs_FromCAS")

# Keep only rows where at least one new ID was actually found
cas_ids_found <- features_ids_from_cas[["TranslatedDF"]] %>%
  dplyr::filter(
    !(is.na(hmdb)    | hmdb    == "") |
    !(is.na(kegg)    | kegg    == "") |
    !(is.na(pubchem) | pubchem == "")
  ) %>%
  # Replace the (empty) original ID columns with the translated ones
  dplyr::mutate(
    HMDB = normalize_id_cell(hmdb),
    KEGG = normalize_id_cell(kegg),
    PUBCHEM = normalize_id_cell(pubchem)
  ) %>%
  dplyr::select(-hmdb, -kegg, -pubchem)

# Merge back: update matching rows in MetaboliteIDs by COMP_ID
MetaboliteID_Translated <- MetaboliteIDs %>%
  dplyr::rows_update(cas_ids_found, by = "COMP_ID")
```
\
We will repeat the `count_id` and `compare_pk` plot after every change to the number of metabolite IDs:
```{r}
ccRCC_CompareIDs_Translated <- MetaProViz::compare_pk(data = list(Biocft = MetaboliteID_Translated |> dplyr::rename("Class"="SUPER_PATHWAY")),
                                           name_col = "Metabolite",
                                           metadata_info = list(Biocft = c("KEGG", "HMDB", "PUBCHEM")),
                                           plot_name = "Overlap of ID types in ccRCC data",
                                           path= "MetaProViz_Results/PK/Translated")

Plot2_HMDB <- MetaProViz::count_id(MetaboliteID_Translated,
                                   delimiter = id_separator,
                                   "HMDB",
                                   fill_colors = c(`No ID` = "#006384", `Single ID` = "#9F0162", `Multiple IDs` = "#BEBEBE"),
                                   binwidth = 1,
                                   path = "MetaProViz_Results/PK/PK_HMDB/Translated")

Plot2_KEGG <- MetaProViz::count_id(MetaboliteID_Translated,
                                   delimiter = id_separator,
                                   fill_colors = c(`No ID` = "#006384", `Single ID` = "#9F0162", `Multiple IDs` = "#BEBEBE"),
                                   "KEGG",
                                   path= "MetaProViz_Results/PK/PK_KEGG/Translated")

Plot2_PubChem <- MetaProViz::count_id(MetaboliteID_Translated,
                                   delimiter = id_separator,
                                   fill_colors = c(`No ID` = "#006384", `Single ID` = "#9F0162", `Multiple IDs` = "#BEBEBE"),
                                   "PUBCHEM",
                                   path= "MetaProViz_Results/PK/PK_PUBCHEM/Translated")
```
\
We will start to compile a summary table that will include the number of IDs  (1,2,n) after each step taken to increase the feature space. So now we will add the The CAS IDs
```{r}
# Helper: count features with at least one ID in one DB from count_id output
count_present <- function(count_id_obj) {
  tbl <- count_id_obj[["Table"]]
  sum(tbl$id_label %in% c("Single ID", "Multiple IDs"), na.rm = TRUE)
}

# Helper: count features with no HMDB, KEGG and PUBCHEM ID at all
count_none_all <- function(df) {
  sum(
    (is.na(df$HMDB)    | trimws(df$HMDB) %in% c("", "NA")) &
    (is.na(df$KEGG)    | trimws(df$KEGG) %in% c("", "NA")) &
    (is.na(df$PUBCHEM) | trimws(df$PUBCHEM) %in% c("", "NA")),
    na.rm = TRUE
  )
}

CAS_ID_Summary <- tibble::tibble(
  HMDB_Original = count_present(Plot1_HMDB),
  HMDB_Translated = count_present(Plot2_HMDB),
  HMDB_deltaOT = HMDB_Translated - HMDB_Original,
  KEGG_Original = count_present(Plot1_KEGG),
  KEGG_Translated = count_present(Plot2_KEGG),
  KEGG_deltaOT = KEGG_Translated - KEGG_Original,
  PUBCHEM_Original = count_present(Plot1_PubChem),
  PUBCHEM_Translated = count_present(Plot2_PubChem),
  PUBCHEM_deltaOT = PUBCHEM_Translated - PUBCHEM_Original,
  None_Original = count_none_all(MetaboliteIDs),
  None_Translated = count_none_all(MetaboliteID_Translated),
  None_deltaOT = None_Translated - None_Original
)

CAS_ID_Summary

# Optional detailed distribution (No ID / Single ID / Multiple IDs) by DB
CAS_ID_Label_Summary <- dplyr::bind_rows(
  Plot1_HMDB[["Table"]] %>% dplyr::count(id_label, name = "HMDB_Original"),
  Plot2_HMDB[["Table"]] %>% dplyr::count(id_label, name = "HMDB_Translated")
) %>%
  dplyr::group_by(id_label) %>%
  dplyr::summarise(
    HMDB_Original = sum(dplyr::coalesce(HMDB_Original, 0L)),
    HMDB_Translated = sum(dplyr::coalesce(HMDB_Translated, 0L)),
    .groups = "drop"
  ) %>%
  dplyr::left_join(
    dplyr::bind_rows(
      Plot1_KEGG[["Table"]] %>% dplyr::count(id_label, name = "KEGG_Original"),
      Plot2_KEGG[["Table"]] %>% dplyr::count(id_label, name = "KEGG_Translated")
    ) %>%
      dplyr::group_by(id_label) %>%
      dplyr::summarise(
        KEGG_Original = sum(dplyr::coalesce(KEGG_Original, 0L)),
        KEGG_Translated = sum(dplyr::coalesce(KEGG_Translated, 0L)),
        .groups = "drop"
      ),
    by = "id_label"
  ) %>%
  dplyr::left_join(
    dplyr::bind_rows(
      Plot1_PubChem[["Table"]] %>% dplyr::count(id_label, name = "PUBCHEM_Original"),
      Plot2_PubChem[["Table"]] %>% dplyr::count(id_label, name = "PUBCHEM_Translated")
    ) %>%
      dplyr::group_by(id_label) %>%
      dplyr::summarise(
        PUBCHEM_Original = sum(dplyr::coalesce(PUBCHEM_Original, 0L)),
        PUBCHEM_Translated = sum(dplyr::coalesce(PUBCHEM_Translated, 0L)),
        .groups = "drop"
      ),
    by = "id_label"
  ) %>%
  dplyr::mutate(
    HMDB_deltaOT = HMDB_Translated - HMDB_Original,
    KEGG_deltaOT = KEGG_Translated - KEGG_Original,
    PUBCHEM_deltaOT = PUBCHEM_Translated - PUBCHEM_Original
  ) %>%
  dplyr::arrange(factor(id_label, levels = c("No ID", "Single ID", "Multiple IDs")))

CAS_ID_Label_Summary

```

## Traverse ID
Next, we fill the gaps by expanding the metabolite identifiers across HMDB, KEGG and PubChem by building and traversing a metabolite ID graph (see methods for details).  This fills the gaps by iteratively going through the cross-database mappings and collecting all connected IDs whilst additionally adding ChEBI IDs to the features.

@Jannik
```{r}

# Harmonize ID separators for traversal (function expects one delimiter)
Input_Traverse <- MetaboliteID_Translated %>%
  dplyr::mutate(
    HMDB = normalize_id_cell(HMDB),
    KEGG = normalize_id_cell(KEGG),
    PUBCHEM = normalize_id_cell(PUBCHEM)
  )

Results_Traverse <- MetaProViz::traverse_ids(
  data = Input_Traverse,
  id_types = c("HMDB", "KEGG", "CHEBI", "PUBCHEM"),
  delimiter = ",",
  path = "MetaProViz_Results/PK/Results_Traverse"
)

## inspect results
traverse_ID_pair_compatibility <- Results_Traverse$ID_pair_compatibility
traverse_ID_edges <- Results_Traverse$ID_Edges_prior_knowledge
traverse_ID_ExpandedDF <- Results_Traverse$ExpandedDF


```
\
First we need to take some decisions regarding detected incompatible seed IDs in 74 rows. Not all seed IDs in these rows are mutually reachable through IDEdges, suggesting they may map to different molecules. This can overexpanded the ID space during traversal. Please manually remove or correct incompatible seed IDs and rerun traverse_ids() freshly on a clean input table.
```{r}
#How can I see what are compatible once for a feature? If e.g. 2 of 3 are, How can I ignore the one an exclude the new IDs coming from this?

# @Christina: The above comment should be sorted out now that we have the pair_compatibility testing

# How do we proceed from here?

# Option 1: 
#   go through the pair_compatibility permutation table
#   exclude all incompatible ID pairs
#       - sometimes there is less than two HMDB, KEGG, PUBCHEM seed IDs.
#         Maybe there is only a CAS or only one of the other ID types.
#         In this case, we don't want to exclude them, because we can not really
#         assess compatibility. Therefore, "all_seed_ids_compatible" is TRUE
#         -> keep all where "all_seed_ids_compatible" == TRUE
#       - in the Results_Traverse$ID_pair_compatibility (traverse_ID_pair_compatibility),
#         we have the column "original_row_id" to re-group seed_id-pairs later
#       - we need to exclude seed_id combinations without a match for all rows where
#         "all_seed_ids_compatible" == FALSE

## a good example is original_row_id == 99 of the original df, the metabolite 4-hydroxyglutamate

# > traverse_ID_pair_compatibility[traverse_ID_pair_compatibility$"original_row_id" == "99", ]

#   original_row_id Metabolite   CAS   SUPER_PATHWAY  ...   PUBCHEM KEGG  HMDB  CHEBI row_id seed1_type seed1_id seed2_type seed2_id pair_compatible compatibility_path n_seed_ids all_seed_ids_compatible
#   <chr>           <chr>        <chr> <chr>          ...   <chr>   <chr> <chr> <chr>  <int> <chr>      <chr>    <chr>      <chr>    <lgl>           <chr>                   <int> <lgl>                  
# 1 99              4-hydroxygl~ 2485~ Amino acid     ...   546007~ C030~ HMDB~ NA        99 HMDB       HMDB000~ KEGG       C03079   FALSE           no_match                    4 FALSE
# 2 99              4-hydroxygl~ 2485~ Amino acid     ...   546007~ C030~ HMDB~ NA        99 HMDB       HMDB000~ PUBCHEM    5460078  FALSE           no_match                    4 FALSE
# 3 99              4-hydroxygl~ 2485~ Amino acid     ...   546007~ C030~ HMDB~ NA        99 HMDB       HMDB000~ PUBCHEM    5460800  FALSE           no_match                    4 FALSE
# 4 99              4-hydroxygl~ 2485~ Amino acid     ...   546007~ C030~ HMDB~ NA        99 KEGG       C03079   PUBCHEM    5460078  TRUE            direct                      4 FALSE 
# 5 99              4-hydroxygl~ 2485~ Amino acid     ...   546007~ C030~ HMDB~ NA        99 KEGG       C03079   PUBCHEM    5460800  TRUE            direct                      4 FALSE 
# 6 99              4-hydroxygl~ 2485~ Amino acid     ...   546007~ C030~ HMDB~ NA        99 PUBCHEM    5460078  PUBCHEM    5460800  TRUE            secondary                   4 FALSE 

## here, we have "all_seed_ids_compatible" == FALSE, but we want to keep some of the combinations
## we also have no_match direct and secondary matches, which makes it a good example
## we want to keep all direct and secondary matches and discard no_match

traverse_ID_pair_compatibility_filtered <- traverse_ID_pair_compatibility %>% 
  filter( 
    (all_seed_ids_compatible == TRUE) | 
    (all_seed_ids_compatible == FALSE & compatibility_path == "direct") | 
    (all_seed_ids_compatible == FALSE & compatibility_path == "secondary") 
  )


# let's assess, if we lost any whole features due to this filtering
## these are these features:
length(unique(traverse_ID_pair_compatibility$Metabolite)) #577
length(unique(traverse_ID_pair_compatibility_filtered$Metabolite)) #544

# that means, 33 whole features would get lost if we filter strictly like above

# let's investigate these "lost" features and see why they are filtered out

lost_rows_after_filtering <- list( 
  setdiff(
      unique(traverse_ID_pair_compatibility$Metabolite),
      unique(traverse_ID_pair_compatibility_filtered$Metabolite)
  )
)

# now look at all seed_id compatibility checks of these lost features
lost_id_pair_rows <- traverse_ID_pair_compatibility[ traverse_ID_pair_compatibility$Metabolite %in% lost_rows_after_filtering[[1]]  , ]

# length(unique(lost_id_pair_rows$Metabolite))
# > 33

## as we can see, these are exactly the 33 rows missing and they have one thing in common:
## They have either 2 or 3 seed_ids AND -> all their seed_id pairs mismatched,
## i.e. they had no single seed_id pair which was compatible

## for these cases, we should not just delete them but handle them differently

# we will now manually investigate all of these cases and decide on one ID to keep
# looking at lost_id_pair_rows, we dedice on an ID to keep and edit the 














# now we want to re-join the filtered seed_id pairs into the original wide-format table

# re-group into a new dataframe "MetaboliteIDs_Translated_filtered" and
# make it ordered in the same way as the "MetaboliteID_Translated" before

MetaboliteIDs_filtered <- traverse_ID_pair_compatibility_filtered %>%
  group_by(original_row_id) %>%
  summarise(
    Metabolite    = first(Metabolite),
    CAS           = first(CAS),
    SUPER_PATHWAY = first(SUPER_PATHWAY),
    SUB_PATHWAY   = first(SUB_PATHWAY),
    COMP_ID       = first(COMP_ID),
    PLATFORM      = first(PLATFORM),
    RI            = first(RI),
    MASS          = first(MASS),
    
    PUBCHEM = {
      x <- trimws(as.character(PUBCHEM))
      x[x == ""] <- NA_character_
      x <- unique(na.omit(x))
      if (length(x) == 0) NA_character_ else paste(x, collapse = ", ")
    },
    
    KEGG = {
      x <- trimws(as.character(KEGG))
      x[x == ""] <- NA_character_
      x <- unique(na.omit(x))
      if (length(x) == 0) NA_character_ else paste(x, collapse = ", ")
    },
    
    HMDB = {
      x <- trimws(as.character(HMDB))
      x[x == ""] <- NA_character_
      x <- unique(na.omit(x))
      if (length(x) == 0) NA_character_ else paste(x, collapse = ", ")
    },
    
    CHEBI = {
      x <- trimws(as.character(CHEBI))
      x[x == ""] <- NA_character_
      x <- unique(na.omit(x))
      if (length(x) == 0) NA_character_ else paste(x, collapse = ", ")
    },
    
    .groups = "drop"
  ) %>%
  select(
    Metabolite,
    CAS,
    SUPER_PATHWAY,
    SUB_PATHWAY,
    COMP_ID,
    PLATFORM,
    RI,
    MASS,
    PUBCHEM,
    KEGG,
    HMDB,
    CHEBI
  )








#Final Output:
MetaboliteID_Traverse <- Results_Traverse[["ExpandedDF"]]
```


\
We will repeat the `count_id` and `compare_pk` plot after every change to the number of metabolite IDs:
```{r}
ccRCC_CompareIDs_Traverse <- MetaProViz::compare_pk(data = list(Biocft = MetaboliteID_Traverse |> dplyr::rename("Class"="SUPER_PATHWAY")),
                                           name_col = "Metabolite",
                                           metadata_info = list(Biocft = c("KEGG", "HMDB", "PUBCHEM", "CHEBI")),
                                           plot_name = "Overlap of ID types in ccRCC data",
                                           path= "MetaProViz_Results/PK/Traverse")

Plot3_HMDB <- MetaProViz::count_id(MetaboliteID_Traverse,
                                   delimiter = id_separator,
                                   "HMDB",
                                   fill_colors = c(`No ID` = "#006384", `Single ID` = "#9F0162", `Multiple IDs` = "#BEBEBE"),
                                   binwidth = 1,
                                   path = "MetaProViz_Results/PK/PK_HMDB/Traverse")

Plot3_KEGG <- MetaProViz::count_id(MetaboliteID_Traverse,
                                   delimiter = id_separator,
                                   fill_colors = c(`No ID` = "#006384", `Single ID` = "#9F0162", `Multiple IDs` = "#BEBEBE"),
                                   "KEGG",
                                   path= "MetaProViz_Results/PK/PK_KEGG/Traverse")

Plot3_PubChem <- MetaProViz::count_id(MetaboliteID_Traverse,
                                   delimiter = id_separator,
                                   fill_colors = c(`No ID` = "#006384", `Single ID` = "#9F0162", `Multiple IDs` = "#BEBEBE"),
                                   "PUBCHEM",
                                   path= "MetaProViz_Results/PK/PK_PUBCHEM/Traverse")


```
\
We will start to compile a summary table that will include the number of IDs  (1,2,n) after each step taken to increase the feature space. So now we will add the new IDs from Traverse ID
```{r}

## extend counting dataframe with traverse_id results


```


## Eqivalent ID
@Jannik --> Do we already have them from tarverse?

```{r}





# Add equivalent IDs:
Res_Equivalent <- MetaProViz::equivalent_id(data= Input_HMDB,
                                                           metadata_info = c(InputID="HMDB"),# ID in the measured data, here we use the HMDB ID
                                                           from = "hmdb",
                                                     path= "MetaProViz_Results/PK/PK_HMDB/PlusEquivalentIDs")


```

\
We will repeat the `count_id` and `compare_pk` plot after every change to the number of metabolite IDs:
```{r}


```
\
We will start to compile a summary table that will include the number of IDs  (1,2,n) after each step taken to increase the feature space. So now we will add the The CAS IDs
```{r}


```







```{r}

## summary couting table including all deltas as bargraph / lollipop graph


```













# 3. Integrate feature IDs with prior knowledge

Update below (is from revision branch). Here we also need to create a summary table according to reviewers request!




Given that in some cases we have multiple KEGG IDs for a measured feature, we will check if this causes mapping to multiple, different entries in the KEGG pathway-metabolite sets:
```{r}
#Load Kegg pathways:
KEGG_Pathways <- metsigdb_kegg()

#check mapping with metadata

ccRCC_to_KEGGPathways <- MetaProViz::checkmatch_pk_to_data(data = Tissue_MetaData_Extended,
                                                           input_pk = KEGG_Pathways,
                                                           metadata_info = c(InputID = "KEGG", PriorID = "MetaboliteID", grouping_variable = "term"))

problems_terms <- ccRCC_to_KEGGPathways[["GroupingVariable_summary"]]%>%
  filter(!Group_Conflict_Notes== "None")
```
```{r, echo=FALSE}
problems_terms[,c(1,5:8)]%>%
  kableExtra::kbl(caption = "Terms in KEGG pathways where the same measured feature mapps with more than one ID, which will inflate the enrichment analysis.") %>%
  kableExtra::kable_classic(full_width = F, html_font = "Cambria", font_size = 12)
```
\
Dependent on the biological question and the organism and prior knowledge, one can either maintain the metabolite ID of the more likely metabolite (e.g. in human its moire likely that we have L-aminoacid than D-aminoacid) or the metabolite ID that is represented in more/less pathways (specificity). If we are looking into the cases where we do have multiple IDs, for cases with no match to the prior knowledge we can just maintain one ID, whilst for cases with exactly one match we should maintain the ID that is found in the prior knowledge. 
```{r}
# Select the cases where a feature has multiple IDs
multipleIDs <- ccRCC_to_KEGGPathways[["data_summary"]]%>%
  filter(original_count>1)
```
```{r, echo=FALSE}
multipleIDs[,c(1,12, 5:6, 3, 9, 11, 13)]%>%
  kableExtra::kbl(caption = "Terms in KEGG pathways where the same measured feature mapps with more than one ID, which may inflate the enrichment analysis.") %>%
  kableExtra::kable_classic(full_width = F, html_font = "Cambria", font_size = 12)
```
\
In case of `ActionRequired=="Check"`, we can look into the column `Action_Specific` which contains additional information. In case of the entry `KeepEachID`, multiple matches to the prior knowledge were found, yet the features are in different pathways (=GroupingVariable). Yet, in case of `KeepOneID`, the different IDs map to the same pathway in the prior knowledge for at least one case and therefore keeping both would inflate the enrichment analysis.\
```{r}
SelectedIDs <- ccRCC_to_KEGGPathways[["data_summary"]]%>%
  #Expand rows where Action == KeepEachID by splitting `matches`
  dplyr::mutate(matches_split = if_else(Action_Specific == "KeepEachID", matches, NA_character_)) %>%
  separate_rows(matches_split, sep = ",\\s*") %>%
  mutate(InputID_select = if_else(Action_Specific  == "KeepEachID", matches_split, InputID_select)) %>%
  select(-matches_split) %>%
  #Select one ID for AcionSpecific==KeepOneID
  dplyr::mutate(InputID_select = case_when(
    Action_Specific == "KeepOneID" & matches == "C03460, C03722" ~ "C03722", # 2-Methylprop-2-enoyl-CoA versus Quinolinate. No evidence, hence we keep the one present in more pathways ( C03722=7 pathways, C03460=2 pathway)
    Action_Specific == "KeepOneID" & matches ==  "C00221, C00031" ~ "C00031", # These are D- and L-Glucose. We have human samples, so in this conflict we will maintain L-Glucose
    Action_Specific == "KeepOneID" & matches ==  "C17737, C00695" ~ "C00695", # Allocholic acid versus Cholic acid. No evidence, hence we keep the one present in more pathways (C00695 = 4 pathways, C17737 = 1 pathway)
    Action_Specific == "KeepOneID" ~ InputID_select,  # Keep NA where not matched manually
    TRUE ~ InputID_select
  )) 
```
\
Lastly, we need to add the column including our selected IDs to the metadata table:
```{r}
Tissue_MetaData_Extended <- merge(x= SelectedIDs %>%
                                        dplyr::select(KEGG, InputID_select),
                                  y= Tissue_MetaData_Extended,
                                  by= "KEGG",
                                  all.y=TRUE)
```


This final table will be used as input for the enrichment analysis





# Session information

```{r session_info, echo=FALSE}
options(width = 120)
sessionInfo()
```

# Bibliography
