---
title: "Enrichment Analysis"
author:
  - name: Christina Schmidt
    affiliation:
    - Heidelberg University
  - name: Jannik Franken
    affiliation:
    - Heidelberg University
output:
  html_document:
    self_contained: true
    toc: true
    toc_float: true
    toc_depth: 5
    code_folding: show
vignette: >
  %\VignetteIndexEntry{Standard Metabolomics}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
bibliography: bibliography.bib
editor_options:
  chunk_output_type: console
---

<style>
.vscroll-plot {
    width: 850px;
    height: 500px;
    overflow-y: scroll;
    overflow-x: hidden;
}
</style>

```{r chunk_setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

```{r message=FALSE, warning=FALSE}
# 1. Install Rtools if you haven’t done this yet, using the appropriate version (e.g.windows or macOS).
# 2. Install the latest development version from GitHub using devtools
# devtools::install_github("https://github.com/saezlab/MetaProViz")

library(MetaProViz)

# dependencies that need to be loaded:
library(magrittr)
library(dplyr)
library(rlang)
library(tidyr)
library(tibble)
library(stringr)
library(readxl)
```

```{r}
# Make folder for results: @Jannik add folders as needed!
if (!dir.exists("MetaProViz_Results/Enrichment")) {
  dir.create("MetaProViz_Results/Enrichment", recursive = TRUE)
}

```

```{r}
# Convert All Text to UTF-8
sanitize_df <- function(df) {
  df[] <- lapply(df, function(col) {
    if (is.character(col)) {
      col <- ifelse(is.na(col), NA, col)  # Ensure NAs remain NAs
      col <- iconv(col, from = "UTF-8", to = "UTF-8", sub = "")  # Remove non-UTF-8 characters
      col <- gsub("[[:cntrl:]]", "", col)  # Remove control characters (but not NA)
      col <- substr(col, 1, 30000)  # Truncate long cells
    }
    return(col)
  })
  return(df)
}

# Open Excel file from BioRxiv branch and read in the data
read_xlsx_all <- function(url) {
  tmp <- tempfile(fileext = ".xlsx")
  download.file(url, tmp, mode = "wb")
  
  sheets <- readxl::excel_sheets(tmp)
  lapply(sheets, function(s) read_excel(tmp, sheet = s)) |>
    setNames(sheets)
}


```

# Load results
## 1. Loading the Feature metadata
Result from Fig2_FeatureMetadata, which include the feature metadata table with extended metabolite IDs and solved ambiguity issues for mapping KEGG IDs to KEGG pathways.

```{r}
# load Results from Fig2_FeatureMetadata

Tissue_MetaData_Extended_cleaned <- 
    read.csv2(
        "MetaProViz_Results/PK/End_Result_Fig2/Tissue_MetaData_Extended_cleaned.csv"
    ) %>%
    select(
        -X
    ) 

```


## 2. Biological regulated clustering Results
Load Extended Data Table 1 , Load Sheet 9 from BiorXiv

```{r}
Extended_Data_Table_1 <- read_xlsx_all(
  "https://raw.githubusercontent.com/saezlab/MetaProViz/BioRxiv/Extended%20Data%20Table%201_PatientsTissue_Metabolomics.xlsx"
)

# names(Extended_Data_Table_1)
DMA_TUMOR_vs_NORMAL_Old <- Extended_Data_Table_1$"DMA_TUMOR_vs_NORMAL_Old"
DMA_TUMOR_vs_NORMAL_Young <- Extended_Data_Table_1$"DMA_TUMOR_vs_NORMAL_Young"


# Some metabolite names in the Extended Data Table 1 contain semicolons, while they are substituted by commas in the Tissue_MetaData_Extended_cleaned
# change semicolons in Metabolite column to commas

DMA_TUMOR_vs_NORMAL_Old <- DMA_TUMOR_vs_NORMAL_Old %>%
  mutate(
      Metabolite = gsub(";", ",", Metabolite)
  )

DMA_TUMOR_vs_NORMAL_Young <- DMA_TUMOR_vs_NORMAL_Young %>%
  mutate(
      Metabolite = gsub(";", ",", Metabolite)
  )

Tissue_MetaData_Extended_cleaned[Tissue_MetaData_Extended_cleaned$Metabolite == "docosapentaenoate (n3 DPA 22:5n3)", "Metabolite"] <- "docosapentaenoate (n3 DPA, 22:5n3)"

```



# Pathway enrichment

Next, we perform Over Representation Analysis (`ORA`) using KEGG pathways for each comparison.\
```{r, warning=FALSE}
# Since we have performed multiple comparisons (all_vs_HK2), we will run ORA for each of this comparison
DM_ORA_res <- list()

KEGG_Pathways <- MetaProViz::metsigdb_kegg()

## manually clean up KEGG_Pathways with the few instances we saw in Fig2 during checkmatch_pk_to_data
KEGG_Pathways_clean <- 
    KEGG_Pathways %>%
    filter(
        !(MetaboliteID == "C01087" & term == "Metabolic pathways"),  # Remove (R)-2-Hydroxyglutarate (C01087) from KEGG pathway "Metabolic Pathways"
        !(MetaboliteID == "C00668" & term %in% c(    # C00092, C00668: remove C00668 from Pathways which contain both of them
            "Metabolic pathways",
            "Biosynthesis of secondary metabolites"
        ))
    )




# continue
# comparison = "Young"
for(comparison in c("Old", "Young")){
  
  df_selected <- 
    if (comparison == "Old") {
        DMA_TUMOR_vs_NORMAL_Old
    } else if (comparison == "Young") {
        DMA_TUMOR_vs_NORMAL_Young
    } else {
        stop("Invalid comparison value")
    }
  
  dma_res <- merge(
      x = Tissue_MetaData_Extended_cleaned,
      y = df_selected,
      by = "Metabolite",
      all = TRUE
  )
  
  
  #@Jannik:
  # 1. Long DF --> Columns KEGG IDs, Log2FC, pval, tval, padj. --> duplicate based on mutliple KEGG IDS --> duplicated Log2fc etc.
  # 2. For duplicated KEGG IDs --> keep higher absolute Log2FC
  # 3. --> KEGG IDs as col names --> input for DMA
  
  
  # Ensure unique IDs and full background --> we include measured features that do not have a KEGG ID.
  dma_res_clean <- 
      dma_res %>%
          select(
              Metabolite,
              KEGG,
              InputID_select,
              Log2FC,
              t.val,
              p.val,
              p.adj
          ) %>%
          mutate(  # make one KEGG column from "KEGG" and "InputID_select"
              KEGG = case_when(
                  InputID_select == "KEGG" ~ KEGG,
                  .default = InputID_select
              )
          ) %>%
          select(
              - InputID_select
          ) %>%
          mutate(  # mark original rows
              .row_id = row_number()
          ) %>%
          separate_rows(  # make long-format for cases with two KEGG IDs
              KEGG,
              sep = ",\\s*"
          ) %>%
          
          group_by( # create index per original row
              .row_id
          ) %>%
          mutate(
            .dup_index = row_number(),
            .n = n()
          ) %>%
          ungroup() %>%
          mutate(   #  only add suffix if there was splitting (n > 1)
              Metabolite = if_else(
                  .n > 1,
                  paste0(Metabolite, "_", .dup_index),
                  Metabolite
              )
          ) %>%
          mutate( 
              KEGG = if_else(
                  is.na(KEGG),
                  paste0("NA_", cumsum(is.na(KEGG))),
                  KEGG
              )
          ) %>%
          group_by(   
              KEGG
          ) %>%
          slice_max(    # For duplicated KEGG IDs --> keep higher absolute Log2FC
              order_by = Log2FC, n = 1, with_ties = FALSE
          ) %>%
          ungroup() %>%
          column_to_rownames(
              "KEGG"
          ) %>%
          rename(
              Metabolite_Name = Metabolite
          ) %>%
          select(
              - ".row_id",
              - ".dup_index",
              - ".n"
          )
  
  
  #Perform ORA
  set.seed(100)
  Res <- MetaProViz::standard_ora(
      data = dma_res_clean,    # Input data requirements: column `t.val` and column `Metabolite`
      metadata_info = c(
          pvalColumn = "p.adj",
          percentageColumn = "t.val",
          PathwayTerm = "term",
          PathwayFeature = "MetaboliteID"
      ),
      input_pathway = KEGG_Pathways_clean,  # Pathway file requirements: column `term`, `Metabolite` and `Description`. Above we loaded the Kegg_Pathways using MetaProViz::Load_KEGG()
      pathway_name = paste0("KEGG_", comparison, sep=""),
      min_gssize = 3,
      max_gssize = 1000,
      cutoff_stat = 0.01,
      cutoff_percentage = 10
  )

  DM_ORA_res[[comparison]] <- Res

  #Select to plot:
  Res_Select <- Res[["ClusterGosummary"]] %>%
    filter(p.adjust < 0.1)%>%
    #filter(pvalue<0.05)%>%
    filter(percentage_of_Pathway_detected > 10)

  if(!is.null(Res_Select)){
      MetaProViz::viz_volcano(
          plot_types = "PEA",
          data = dma_res_clean, # Must be the data you have used as an input for the pathway analysis
          data2 = as.data.frame(Res_Select) %>% dplyr::rename("term"="ID"),
          metadata_info = c(
              PEA_Pathway = "term",# Needs to be the same in both, metadata_feature and data2.
              PEA_stat = "p.adjust",#Column data2
              PEA_score = "GeneRatio",#Column data2
              PEA_Feature = "MetaboliteID"
          ), # Column metadata_feature (needs to be the same as row names in data)
          metadata_feature = KEGG_Pathways,#Must be the pathways used for pathway analysis
          plot_name = paste("KEGG_", comparison, sep = ""),
          select_label = NULL,
          subtitle = "PEA" 
      )
  }
}
```

## Use the results to showcase networks

```{r}

cluster_pk_input <- 
    DM_ORA_res$Young$ClusterGosummary %>%
      filter(
          p.adjust < 0.05
      )

cluster_pk_res <- 
    cluster_pk(
        cluster_pk_input,
        metadata_info = c(
            metabolite_column = "Metabolites_in_pathway",
            pathway_column = "ID"
        ),
        input_format = "enrichment",
        similarity = "jaccard",
        threshold = 0.6,
        plot_threshold = 0.4,
        clust = "community",
        min = 1,
        node_size_column = "percentage_of_Pathway_detected",
        save_plot = NULL,
        plot_name = "Enrichment_example",
        print_plot = FALSE,
        min_degree = 0,
        show_density = TRUE,
        max_nodes = 1000
    )

head(cluster_pk_res$cluster_summary)
cluster_pk_res$graph_plot

set.seed(123)
viz_graph(
    similarity_matrix = cluster_pk_res$similarity_matrix,
    clusters = cluster_pk_res$clusters,
    plot_threshold = 0,
    plot_name = "Enrichment_example_plot_threshold_0",
    max_nodes = 1000,
    min_degree = 0,
    node_sizes = cluster_pk_res$node_sizes,
    show_density = TRUE,
    save_plot = "svg",
    print_plot = FALSE,
    path = "MetaProViz_Results/Enrichment"
)

```


# Save the results

```{r}
# Previous Extended Data Table 1 sheets:
# "SheetOverview"
# "Feature metadata"
# "NormalisedData"
# "MetaAnalysis_AOV"
# "MetaAnalysis_Features"
# "MetaAnalysis_TopBottom"
# "DMA_TUMOR_vs_NORMAL"
# "DMA_TUMOR_vs_NORMAL_Old"
# "DMA_TUMOR_vs_NORMAL_Young"
# "MCA2Cond_Old-Young"
# "MCA2Cond_Summary"
# "KEGG_PathwayEnrichment_Old"    -> replace
# "KEGG_PathwayEnrichment_Young"  -> replace

# new results generated here:
# 1. DM_ORA_res$Old      -> replace old one
# 2. DM_ORA_res$Young    -> replace old one
# 3. cluster_pk_res
#     a.  $data                   -> original data plus cluster association column
#     b.  $cluster_summary        -> cluster, n_terms, pct_terms
#     c.  $clusters               -> long format cluster to PK term mapping
#     d.  $similarity_matrix      -> [0-1], 0 = no overlap, 1 = equality, diagonal 1
#     e.  $distance_matrix        -> [0-1], = 1 - similarity matrix, diagonal 0
#     f.  $node_sizes             -> long formal node_size to term mapping
#     g.  $graph_plot             -> S3 igraph object

write.csv2(
    DM_ORA_res$Old$ClusterGosummary,
    "MetaProViz_Results/Enrichment/DM_ORA_res_OLD.csv"
)

write.csv2(
    DM_ORA_res$Young$ClusterGosummary,
    "MetaProViz_Results/Enrichment/DM_ORA_res_YOUNG.csv"
)

write.csv2(
    cluster_pk_res$data,
    "MetaProViz_Results/Enrichment/data_with_clusters.csv"
)

write.csv2(
    cluster_pk_res$cluster_summary,
    "MetaProViz_Results/Enrichment/cluster_summary.csv"
)

write.csv2(
    cluster_pk_res$similarity_matrix,
    "MetaProViz_Results/Enrichment/similarity_matrix.csv"
)


```





# Session information

```{r session_info, echo=FALSE}
options(width = 120)
sessionInfo()
```

# Bibliography
