---
title: "Enrichment Analysis"
author:
  - name: Christina Schmidt
    affiliation:
    - Heidelberg University
  - name: Jannik Franken
    affiliation:
    - Heidelberg University
output:
  html_document:
    self_contained: true
    toc: true
    toc_float: true
    toc_depth: 5
    code_folding: show
vignette: >
  %\VignetteIndexEntry{Standard Metabolomics}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
bibliography: bibliography.bib
editor_options:
  chunk_output_type: console
---

<style>
.vscroll-plot {
    width: 850px;
    height: 500px;
    overflow-y: scroll;
    overflow-x: hidden;
}
</style>

```{r chunk_setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

```{r message=FALSE, warning=FALSE}
# 1. Install Rtools if you haven’t done this yet, using the appropriate version (e.g.windows or macOS).
# 2. Install the latest development version from GitHub using devtools
# devtools::install_github("https://github.com/saezlab/MetaProViz")

library(MetaProViz)

# dependencies that need to be loaded:
library(magrittr)
library(dplyr)
library(rlang)
library(tidyr)
library(tibble)
library(stringr)
library(readxl)
```

```{r}
# Make folder for results: @Jannik add folders as needed!
if (!dir.exists("MetaProViz_Results/Enrichment")) {
  dir.create("MetaProViz_Results/Enrichment", recursive = TRUE)
}

```

```{r}
# Convert All Text to UTF-8
sanitize_df <- function(df) {
  df[] <- lapply(df, function(col) {
    if (is.character(col)) {
      col <- ifelse(is.na(col), NA, col)  # Ensure NAs remain NAs
      col <- iconv(col, from = "UTF-8", to = "UTF-8", sub = "")  # Remove non-UTF-8 characters
      col <- gsub("[[:cntrl:]]", "", col)  # Remove control characters (but not NA)
      col <- substr(col, 1, 30000)  # Truncate long cells
    }
    return(col)
  })
  return(df)
}

# Open Excel file from BioRxiv branch and read in the data
read_xlsx_all <- function(url) {
  tmp <- tempfile(fileext = ".xlsx")
  download.file(url, tmp, mode = "wb")
  
  sheets <- readxl::excel_sheets(tmp)
  lapply(sheets, function(s) read_excel(tmp, sheet = s)) |>
    setNames(sheets)
}


```

# Load results
## 1. Loading the Feature metadata
Result from Fig2_FeatureMetadata, which include the feature metadata table with extended metabolite IDs and solved ambiguity issues for mapping KEGG IDs to KEGG pathways.
@Jannik

```{r}
# load Results from Fig2_FeatureMetadata

Tissue_MetaData_Extended_cleaned <- 
    read.csv2(
        "MetaProViz_Results/PK/End_Result_Fig2/Tissue_MetaData_Extended_cleaned.csv"
    )

```


## 2. Biological regulated clustering Results
@Jannik: Load Extended Data Table 1 , Load Sheet 9 from BiorXiv

```{r}
Extended_Data_Table_1 <- read_xlsx_all(
  "https://raw.githubusercontent.com/saezlab/MetaProViz/BioRxiv/Extended%20Data%20Table%201_PatientsTissue_Metabolomics.xlsx"
)

# names(Extended_Data_Table_1)
DMA_TUMOR_vs_NORMAL_Old <- Extended_Data_Table_1$"DMA_TUMOR_vs_NORMAL_Old"
DMA_TUMOR_vs_NORMAL_Young <- Extended_Data_Table_1$"DMA_TUMOR_vs_NORMAL_Young"
MCA2Cond_Old_Young <- Extended_Data_Table_1$"MCA2Cond_Old-Young"
```



# Pathway enrichment
@Jannik: This needs to be rerun --> new sheet 11 and 12
Below is the previously used code:


Next, we perform Over Representation Analysis (`ORA`) using KEGG pathways for each comparison.\
```{r, warning=FALSE}
# Since we have performed multiple comparisons (all_vs_HK2), we will run ORA for each of this comparison
DM_ORA_res <- list()

KEGG_Pathways <- MetaProViz::metsigdb_kegg()

## manually clean up KEGG_Pathways with the few instances we saw in Fig2 during checkmatch_pk_to_data
## to-do @Jannik


# continue
# comparison = "Old"
for(comparison in c("Old", "Young")){
  
  dma_res <- merge(
      x = Tissue_MetaData_Extended_cleaned,
      y = case_when(
              comparison == "Old" ~ DMA_TUMOR_vs_NORMAL_Old,
              comparison == "Young" ~ DMA_TUMOR_vs_NORMAL_Young
          ),
      by = "Metabolite",
      all = TRUE
  )
  
  # Ensure unique IDs and full background --> we include measured features that do not have a KEGG ID.
  dma_res <- 
      dma_res[,c(3,21:25)] %>%
      mutate(  # remove duplications and keep the higher Log2FC measurement
          InputID_select = if_else(
              is.na(InputID_select),
              paste0("NA_", cumsum(is.na(InputID_select))),
              InputID_select
              )
      ) %>%
      group_by(
          InputID_select
      ) %>%
      slice_max(
          order_by = Log2FC, n = 1, with_ties = FALSE
      ) %>%
      ungroup() %>%
      remove_rownames() %>%
      tibble::column_to_rownames(
          "InputID_select"
      )
  
  
  #Perform ORA
  set.seed(100)
  Res <- MetaProViz::standard_ora(data = dma_res, #Input data requirements: column `t.val` and column `Metabolite`
                                 metadata_info=c(pvalColumn="p.adj", percentageColumn="t.val", PathwayTerm= "term", PathwayFeature= "MetaboliteID"),
                                 input_pathway=KEGG_Pathways,#Pathway file requirements: column `term`, `Metabolite` and `Description`. Above we loaded the Kegg_Pathways using MetaProViz::Load_KEGG()
                                 pathway_name=paste0("KEGG_", comparison, sep=""),
                                 min_gssize=3,
                                 max_gssize=1000,
                                 cutoff_stat=0.01,
                                 cutoff_percentage=10)

  DM_ORA_res[[comparison]] <- Res

  #Select to plot:
  Res_Select <- Res[["ClusterGosummary"]]%>%
    filter(p.adjust<0.1)%>%
    #filter(pvalue<0.05)%>%
    filter(percentage_of_Pathway_detected>10)

  if(is.null(Res_Select)==FALSE){
      MetaProViz::viz_volcano(plot_types="PEA",
                       data= dma_res, #Must be the data you have used as an input for the pathway analysis
                       data2=as.data.frame(Res_Select)%>%dplyr::rename("term"="ID"),
                       metadata_info= c(PEA_Pathway="term",# Needs to be the same in both, metadata_feature and data2.
                                       PEA_stat="p.adjust",#Column data2
                                       PEA_score="GeneRatio",#Column data2
                                       PEA_Feature="MetaboliteID"),# Column metadata_feature (needs to be the same as row names in data)
                       metadata_feature= KEGG_Pathways,#Must be the pathways used for pathway analysis
                       plot_name= paste("KEGG_", comparison, sep=""),
                       select_label = NULL,
                       subtitle= "PEA" )
  }
}
```

## Use the results to showcase networks
@Jannik here would be cluster_pk


# Save new Extended Data Table
@Jannik now you would create the complete new Extended Datatable. We will need to include some of my results too, so we can discuss here once you have the previous order.





# Session information

```{r session_info, echo=FALSE}
options(width = 120)
sessionInfo()
```

# Bibliography
