---
title: "Sample Metadata Analysis"
author:
  - name: Christina Schmidt
    affiliation:
    - Heidelberg University
output:
  html_document:
    self_contained: true
    toc: true
    toc_float: true
    toc_depth: 5
    code_folding: show
vignette: >
  %\VignetteIndexEntry{Standard Metabolomics}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
bibliography: bibliography.bib
editor_options:
  chunk_output_type: console
---

<style>
.vscroll-plot {
    width: 850px;
    height: 500px;
    overflow-y: scroll;
    overflow-x: hidden;
}
</style>

```{r chunk_setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

```{r message=FALSE, warning=FALSE}
# 1. Install Rtools if you haven’t done this yet, using the appropriate version (e.g.windows or macOS).
# 2. Install the latest development version from GitHub using devtools
# devtools::install_github("https://github.com/saezlab/MetaProViz")

#remotes::install_github('saezlab/OmnipathR@devel')
#remotes::install_github("saezlab/MetaProViz@devel")
library(MetaProViz)

# dependencies that need to be loaded:
library(magrittr)
library(dplyr)
library(rlang)
library(tidyr)
library(tibble)
library(stringr)
```


# 1. Loading the data
Here we choose an example datasets, which is publicly available in the [paper](https://www.cell.com/cancer-cell/comments/S1535-6108(15)00468-7#supplementaryMaterial) "An Integrated Metabolic Atlas of Clear Cell Renal Cell Carcinoma", which includes metabolomic profiling on 138 matched clear cell renal cell carcinoma (ccRCC)/normal tissue pairs [@Hakimi2016]. Metabolomics was done using The company Metabolon, so this is untargeted metabolomics.\
We have updated the PubChem IDs as there were PubChem entries in the excel file that are not PubChem IDs. It is likely that this is an error caused by excel conversion as it seems in some cases the first part of the ID corresponds to one of the other IDs, so maybe these where multiple potential PubChem IDs.To see details on how this was done check out our example data branch: https://github.com/saezlab/MetaProViz/blob/ExampleData/ccRCC_Hakimi-et-all.html

Load the feature metadata:
```{r}
load("C:/Users/chris/OneDrive/docs/Desktop/tissue_meta.rda")
#data(tissue_meta)

MetaboliteIDs <- tissue_meta %>%
  dplyr::filter(!stringr::str_detect(Metabolite, "^X\\s*-\\s*\\d+$"))%>%# only keep entries with trivial names
   dplyr::mutate(
    SUPER_PATHWAY = dplyr::coalesce(SUPER_PATHWAY, "None"),
    SUB_PATHWAY   = dplyr::coalesce(SUB_PATHWAY, "None")
  )
```

# MetaProViz Metabolite ID workflow
```{r}
# Make folder for results:
if (!dir.exists("MetaProViz_Results/PK")) {
  dir.create("MetaProViz_Results/PK", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/Original")) {
  dir.create("MetaProViz_Results/PK/Original", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PlusTranslatedIDs_FromCAS")) {
  dir.create("MetaProViz_Results/PK/PlusTranslatedIDs_FromCAS", recursive = TRUE)
}

##------------------------------------------------------------------
# Individual DBs
## HMDB
if (!dir.exists("MetaProViz_Results/PK/PK_HMDB")) {
  dir.create("MetaProViz_Results/PK/PK_HMDB", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PK_HMDB/Original")) {
  dir.create("MetaProViz_Results/PK/PK_HMDB/Original", recursive = TRUE)
}

## KEGG
if (!dir.exists("MetaProViz_Results/PK/PK_KEGG")) {
  dir.create("MetaProViz_Results/PK/PK_KEGG", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PK_KEGG/Original")) {
  dir.create("MetaProViz_Results/PK/PK_KEGG/Original", recursive = TRUE)
}

## PUBCHEM
if (!dir.exists("MetaProViz_Results/PK/PK_PUBCHEM")) {
  dir.create("MetaProViz_Results/PK/PK_PUBCHEM", recursive = TRUE)
}

if (!dir.exists("MetaProViz_Results/PK/PK_PUBCHEM/Original")) {
  dir.create("MetaProViz_Results/PK/PK_PUBCHEM/Original", recursive = TRUE)
}
```

## Original Feature Space Exploration
1. Present metabolite IDs:
```{r, warning=FALSE}
ccRCC_CompareIDs <- MetaProViz::compare_pk(data = list(Biocft = MetaboliteIDs |> dplyr::rename("Class"="SUPER_PATHWAY")),
                                           name_col = "Metabolite",
                                           metadata_info = list(Biocft = c("KEGG", "HMDB", "PUBCHEM")),
                                           plot_name = "Overlap of ID types in ccRCC data",
                                           path= "MetaProViz_Results/PK/Original")
```
\
Here we notice that 76 features have no metabolite ID assigned, yet have a trivial name and a metabolite class assigned. For 135 metabolites we only have a pubchem ID, yet no HMDB or KEGG ID. 

2. Metabolite IDs per feature:
2.1. HMDB:
```{r}
Plot1_HMDB <- MetaProViz::count_id(MetaboliteIDs, 
                                   delimiter =",",
                                   "HMDB",
                                   fill_colors = c(`No ID` = "#006384", `Single ID` = "#9F0162", `Multiple IDs` = "#BEBEBE"),
                                   binwidth = 1,
                                   path = "MetaProViz_Results/PK/PK_HMDB/Original")
```

2.2. KEGG:
```{r}
Plot1_KEGG <- MetaProViz::count_id(MetaboliteIDs, 
                                   delimiter =",",
                                   fill_colors = c(`No ID` = "#006384", `Single ID` = "#9F0162", `Multiple IDs` = "#BEBEBE"),
                                   "KEGG",
                                   path= "MetaProViz_Results/PK/PK_KEGG/Original")
```

2.3.: PubChem:
```{r}
Plot1_PubChem <- MetaProViz::count_id(MetaboliteIDs,
                                   delimiter ="; ",
                                   fill_colors = c(`No ID` = "#006384", `Single ID` = "#9F0162", `Multiple IDs` = "#BEBEBE"),
                                   "PUBCHEM",
                                   path= "MetaProViz_Results/PK/PK_PUBCHEM/Original")
```

3. Features with no HMDB, KEGG or PubChem ID
```{r}
# Extract features with No HMDB, KEGG or PubChem ID:
features_no_ids <- MetaboliteIDs %>%
  dplyr::filter(
    (is.na(HMDB)    | trimws(HMDB) %in% c("", "NA")) &
    (is.na(KEGG)    | trimws(KEGG) %in% c("", "NA")) &
    (is.na(PUBCHEM) | trimws(PUBCHEM) %in% c("", "NA"))
  )

#Of those some may have a CAS number:
features_no_ids_with_cas <- features_no_ids %>%
  dplyr::filter(!(is.na(CAS) | trimws(CAS) %in% c("", "NA")))

print(paste0("We have ", nrow(features_no_ids_with_cas), " features with CAS number but no other metabolite ID and ", nrow(features_no_ids)- nrow(features_no_ids_with_cas), " features with no CAS number and no other metabolite ID."))
```

## Translate Cas to metabolite ID
There are some cases in which we have a CAS number, yet no HMDB, PubChem or KEGG ID. We can use the CAS number to check if we can find a metabolite ID for these features.\
```{r}
MetaProViz::translate_id(data = features_no_ids, 
                         metadata_info = list(InputID = "PUBCHEM", grouping_variable = "SUPER_PATHWAY"), 
                         from = "CAS", 
                         to = c("hmdb", "kegg", "pubchem"), 
                         path="MetaProViz_Results/PK/PlusTranslatedIDs_FromCAS")

```
\
We will repeat the `count_id` and `compare_pk` plot after every change to the number of metabolite IDs:
```{r}


```
\
We will start to compile a summary table that will include the number of IDs  (1,2,n) after each step taken to increase the feature space. So now we will add the The CAS IDs
```{r}


```

## Traverse ID


## 

```{r}



# Add equivalent IDs:
tissue_meta_AddIDs_HMDB <- MetaProViz::equivalent_id(data= Input_HMDB,
                                                           metadata_info = c(InputID="HMDB"),# ID in the measured data, here we use the HMDB ID
                                                           from = "hmdb",
                                                     path= "MetaProViz_Results/PK/PK_HMDB/PlusEquivalentIDs")


```


## Load the differential metabolite results and add new IDs
Load differential metabolite results from BioRXiv branch and add the new KEGG IDs. Afterwards we can rerun ORA and biological regulated clustering using the updated feature space.

### ORA
We can perform Over Representation Analysis (`ORA`) using KEGG pathways for each comparison and plot significant pathways. Noteworthy, since not all metabolites have KEGG IDs, we will lose information.\
\
Given that in some cases we have multiple KEGG IDs for a measured feature, we will check if this causes mapping to multiple, different entries in the KEGG pathway-metabolite sets:
```{r}
#Load Kegg pathways:
KEGG_Pathways <- metsigdb_kegg()

#check mapping with metadata

ccRCC_to_KEGGPathways <- MetaProViz::checkmatch_pk_to_data(data = Tissue_MetaData_Extended,
                                                           input_pk = KEGG_Pathways,
                                                           metadata_info = c(InputID = "KEGG", PriorID = "MetaboliteID", grouping_variable = "term"))

problems_terms <- ccRCC_to_KEGGPathways[["GroupingVariable_summary"]]%>%
  filter(!Group_Conflict_Notes== "None")
```
```{r, echo=FALSE}
problems_terms[,c(1,5:8)]%>%
  kableExtra::kbl(caption = "Terms in KEGG pathways where the same measured feature mapps with more than one ID, which will inflate the enrichment analysis.") %>%
  kableExtra::kable_classic(full_width = F, html_font = "Cambria", font_size = 12)
```
\
Dependent on the biological question and the organism and prior knowledge, one can either maintain the metabolite ID of the more likely metabolite (e.g. in human its moire likely that we have L-aminoacid than D-aminoacid) or the metabolite ID that is represented in more/less pathways (specificity). If we are looking into the cases where we do have multiple IDs, for cases with no match to the prior knowledge we can just maintain one ID, whilst for cases with exactly one match we should maintain the ID that is found in the prior knowledge. 
```{r}
# Select the cases where a feature has multiple IDs
multipleIDs <- ccRCC_to_KEGGPathways[["data_summary"]]%>%
  filter(original_count>1)
```
```{r, echo=FALSE}
multipleIDs[,c(1,12, 5:6, 3, 9, 11, 13)]%>%
  kableExtra::kbl(caption = "Terms in KEGG pathways where the same measured feature mapps with more than one ID, which may inflate the enrichment analysis.") %>%
  kableExtra::kable_classic(full_width = F, html_font = "Cambria", font_size = 12)
```
\
In case of `ActionRequired=="Check"`, we can look into the column `Action_Specific` which contains additional information. In case of the entry `KeepEachID`, multiple matches to the prior knowledge were found, yet the features are in different pathways (=GroupingVariable). Yet, in case of `KeepOneID`, the different IDs map to the same pathway in the prior knowledge for at least one case and therefore keeping both would inflate the enrichment analysis.\
```{r}
SelectedIDs <- ccRCC_to_KEGGPathways[["data_summary"]]%>%
  #Expand rows where Action == KeepEachID by splitting `matches`
  dplyr::mutate(matches_split = if_else(Action_Specific == "KeepEachID", matches, NA_character_)) %>%
  separate_rows(matches_split, sep = ",\\s*") %>%
  mutate(InputID_select = if_else(Action_Specific  == "KeepEachID", matches_split, InputID_select)) %>%
  select(-matches_split) %>%
  #Select one ID for AcionSpecific==KeepOneID
  dplyr::mutate(InputID_select = case_when(
    Action_Specific == "KeepOneID" & matches == "C03460, C03722" ~ "C03722", # 2-Methylprop-2-enoyl-CoA versus Quinolinate. No evidence, hence we keep the one present in more pathways ( C03722=7 pathways, C03460=2 pathway)
    Action_Specific == "KeepOneID" & matches ==  "C00221, C00031" ~ "C00031", # These are D- and L-Glucose. We have human samples, so in this conflict we will maintain L-Glucose
    Action_Specific == "KeepOneID" & matches ==  "C17737, C00695" ~ "C00695", # Allocholic acid versus Cholic acid. No evidence, hence we keep the one present in more pathways (C00695 = 4 pathways, C17737 = 1 pathway)
    Action_Specific == "KeepOneID" ~ InputID_select,  # Keep NA where not matched manually
    TRUE ~ InputID_select
  )) 
```
\
Lastly, we need to add the column including our selected IDs to the metadata table:
```{r}
Tissue_MetaData_Extended <- merge(x= SelectedIDs %>%
                                        dplyr::select(KEGG, InputID_select),
                                  y= Tissue_MetaData_Extended,
                                  by= "KEGG",
                                  all.y=TRUE)
```
\
For results with p.adjusted value < 0.1 and a minimum of 10% of the pathway detected will be visualized as Volcano plots:\
```{r, eval=FALSE}
# Since we have performed multiple comparisons (all_vs_HK2), we will run ORA for each of this comparison
DM_ORA_res<- list()

for(comparison in names(ResList)){#Res list includes the different comparisons we performed above <-
  #Ensure that the Metabolite names match with KEGG IDs or KEGG trivial names.
  dma_res <- merge(x= Tissue_MetaData_Extended,
                   y= ResList[[comparison]][["dma"]][["TUMOR_vs_NORMAL"]], 
                   by="Metabolite", 
                   all=TRUE)
  
  #Ensure unique IDs and full background --> we include measured features that do not have a KEGG ID.
  dma_res <- dma_res[,c(3,21:25)]%>%
    dplyr::mutate(InputID_select = if_else(
      is.na(InputID_select),
      paste0("NA_", cumsum(is.na(InputID_select))),
      InputID_select
      ))%>% #remove duplications and keep the higher Log2FC measurement
    group_by(InputID_select) %>%
    slice_max(order_by = Log2FC, n = 1, with_ties = FALSE) %>%
    ungroup()%>%
    remove_rownames()%>%
    tibble::column_to_rownames("InputID_select")
  
  #Perform ORA
  set.seed(100)
  Res <- MetaProViz::standard_ora(data= dma_res, #Input data requirements: column `t.val` and column `Metabolite`
                                 metadata_info=c(pvalColumn="p.adj", percentageColumn="t.val", PathwayTerm= "term", PathwayFeature= "MetaboliteID"),
                                 input_pathway=KEGG_Pathways,#Pathway file requirements: column `term`, `Metabolite` and `Description`. Above we loaded the Kegg_Pathways using MetaProViz::Load_KEGG()
                                 pathway_name=paste0("KEGG_", comparison, sep=""),
                                 min_gssize=3,
                                 max_gssize=1000,
                                 cutoff_stat=0.01,
                                 cutoff_percentage=10)

  DM_ORA_res[[comparison]] <- Res

  #Select to plot:
  Res_Select <- Res[["ClusterGosummary"]]%>%
    filter(p.adjust<0.1)%>%
    #filter(pvalue<0.05)%>%
    filter(percentage_of_Pathway_detected>10)

  if(is.null(Res_Select)==FALSE){
     MetaProViz::viz_volcano(plot_types="PEA",
                       data= dma_res, #Must be the data you have used as an input for the pathway analysis
                       data2=as.data.frame(Res_Select )%>%dplyr::rename("term"="ID"),
                       metadata_info= c(PEA_Pathway="term",# Needs to be the same in both, metadata_feature and data2.
                                       PEA_stat="p.adjust",#Column data2
                                       PEA_score="GeneRatio",#Column data2
                                       PEA_Feature="MetaboliteID"),# Column metadata_feature (needs to be the same as row names in data)
                       metadata_feature= KEGG_Pathways,#Must be the pathways used for pathway analysis
                       plot_name= paste("KEGG_", comparison, sep=""),
                       subtitle= "PEA" )
  }
}
```
\
<div class="vscroll-plot">
```{r, echo=FALSE, warning=FALSE, fig.align="left", fig.width=7, fig.height=5}
# Since we have performed multiple comparisons (all_vs_HK2), we will run ORA for each of this comparison
DM_ORA_res<- list()

for(comparison in names(ResList)){#Res list includes the different comparisons we performed above <-
  #Ensure that the Metabolite names match with KEGG IDs or KEGG trivial names.
  dma_res <- merge(x= Tissue_MetaData_Extended,
                   y= ResList[[comparison]][["dma"]][["TUMOR_vs_NORMAL"]], 
                   by="Metabolite", 
                   all=TRUE)
  
  #Ensure unique IDs and full background --> we include measured features that do not have a KEGG ID.
  dma_res <- dma_res[,c(3,21:25)]%>%
    dplyr::mutate(InputID_select = if_else(
      is.na(InputID_select),
      paste0("NA_", cumsum(is.na(InputID_select))),
      InputID_select
      ))%>% #remove duplications and keep the higher Log2FC measurement
    group_by(InputID_select) %>%
    slice_max(order_by = Log2FC, n = 1, with_ties = FALSE) %>%
    ungroup()%>%
    remove_rownames()%>%
    tibble::column_to_rownames("InputID_select")
  
  #Perform ORA
  set.seed(100)
  Res <- MetaProViz::standard_ora(data= dma_res, #Input data requirements: column `t.val` and column `Metabolite`
                                 metadata_info=c(pvalColumn="p.adj", percentageColumn="t.val", PathwayTerm= "term", PathwayFeature= "MetaboliteID"),
                                 input_pathway=KEGG_Pathways,#Pathway file requirements: column `term`, `Metabolite` and `Description`. Above we loaded the Kegg_Pathways using MetaProViz::Load_KEGG()
                                 pathway_name=paste0("KEGG_", comparison, sep=""),
                                 min_gssize=3,
                                 max_gssize=1000,
                                 cutoff_stat=0.01,
                                 cutoff_percentage=10)

  DM_ORA_res[[comparison]] <- Res

  #Select to plot:
  Res_Select <- Res[["ClusterGosummary"]]%>%
    filter(p.adjust<0.1)%>%
    #filter(pvalue<0.05)%>%
    filter(percentage_of_Pathway_detected>10)

  if(is.null(Res_Select)==FALSE){
     MetaProViz::viz_volcano(plot_types="PEA",
                       data= dma_res, #Must be the data you have used as an input for the pathway analysis
                       data2=as.data.frame(Res_Select )%>%dplyr::rename("term"="ID"),
                       metadata_info= c(PEA_Pathway="term",# Needs to be the same in both, metadata_feature and data2.
                                       PEA_stat="p.adjust",#Column data2
                                       PEA_score="GeneRatio",#Column data2
                                       PEA_Feature="MetaboliteID"),# Column metadata_feature (needs to be the same as row names in data)
                       metadata_feature= KEGG_Pathways,#Must be the pathways used for pathway analysis
                       plot_name= paste("KEGG_", comparison, sep=""),
                       subtitle= "PEA" )
  }
}
```
</div>
\
\
\

### Biological regulated clustering
to understand which metabolites are changing independent of the patients age, hence only due to tumour versus normal, and which metabolites change independent of tumour versus normal, hence due to the different age,
we can use the `MetaProViz::mca_2cond()` function.\
Metabolite Clustering Analysis (`MCA`) enables clustering of metabolites into groups based on logical regulatory rules. Here we set two different thresholds, one for the differential metabolite abundance (Log2FC) and one for the significance (e.g. p.adj). This will define if a feature (= metabolite) is assigned into:\
1. “UP”, which means a metabolite is significantly up-regulated in the underlying comparison.\
2. “DOWN”, which means a metabolite is significantly down-regulated in the underlying comparison.\
3. “No Change”, which means a metabolite does not change significantly in the underlying comparison and/or is not defined as up-regulated/down-regulated based on the Log2FC threshold chosen.\
\
Thereby “No Change” is further subdivided into four states:\
1. “Not Detected”, which means a metabolite is not detected in the underlying comparison.\
2. “Not Significant”, which means a metabolite is not significant in the underlying comparison.\
3. “Significant positive”, which means a metabolite is significant in the underlying comparison and the differential metabolite abundance is positive, yet does not meet the threshold set for “UP” (e.g. Log2FC >1 = “UP” and we have a significant Log2FC=0.8).\
4. “Significant negative”, which means a metabolite is significant in the underlying comparison and the differential metabolite abundance is negative, yet does not meet the threshold set for “DOWN”.\
\
For more information you can also check out the other vignettes.
```{r}
MCAres <-  MetaProViz::mca_2cond(data_c1=ResList[["Young"]][["dma"]][["TUMOR_vs_NORMAL"]],
                                 data_c2=ResList[["Old"]][["dma"]][["TUMOR_vs_NORMAL"]],
                                 metadata_info_c1=c(ValueCol="Log2FC",StatCol="p.adj", cutoff_stat= 0.05, ValueCutoff=1),
                                 metadata_info_c2=c(ValueCol="Log2FC",StatCol="p.adj", cutoff_stat= 0.05, ValueCutoff=1),
                                 feature = "Metabolite",
                                 save_table = "csv",
                                 method_background="C1&C2"#Most stringend background setting, only includes metabolites detected in both comparisons
                                 )
```
\
Now we can use this information to colour code our volcano plot. We will plot individual vocano plots for each metabolite pathway as defined by the feature metadata provided as part of the data in [@Hakimi2016].
```{r, eval=FALSE}
# Add metabolite information such as KEGG ID or pathway to results
MetaData_Metab <- merge(x=Tissue_MetaData,
                   y= MCAres[["MCA_2Cond_Results"]][, c(1, 14:15)]%>%tibble::column_to_rownames("Metabolite"),
                   by=0,
                   all.y=TRUE)%>%
  tibble::column_to_rownames("Row.names")

MetaProViz::viz_volcano(plot_types="Compare",
                       data=ResList[["Young"]][["dma"]][["TUMOR_vs_NORMAL"]]%>%tibble::column_to_rownames("Metabolite"),
                       data2= ResList[["Old"]][["dma"]][["TUMOR_vs_NORMAL"]]%>%tibble::column_to_rownames("Metabolite"),
                       name_comparison= c(data="Young", data2= "Old"),
                       metadata_feature =  MetaData_Metab,
                      plot_name= "Young-TUMOR_vs_NORMAL compared to Old-TUMOR_vs_NORMAL",
                       subtitle= "Results of dma",
                       metadata_info = c(individual = "SUPER_PATHWAY",
                                        color = "RG2_Significant"))

MetaProViz::viz_volcano(plot_types="Compare",
                       data=ResList[["Young"]][["dma"]][["TUMOR_vs_NORMAL"]]%>%tibble::column_to_rownames("Metabolite"),
                       data2= ResList[["Old"]][["dma"]][["TUMOR_vs_NORMAL"]]%>%tibble::column_to_rownames("Metabolite"),
                       name_comparison= c(data="Young", data2= "Old"),
                       metadata_feature =  MetaData_Metab,
                      plot_name= "Young-TUMOR_vs_NORMAL compared to Old-TUMOR_vs_NORMAL_Sub",
                       subtitle= "Results of dma",
                       metadata_info = c(individual = "SUB_PATHWAY",
                                        color = "RG2_Significant"))
```
\
<div class="vscroll-plot">
```{r, echo=FALSE, warning=FALSE, fig.align="left", fig.width=7, fig.height=5}
# Add metabolite information such as KEGG ID or pathway to results
MetaData_Metab <- merge(x=Tissue_MetaData,
                   y= MCAres[["MCA_2Cond_Results"]][, c(1, 14:15)]%>%tibble::column_to_rownames("Metabolite"),
                   by=0,
                   all.y=TRUE)%>%
  tibble::column_to_rownames("Row.names")

MetaProViz::viz_volcano(plot_types="Compare",
                       data=ResList[["Young"]][["dma"]][["TUMOR_vs_NORMAL"]]%>%tibble::column_to_rownames("Metabolite"),
                       data2= ResList[["Old"]][["dma"]][["TUMOR_vs_NORMAL"]]%>%tibble::column_to_rownames("Metabolite"),
                       name_comparison= c(data="Young", data2= "Old"),
                       metadata_feature =  MetaData_Metab,
                      plot_name= "Young-TUMOR_vs_NORMAL compared to Old-TUMOR_vs_NORMAL",
                       subtitle= "Results of dma",
                       metadata_info = c(individual = "SUPER_PATHWAY",
                                        color = "RG2_Significant"),
                       color_palette = c(Cond1_DOWN="#0694B9", Cond2_DOWN="#38BEEF",Cond1_UP="#F1E644", Cond2_UP= "#DE963F", core_DOWN="#2F358B", core_UP="#BD1F2D", Opposite="#10AC8D", None="#969697"))

MetaProViz::viz_volcano(plot_types="Compare",
                       data=ResList[["Young"]][["dma"]][["TUMOR_vs_NORMAL"]]%>%tibble::column_to_rownames("Metabolite"),
                       data2= ResList[["Old"]][["dma"]][["TUMOR_vs_NORMAL"]]%>%tibble::column_to_rownames("Metabolite"),
                       name_comparison= c(data="Young", data2= "Old"),
                       metadata_feature =  MetaData_Metab,
                      plot_name= "Young-TUMOR_vs_NORMAL compared to Old-TUMOR_vs_NORMAL_Sub",
                       subtitle= "Results of dma",
                       metadata_info = c(individual = "SUB_PATHWAY",
                                        color = "RG2_Significant"),
                       color_palette = c(Cond1_DOWN="#0694B9", Cond2_DOWN="#38BEEF",Cond1_UP="#F1E644", Cond2_UP= "#DE963F", core_DOWN="#2F358B", core_UP="#BD1F2D", Opposite="#10AC8D", None="#969697"))

```
</div>
\
\
\

### Pathway enrichment
Next, we perform Over Representation Analysis (`ORA`) using KEGG pathways for each comparison.\
```{r, eval=FALSE}
# Since we have performed multiple comparisons (all_vs_HK2), we will run ORA for each of this comparison
DM_ORA_res<- list()

KEGG_Pathways <- MetaProViz::metsigdb_kegg()

for(comparison in names(ResList)){
  dma_res <- merge(x= Tissue_MetaData_Extended,
                   y=ResList[[comparison]][["dma"]][["TUMOR_vs_NORMAL"]], 
                   by="Metabolite", 
                   all=TRUE)
  
  #Ensure unique IDs and full background --> we include measured features that do not have a KEGG ID.
  dma_res <- dma_res[,c(3,21:25)]%>%
    dplyr::mutate(InputID_select = if_else(
      is.na(InputID_select),
      paste0("NA_", cumsum(is.na(InputID_select))),
      InputID_select
      ))%>% #remove duplications and keep the higher Log2FC measurement
    group_by(InputID_select) %>%
    slice_max(order_by = Log2FC, n = 1, with_ties = FALSE) %>%
    ungroup()%>%
    remove_rownames()%>%
    tibble::column_to_rownames("InputID_select")
  
  
  #Perform ORA
  set.seed(100)
  Res <- MetaProViz::standard_ora(data= dma_res, #Input data requirements: column `t.val` and column `Metabolite`
                                 metadata_info=c(pvalColumn="p.adj", percentageColumn="t.val", PathwayTerm= "term", PathwayFeature= "MetaboliteID"),
                                 input_pathway=KEGG_Pathways,#Pathway file requirements: column `term`, `Metabolite` and `Description`. Above we loaded the Kegg_Pathways using MetaProViz::Load_KEGG()
                                 pathway_name=paste0("KEGG_", comparison, sep=""),
                                 min_gssize=3,
                                 max_gssize=1000,
                                 cutoff_stat=0.01,
                                 cutoff_percentage=10)

  DM_ORA_res[[comparison]] <- Res

  #Select to plot:
  Res_Select <- Res[["ClusterGosummary"]]%>%
    filter(p.adjust<0.1)%>%
    #filter(pvalue<0.05)%>%
    filter(percentage_of_Pathway_detected>10)

  if(is.null(Res_Select)==FALSE){
      MetaProViz::viz_volcano(plot_types="PEA",
                       data= dma_res, #Must be the data you have used as an input for the pathway analysis
                       data2=as.data.frame(Res_Select)%>%dplyr::rename("term"="ID"),
                       metadata_info= c(PEA_Pathway="term",# Needs to be the same in both, metadata_feature and data2.
                                       PEA_stat="p.adjust",#Column data2
                                       PEA_score="GeneRatio",#Column data2
                                       PEA_Feature="MetaboliteID"),# Column metadata_feature (needs to be the same as row names in data)
                       metadata_feature= KEGG_Pathways,#Must be the pathways used for pathway analysis
                       plot_name= paste("KEGG_", comparison, sep=""),
                       select_label = NULL,
                       subtitle= "PEA" )
  }
}
```
\
<div class="vscroll-plot">
```{r, echo=FALSE, warning=FALSE, fig.align="left", fig.width=7, fig.height=5}
# Since we have performed multiple comparisons (all_vs_HK2), we will run ORA for each of this comparison
DM_ORA_res<- list()

KEGG_Pathways <- MetaProViz::metsigdb_kegg()

for(comparison in names(ResList)){
  dma_res <- merge(x= Tissue_MetaData_Extended,
                   y=ResList[[comparison]][["dma"]][["TUMOR_vs_NORMAL"]], 
                   by="Metabolite", 
                   all=TRUE)
  
  #Ensure unique IDs and full background --> we include measured features that do not have a KEGG ID.
  dma_res <- dma_res[,c(3,21:25)]%>%
    dplyr::mutate(InputID_select = if_else(
      is.na(InputID_select),
      paste0("NA_", cumsum(is.na(InputID_select))),
      InputID_select
      ))%>% #remove duplications and keep the higher Log2FC measurement
    group_by(InputID_select) %>%
    slice_max(order_by = Log2FC, n = 1, with_ties = FALSE) %>%
    ungroup()%>%
    remove_rownames()%>%
    tibble::column_to_rownames("InputID_select")
  
  
  #Perform ORA
  set.seed(100)
  Res <- MetaProViz::standard_ora(data= dma_res, #Input data requirements: column `t.val` and column `Metabolite`
                                 metadata_info=c(pvalColumn="p.adj", percentageColumn="t.val", PathwayTerm= "term", PathwayFeature= "MetaboliteID"),
                                 input_pathway=KEGG_Pathways,#Pathway file requirements: column `term`, `Metabolite` and `Description`. Above we loaded the Kegg_Pathways using MetaProViz::Load_KEGG()
                                 pathway_name=paste0("KEGG_", comparison, sep=""),
                                 min_gssize=3,
                                 max_gssize=1000,
                                 cutoff_stat=0.01,
                                 cutoff_percentage=10)

  DM_ORA_res[[comparison]] <- Res

  #Select to plot:
  Res_Select <- Res[["ClusterGosummary"]]%>%
    filter(p.adjust<0.1)%>%
    #filter(pvalue<0.05)%>%
    filter(percentage_of_Pathway_detected>10)

  if(is.null(Res_Select)==FALSE){
  MetaProViz::viz_volcano(plot_types="PEA",
                       data= dma_res, #Must be the data you have used as an input for the pathway analysis
                       data2=as.data.frame(Res_Select)%>%dplyr::rename("term"="ID"),
                       metadata_info= c(PEA_Pathway="term",# Needs to be the same in both, metadata_feature and data2.
                                       PEA_stat="p.adjust",#Column data2
                                       PEA_score="GeneRatio",#Column data2
                                       PEA_Feature="MetaboliteID"),# Column metadata_feature (needs to be the same as row names in data)
                       metadata_feature= KEGG_Pathways,#Must be the pathways used for pathway analysis
                       plot_name= paste("KEGG_", comparison, sep=""),
                       select_label = NULL,
                       subtitle= "PEA" )
  }
}
```
</div>
\
\
\


## Use the results to showcase networks



# Session information

```{r session_info, echo=FALSE}
options(width = 120)
sessionInfo()
```

# Bibliography
